### 2013-10-18 publications/forecasting_future_oil_production/peak_oil.tex
Added Zalan's code and publication.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.2 (15/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

\usepackage{lipsum} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage{graphicx}

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}}
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyhead[C]{Forecasting future oil production $\bullet$ July 2013} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Forecasting future oil production}} % Article title

\author{
\large
%\textsc{John Smith}\thanks{A thank you or further information}, % Your name
\textsc{Z. Forr\'o}, % Your name
\textsc{P. Cauwels},
\textsc{L. Fi\'evet}, 
\textsc{M. del Degan} and 
\textsc{D. Sornette}\\[2mm] 
\normalsize D-MTEC, ETHZ, Z\"urich \\ % Your institution
%\normalsize \href{mailto:zforro@ethz.ch}{zforro@ethz.ch} % Your email address
\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

\thispagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

\noindent We  have developed a Monte-Carlo methodology to forecast the crude oil production of Norway, the U.K. and the U. S. based on the current/past performances of individual oil fields. By extrapolating the future production of these fields and the frequency of new discoveries, we are able to forecast the oil production of these countries. Our results indicate that standard methodologies tend to underestimate remaining oil reserves. We compare our model to those methodologies by making predictions from various times points in the past (up to 20 years back for the US). It shows that our model gives a better description (hopefully soon) of the evolution of the oil production.
\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

\section{Introduction}

\lettrine[nindent=0em,lines=3]
Forecasting future oil production has been a topic of active interest since the beginning of the past century given oil's central role in our economy. Its importance ranges from energy production, through manufacturing to pharmaceuticals industry. Petroleum is a non-renewable, finite resource.  It is primordial to be able to forecast future oil production since a misestimation of its reserves can have huge consequences on our society. \\
The methodology behind forecasting future oil production has not evolved much since Hubbert who, in 1956, famously predicted that the U.S. oil production would peak around 1965-1970. That prediction proved to be correct. His main argument was based on the finiteness of oil reserves and used the logistic differential equation  (equation \ref{logistic}) to model oil production. The logistic differential equation is characterized by an initial exponential growth, the growth decreasing to zero as the total oil extracted reaches saturation (no more oil is to be found).

\begin{equation} \label{logistic}
\frac{dP}{dt} = r P \left( 1 - \frac{P}{K} \right)
\end{equation}

\noindent $r$ is commonly referred to as the growth rate, and $K$ as the carrying capacity (total quantity of oil extracted). If $P(t)$ is the the amount of oil extracted up to time $t$, then $\frac{dP}{dt}$ is the oil production rate, the quantity that M. King Hubbert predicted would peak with surprising accuracy. From a methodological point of view, the Hubbert model has enjoyed a longstanding popularity in modeling future oil production given its simplicity. \\
\noindent In this paper, we introduce a new methodology to forecast future oil production. Instead of taking the aggregate oil production profile and fitting it with the Hubbert curve or its variants (such as the multi cyclic Hubbert curve), we use the production profile of each individual oil field. By extending their production into the future and extrapolating the future rate of discovery of new fields, we are able to forecast future oil production by the means of a Monte Carlo simulation. To demonstrate the universality of the methodology presented here, we  apply it to 3 major oil producing countries: Norway, the U.K. and the U.S.

%------------------------------------------------

\section{Methodology}

The idea behind the methodology is to model the future aggregate oil production of a country by studying the production dynamics of its individual constituents, the oil fields. The main benefit of this approach, compared to working directly with aggregate production data, is the possibility to forecast non-trivial oil production profiles arising from the combination of all the individual fields' dynamics. In order to achieve that, one must be able 1) to extend the oil production of each individual field into the future and 2) to extrapolate the rate of discoveries of new oil fields.

\subsection{Extending the oil production of individual fields}
To be able to forecast the oil production of each individual field, regularity needs to be found in the production's dynamics. Modeling the whole production profile from the beginning of extraction seems elusive due to the variety of the forms it can take . Fortunately, modeling the decay process is sufficient in order to extrapolate future oil production. A preliminary classification is necessary to achieve that goal. Figure \ref{regular_and_irregular} shows that independent of the country, oil fields can be classified into 2 main categories. %4 categories.

\begin{itemize}
\item Regular fields. The fields whose decay show some regularity. 
\item Irregular fields. Irregular fields are the ones that don't decay in a regular fashion or that don't decay yet. As such, there is no easy way to forecast their future oil production based on past data.
%\item Empty fields. Empty fields have no more oil production.
%tem Incomplete fields. Incomplete fields are the ones which don't have enough data (because they are too new usually).
\end{itemize}

\begin{figure}[H]
    \includegraphics[width=0.5\textwidth]{regular_and_irregular.pdf}
    \caption{Example of a regular and an irregular field.}
    \label{regular_and_irregular}
\end{figure}

\noindent As of January 2013, regular fields make up 85\%, 87\% and z\% of the number of fields and 94\%, 71\% and z'\% of the total oil volume in the Norway, the U.K. and the U.S. respectively. As such, being able to model them is crucial. As can be seen on figure \ref{stretched_exponential}, the stretched exponential (equation \ref{stretched_exponential}) is a good functional form to fit the decay process of regular fields.

\begin{equation} \label{stretched_exponential}
P(t) = P_0 e^{\left (\frac{t}{\tau} \right)^\beta}
\end{equation}

For the minority of irregular fields, we simply modeled their decay as follows. We picked $\tau$ to be the average $\tau$ over the regular fields. We then fixed $\beta$ so that the sum of the fields' production over its lifetime be equal to the official ultimate recovery estimates, when such an estimate is available. %More details can be found in the supplementary materials.

\begin{figure}[H]
    \includegraphics[width=0.5\textwidth]{stretched_exponential.pdf}
    \caption{Stretched exponential fit to the production of Welton. This functional form fits the data well.}
    \label{stretched_exponential}
\end{figure}

\subsection{Discovery rate of new fields}
Knowing the future production rate of existing fields is not enough; new fields will be discovered in the future. The model describing the discovery rate of new fields should satisfy two fundamental observations.

\begin{enumerate}
\item The rate of new discoveries should tend to zero as time goes to infinity. This is a consequence of the finiteness of the number of oil fields. 
\item The rate of new discoveries should depend on the size of the oil fields. As of today, giant oil fields are discovered much less frequently than dwarfs.
\end{enumerate}

A natural choice for such a model is a non-homogenous poisson process. The Poisson process is a process that generates independent events at  a rate $\lambda$. It is inhomogeneous if the rate is time-dependent, $\lambda \rightarrow \lambda(t)$. The standard way to measure $\lambda(t)$ is to find a functional form for $N(t)$, the number of events (discoveries) up to time $t$. Then, $\lambda(t)$ is simply $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery} shows $N(t)$ for Norwegian fields classified according to their size.The logistic curve is a good fit to the data (integral form of equation \ref{logistic}). This implies that after an initial increase, the rate of new discoveries reaches a peak followed by a decrease until no more oil fields are to be found, consistent with our fundamental observations. The results are qualitatively similar for the U.K. and the U.S.

\begin{figure}[H]
    \includegraphics[width=0.5\textwidth]{rate_of_discoveries.pdf}
    \caption{Logistic fit to the number of discoveries for Norway. The discovery rate of new oil fields is dependent on their size. The logistic model suggests that the majority of small oil  fields have not yet been discovered, while the rate of discovery of medium oil fields is in its decreasing phase. Big oil fields are not expected to be discovered anymore. }
    \label{rate_of_discovery}
\end{figure}

%------------------------------------------------

\section{Results}

Simulating future oil production was straightforward. For each country, new oil fields were generated according to the poisson process, with $\lambda(t)$ chosen according to their size. For each of the generated oil fields, its production profile was chosen randomly from one of existing fields within the same size category. Figure \ref{no_uk_2013} shows the average of 1000 (50 on these plots) simulations. For each country, we can immediately notice the non-symmetric shape of the production's dynamics contrary to what Hubbert would have predicted.

\begin{table}[H]
\caption{Remaining oil reserves}
\centering
\begin{tabular}{lllr}
%\toprule
%\multicolumn{2}{c}{Name} \\
%\cmidrule(r){1-2}
Country & Hubbert & our model & $\Delta$ \\
\midrule
Norway & $81 \cdot 10^6$ & $270 \cdot 10^6$ & $-70\%$\\
U.K. & $275 \cdot 10^6$ & $959 \cdot 10^6$ & $-71\%$ \\
U.S.A. & $x$ & $y$ & $z\%$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[H]
    \includegraphics[width=0.5\textwidth]{no_and_uk_2013.pdf}
    \caption{Monte-Carlo and Hubbert forecast based on past production data for Norway (top) and the U.K. (bottom).  In both cases, our model forecasts a significantly slower decay than the Hubbert model.}
    \label{no_uk_2013}
\end{figure}

The difference between our and the Hubbert-based forecast is striking. In the case, of Norway, the standard Hubbert model was used to fit the production data. According to it, Norway's future oil production would decay much faster than in the Monte-Carlo case. As such, the estimated remaining reserves are less than one third of the forecasted value of the Monte-Carlo methodology. In the U.K., oil production faced a change of regime during the early nineties due to technological innovation, giving rise to the inverted "w shape" of the oil production profile. The standard procedure to model oil production in those cases is to use a multi-cyclic Hubbert curve. The multi-cyclic Hubbert model is a generalization of the standard one. Conceptually, it is just a superposition of several standard (single-cyclic) Hubbert curves. Two cycles are commonly used to fit the UK's oil production. The difference between the Hubbert-based methodology and the Monte-Carlo one is very similar to the Norwegian case. The former underestimates the remaining oil reserves by about $70\%$ compared to the latter.\\
Which of the two models is more thrust worthy? Clearly, the implications in adopting one methodology over the other are significant. The only way to answer this question is to backtest them. In other words we ask: "What would have each of the models predicted had we used them 10, 15, 20 years in the past?"

%\begin{table}[H]
%\caption{Example table}
%\centering
%\begin{tabular}{llr}
%\toprule
%\multicolumn{2}{c}{Name} \\
%\cmidrule(r){1-2}
%First name & Last Name & Grade \\
%\midrule
%John & Doe & $7.5$ \\
%Richard & Miles & $2$ \\
%\bottomrule
%\end{tabular}
%\end{table}

%------------------------------------------------

\section{Validation}

For each of the three countries, namely Norway, the U.K. and the U.S.A., we went back in time 10, 15 and 20 years respectively. The reason for chosing an earlier starting point for our backtest in the case of the U.S.A. compared to say, Norway has to do with the number of oil fields available at the time of the backtest. The U.S.A. having the biggest number of oil fields, we could chose an earlier starting point without compromising the quality of our results. Figure \ref{bt} shows the outcome of those tests. In the case of Norway, the production data was taken into account until 2003. Comparing the forecast of both models with the oil production of the subsequent 10 years, we find the predictive power of the Hubbert model is impressive. It perfectlty captures the Norwegian oil production's dynamic between 2003 and 2013. The Monte-Carlo method does not perform as well. After underestimating the Norwegian oil production between 2003 and 2009, it overestimates it between 2009 and 2013. For the U.K. the backtesting was performed on data until 1998. The conclusion is not as obvious as for the previous case. None of the two methods captures perfectly the oil production between 1998 and 2013. The following table summarizes the difference between the two models.

\begin{table}[H]
\caption{Remaining oil reserves}
\centering
\begin{tabular}{llllr}
%\toprule
%\multicolumn{2}{c}{Name} \\
%\cmidrule(r){1-2}
Start & Country & Hubbert & our model & $\Delta$ \\
\midrule
2003 & Norway & $66 \cdot 10^6$ & $251 \cdot 10^6$ & $-74\%$\\
1998 & U.K. & $91 \cdot 10^6$ & $2.1 \cdot 10^9$ & $-43\%$ \\
1993 & U.S.A. & $x'$ & $y'$ & $z'\%$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[H]
    \includegraphics[width=0.5\textwidth]{no_and_uk_bt.pdf}
    \caption{Monte-Carlo and Hubbert forecast based on past production data up to 2003 for Norway (top) and up to 1998 for the U.K. (bottom).  The results can be compared with the subsequent oil production (crosses). In both cases, the Hubbert forecast is more accurate.}
    \label{bt}
\end{figure}

%------------------------------------------------

\section{Conclusion}
We have presented a Monte-Carlo based methodology to forecast future oil production. By extending the oil production of current fields into the future and modeling the discovery rate of new fields, we were able to give forecasts on the future oil production of Norway, the U.K. and the U.S. These forecasts offer significantly scenarios from the ones obtained with standard Hubbert-based methodologies. Indeed, our model gives a 3 times superior estimation of the remaining oil reserves than the standard one. However, our backtesting results seriously questions the credibility of that forecast, since the Hubbert-model clearly outperforms the Monte-Carlo simulation. 
%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

%\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

%\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
%Figueredo, A.~J. and Wolf, P. S.~A. (2009).
%\newblock Assortative pairing and life history strategy - a cross-cultural
 %study.
%\newblock {\em Human Nature}, 20:317--330.

%\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{multicols}

\end{document}

### 2013-10-18 tomorrow_django/oil_and_gas/analysis.py

@@ -1,0 +1,741 @@
+from constants import MIN_PROD
+from constants import M3_TO_BARRELS
+from constants import START
+from constants import BOUNDS
+from constants import MAX_DATE
+from fitting import fit_exponential
+from fitting import fit_stretched_exponential
+from fitting import fit_power_law
+from fitting import prepare_xy
+from fitting import chop_xy
+from fitting import cost_function
+from fitting import xmin2xmin
+from get_data import get_production
+from get_data import get_classification
+from config import DPATH
+
+from fit_logistic import fit_logistic
+from fit_logistic import compute_logistic
+
+from pandas import date_range
+from pandas.tseries.offsets import *
+from pandas.tseries.offsets import DateOffset
+from pandas import concat
+from pandas import Series
+from pandas import DataFrame
+from pandas import Timestamp
+
+from numpy import array
+from numpy import exp
+from numpy import log
+from numpy import sign
+from numpy import arange
+from numpy import log10
+from numpy import ceil
+from numpy import inf
+from numpy.random import choice
+from numpy.random import rand 
+
+from pickle import load
+from pickle import dump
+
+import matplotlib.pyplot as plt
+
+import os
+
+def get_fit_parms(country, until='2013-02-01', 
+    fit_style='stretched exponential', xmin='my_xmin'):
+    """Return a DataFrame with the fields as the index and name of the relevant 
+    fit parameters as columns.
+
+    Args:
+        country -> str.
+            The string representing the name of the country.
+        until -> str.
+            The string representing the date until which we take the monthly
+            oil production into account. This is especially useful for 
+            backtesting.
+        fit_style -> str.
+            The string representing the fit style.
+        xmin -> str.
+            The string representing the x-coordinate of the production 
+            timeseries from which the fitting is done.
+
+    Return:
+        fit_parms -> DataFrame.
+            The DataFrame containing the fit parameters. 
+    """
+    
+    fit_parms = {}
+    if fit_style == 'exponential':
+        fit = fit_exponential
+        columns = ['tau', 'y0']
+    elif fit_style == 'stretched exponential':
+        fit = fit_stretched_exponential
+        columns = ['tau', 'beta', 'y0']
+    elif fit_style == 'power law':
+        fit = fit_power_law
+        columns = ['alpha', 'y0']
+
+    fpath = os.path.join(DPATH, '%s_fit_parms_%s_%s' % (country, until[:4], 
+                fit_style[:2]))
+
[…]

### 2013-10-18 tomorrow_django/oil_and_gas/config.py

@@ -1,0 +1,6 @@
+import os
+
+#The directory with the Norwegian oil data. Frequently used information will
+#also be stored there.
+DPATH = os.path.join(os.environ['HOME'], 'work', 'eth', 'peak_oil', 'data')
+

### 2013-10-18 tomorrow_django/oil_and_gas/constants.py

@@ -1,0 +1,15 @@
+from pandas import Timestamp
+
+MIN_PROD = 1.58
+
+M3_TO_BARRELS = 1. / 6.2898
+
+START = {'NO': Timestamp('1971-06-01'), 'UK': Timestamp('1975-06-01')}
+
+BOUNDS = {'NO': {'2013-02-01': [1e6, 1e7],
+               '2003-02-01': [1e7],},
+        'UK': {'2013-02-01': [1e6, 1e7],
+               '1998-02-01': [1e6, 1e7],
+               '1993-02-01': [1e6, 1e7],},
+       }
+MAX_DATE = Timestamp('2260-01-01')

### 2013-10-18 tomorrow_django/oil_and_gas/fit_logistic.py

@@ -1,0 +1,244 @@
+from numpy import array, exp, arange, isnan, where
+from datetime import datetime
+from scipy.stats import linregress, t
+from scipy.optimize import fmin
+import matplotlib.pyplot as plt
+from matplotlib.dates import DateFormatter
+from pandas import Timestamp
+
+#This code is based on Peter Cauwel's matlab code.
+#Recommended article to read (explaining part of the methodology:
+#Quis pendit ipsa pretia: facebook valuation and diagnostic of a bubble based
+#on nonlinear demographic dynamics. Peter Cauwels, Didier Sornette (2011).
+#link: http://arxiv.org/pdf/1110.1319v2.pdf
+from erpy.ipshell import ipshell
+
+def fit_logistic(dates, values, start_date, confid=None, show_plot=False, 
+        xlabel='x', ylabel='y', init_conds=(2., 5.)):
+    """Fits logistic function to data and returns r, k and p. 
+    Required arguments are the dates and values as well as the starting date.
+    confid, the confidence interval can be specified as well."""
+
+    dates = array([float(d.toordinal()) for d in dates])
+    values = array([float(v) for v in values])
+    start_date = start_date.toordinal()
+    dates = dates - start_date
+
+    valmax = values.max()
+    idxmax = where(values==values.max())[0][0]
+    values /= valmax
+
+    #It can be shown that there is a linear dependance between the population 
+    #and the growth rate.
+    r, k, p = 1. / dates[idxmax], 1., values[0]
+
+    if p == 0:
+        p = 0.01 
+    #Sometimes the initial estimates of r, k, p are too far from the real values
+    #for the optimzer to converge. We therefore, chose other starting points.
+    #p0 is usually the problematic parameter.
+    mult, len_grid = init_conds
+    grid = mult ** arange(-len_grid, len_grid+1)
+    ps = p * grid
+    rs = r * grid
+    ks = k * grid    
+
+    rkp_f = []
+    for k in ks:
+        for r in rs:
+            for p in ps:
+                #Here we optimize (r, k, p) without constraint.
+                if confid == None:
[…]

### 2013-10-18 tomorrow_django/oil_and_gas/fitting.py

@@ -1,0 +1,170 @@
+from datetime import datetime
+
+from numpy import array
+from numpy import ndarray
+from numpy import exp
+from numpy import log
+from numpy import inf
+from numpy import sign
+
+from scipy.optimize import fmin
+from scipy.stats import linregress
+
+import matplotlib.pyplot as plt
+
+from pandas import DataFrame
+from pandas import Series
+from pandas import Timestamp
+
+def fit_exponential(_x, _y=None, xmin=None, xmax=None, show=False, 
+    xscale='linear', yscale='linear', ax='None', title='None'):
+    """Fits an exponential y0 exp(-x / tau) to the data for each x > xmin.
+    xmin defines the point from which the exponential regime holds."""
+
+    x_, y_ = prepare_xy(_x, _y)
+    xmin = xmin2xmin(xmin, x_, y_, _x, _y)
+    x, y = chop_xy(x_, y_, xmin, xmax)
+
+    logy = log(y)
+    lamda, logy0, _, _, _ = linregress(x, logy)
+    tau, y0 = 1. / lamda, exp(logy0)
+
+    if show:
+        show_fit('exponential', x_=x_, y_=y_, x=x, y=y, y0=y0, tau=tau,
+            xscale=xscale, yscale=yscale, ax=ax, title=title)
+
+    return (tau, y0)
+
+def fit_power_law(_x, _y=None, xmin=None, xmax=None, show=False, 
+    xscale='linear', yscale='linear', ax='None', title='None'):
+    """Fits a power law A x ^ -alpha for each x > xmin. xmin defines the point
+    from which the power law regime holds."""
+
+    x_, y_ = prepare_xy(_x, _y)
+    xmin = xmin2xmin(xmin, x_, y_, _x, _y)
+    x, y = chop_xy(x_, y_, xmin, xmax)
+
+    logx, logy = log(x), log(y)
+    alpha, logy0, _, _, _ = linregress(logx[1:], logy[1:])
+    y0 = exp(logy0)
+
+    if show:
+        show_fit('power-law', x_=x_, y_=y_, x=x, y=y, y0=y0, alpha=alpha,
+            xscale=xscale, yscale=yscale, ax=ax, title=title)
+
+    return (alpha, y0)
+
+def fit_stretched_exponential(_x, _y=None, xmin=None, xmax=None, 
+    show=False, xscale='linear', yscale='linear', ax='None', title='None'):
[…]

### 2013-10-18 tomorrow_django/oil_and_gas/get_data.py

@@ -1,0 +1,164 @@
+from pandas import DataFrame
+from pandas import Timestamp
+from pandas import MultiIndex
+from pandas import concat
+from pandas import read_csv
+from pandas import read_html
+
+from pickle import dump
+from pickle import load
+
+from numpy import nan
+
+from itertools import product
+
+import os
+
+import re
+
+from constants import M3_TO_BARRELS
+
+from config import DPATH
+
+def get_production(country, until='2013-02-01'):
+    """Return a DataFrame where the index are the dates and columns are the
+    monthly production of each of the country's oil fields. File downloadable on
+    http://factpages.npd.no/factpages/ under Field -> Production.
+
+    Args:
+        country -> str.
+            The string representing the name of the coutry. Ex: 'NO' for Norway,
+            'UK' for the United Kingdom.
+        until -> str.
+            The string representing the date until which the production should
+            be returned.
+    Return:
+        production -> DataFrame.
+            The DataFrame containing the timeseries of monthly oil production.
+    """ 
+
+    file_path = os.path.join(DPATH, '%s_monthly_oil_production.pkl' % country)
+
+    if os.path.exists(file_path):
+        f = open(file_path)
+        df = load(f)
+        return df[:until].dropna(how='all', axis=1).dropna(how='all', axis=0)
+    else:
+        print 'Getting the data...'
+        f = open(file_path, 'w')
+
+    os.chdir(DPATH)
+
+    if country == 'NO':
+        #Reading in the csv file. We get a MultiIndexed DataFrame with the 
+        #hierarchies being field_name, year, month.
+        df = read_csv('oil_production_by_fields.csv', index_col=[0, 1, 2])
+        #We only need the oil production
+        df = df['prfPrdOilNetMillSm3'].dropna()
+        mind = df.index
+        _mind = [(field, Timestamp('%d-%d-01' % (int(year), int(month)))) 
+                    for (field, year, month) in mind]
+        #Dealing with Danish special characters...
+        _mind = [(m[0].decode('utf-8'), m[1]) for m in _mind]
+        _mind = [(re.sub(r'\xc5', 'A', m[0]), m[1]) for m in _mind]
+        _mind = [(re.sub(r'\xd8', 'O', m[0]), m[1]) for m in _mind]
+        #I cannot call MultiIndex(_mind) or MultiIndex(zip(*_mind)). I don't
+        #understand why. So I go through the following construction:
[…]

### 2013-10-18 tomorrow_django/oil_and_gas/plotting.py

@@ -1,0 +1,474 @@
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+from get_data import get_production, get_classification
+from analysis import get_fit_parms, classify_fields_according_to_urr
+from fitting import prepare_xy
+
+from numpy import exp, sign, array, arange
+
+from pandas import Series, date_range
+
+from pickle import load, dump
+
+import os
+
+from constants import START
+
+mpl.rcParams['font.size'] = 15.0
+mpl.rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+mpl.rc('text', usetex=True)
+
+def irregular_vs_regular(w=8, l=10, save=False):
+    uk_regular = 'WELTON'
+    no_irregular = 'VALE'
+
+    uk_prod = get_production('UK')[uk_regular].dropna()
+    no_prod = get_production('NO')[no_irregular].dropna()
+
+    fig = plt.figure(figsize=(w, l))
+    ax = fig.add_subplot(211)
+    uk_prod.plot(ax=ax, marker='x', ls='', color='b')
+    l1 = ax.get_lines()[0]
+
+    ax.set_ylabel('Production [barrels/day]')
+    ax.set_title('Example of a regular field')
+    ax.legend([l1], [uk_regular.lower().capitalize() + ' (UK)'])
+   
+
+    ax2 = fig.add_subplot(212)
+    no_prod.plot(ax=ax2, marker='o', ls='', color='b')
+    l2 = ax2.get_lines()[0]    
+
+    ax2.set_xlabel('Time')
+    ax2.set_ylabel('Production [barrels/day]')
+    ax2.set_title('Example of an irregular field')
+    ax2.legend([l2], [no_irregular.lower().capitalize() + ' (NO)'], 
+               loc='upper left')
+
+    fig.tight_layout()
+
+    if save:
+        os.chdir('article')
+        fig.savefig('regular_and_irregular.pdf')
+        os.chdir('..')
+    else:
+        plt.show()
+
+
+def stretched_exponential(w=8, l=10, save=False):
+    field = 'WELTON'
+    prod = get_production('UK')[field].dropna()
+    fit_parms = get_fit_parms('UK', '2013-02-01').ix[field]    
+
[…]

### 2013-10-18 tomorrow_django/oil_and_gas/analysis.py
Started implementing API.

@@ -7,9 +7,6 @@ from fitting import fit_exponential
 from fitting import fit_stretched_exponential
 from fitting import fit_power_law
 from fitting import prepare_xy
-from fitting import chop_xy
-from fitting import cost_function
-from fitting import xmin2xmin
 from get_data import get_production
 from get_data import get_classification
 from config import DPATH
@@ -19,7 +16,6 @@ from fit_logistic import compute_logistic
 
 from pandas import date_range
 from pandas.tseries.offsets import *
-from pandas.tseries.offsets import DateOffset
 from pandas import concat
 from pandas import Series
 from pandas import DataFrame

### 2013-10-18 tomorrow_django/oil_and_gas/models.py

@@ -8,9 +8,15 @@ class Production(models.Model):
     logger.info("Created production.")
     name = models.CharField(max_length=50, default="", unique_for_date="date")
     date = models.DateField()
+    # In BBL (barrels)
     production_oil = models.PositiveIntegerField(default=0)
+    # In MCF (million cubic feet)
     production_gas = models.PositiveIntegerField(null=True, default=None)
     production_water = models.PositiveIntegerField(null=True, default=None)
+
+    @property
+    def production(self):
+        return self.production_oil + (self.production_gas * 1000 / 5800)
 
     def __str__(self):
         return self.name + " ; " + self.date.__str__() + " ; "

### 2013-10-18 tomorrow_django/oil_and_gas/serializers.py

@@ -1,0 +1,23 @@
+__author__ = 'lucas.fievet'
+
+from rest_framework.relations import HyperlinkedRelatedField, HyperlinkedIdentityField
+from rest_framework import serializers
+
+from .models import (FieldProduction)
+
+import logging
+logger = logging.getLogger("UsLoader")
+
+
+class FieldProductionSerializer(serializers.ModelSerializer):
+    production = serializers.IntegerField(source='production', read_only=True)
+
+    class Meta:
+        model = FieldProduction
+        fields = ('date', 'production')
+
+
+class FieldSerializer(serializers.ModelSerializer):
+    class Meta:
+        model = FieldProduction
+        fields = ('name',)

### 2013-10-18 tomorrow_django/oil_and_gas/urls.py

@@ -1,0 +1,15 @@
+__author__ = 'lucas.fievet'
+
+from django.conf.urls import patterns, url
+
+from . import views
+
+oil_and_gas_urlpatterns = patterns(
+    '',
+    url(r'^fields/',
+        views.FieldList.as_view(),
+        name='field-list'),
+    url(r'^productions/(?P<name>[\w-]*)/$$',
+        views.FieldProductionList.as_view(),
+        name='production-list'),
+)

### 2013-10-18 tomorrow_django/oil_and_gas/views.py

@@ -1,1 +1,85 @@
-# Create your views here.
+from django.conf import settings
+from django.http import HttpResponse
+from django.utils.translation import ugettext_lazy as _
+from django.contrib.staticfiles.views import serve
+
+from rest_framework.generics import ListAPIView
+from rest_framework import permissions, generics, status, views, decorators
+from rest_framework.reverse import reverse
+from rest_framework.response import Response
+
+from .models import (FieldProduction)
+from .serializers import (FieldProductionSerializer, FieldSerializer)
+
+import logging
+logger = logging.getLogger("rh")
+
+
+def hello_world(request):
+    return HttpResponse(_("hello world"))
+
+
+class AuthenticatedOrReadOnlyView(object):
+    permission_classes = (permissions.IsAuthenticatedOrReadOnly,)
+
+
+class AuthenticatedView(object):
+    permission_classes = (permissions.IsAuthenticated,)
+
+
+class LoggedViewMixin(object):
+    def dispatch(self, request, *args, **kwargs):
+        response = views.APIView.dispatch(self, request, *args, **kwargs)
+        if response.status_code > 399:
+            log_method = logger.warning
+        elif request.method == 'GET':
+            log_method = logger.debug
+        else:
+            log_method = logger.info
+
+        if "slug" in kwargs:
+            append = " using slug key " + kwargs["slug"]
+        elif "pk" in kwargs:
+            append = " using primary key " + kwargs["pk"]
+        else:
+            append = " with no recognised key"
+        log_method("User " + request.user.username +
+                   " accessed API " + self.view_name + " with method " + request.method + append)
+        return response
+
+
+@decorators.api_view(('GET',))
+@decorators.permission_classes((permissions.IsAuthenticated,))
+def api_root(request, format=None):
+    logger.info("User " + request.user.username + " accessed API root.")
+    return Response({
+        'field-list': reverse(
+            'field-list',
+            request=request,
+            format=format,
+        ),
+        'production-list': reverse(
+            'production-list',
+            request=request,
+            format=format,
+            args=("00335",),
+        ),
+    })
+
+
+class FieldList(AuthenticatedView, LoggedViewMixin, ListAPIView):
+    model = FieldProduction
+    serializer_class = FieldSerializer
+    view_name = "field list"
+
+    def get_queryset(self):
+        return FieldProduction.objects.all().distinct("name")
+
+
[…]

### 2013-10-18 tomorrow_django/tomorrow_django/settings.py

@@ -124,6 +124,7 @@ INSTALLED_APPS = (
     'django.contrib.admin',
     # Uncomment the next line to enable admin documentation:
     # 'django.contrib.admindocs',
+    'rest_framework',
     'oil_and_gas',
     'us_loader',
     'uk_loader',
@@ -169,7 +170,7 @@ LOGGING = {
         },
     },
     'loggers': {
-        'BseeLoader': {
+        'UsLoader': {
             'handlers': ['default', 'console'],
             'level': 'INFO',
             'propagate': True
@@ -201,5 +202,17 @@ CELERYBEAT_SCHEDULE = {
     #     'task': 'uk_loader.tasks.updateUk',
     #     'schedule': timedelta(seconds=30),
     # },
+}
+
+REST_FRAMEWORK = {
+    # Use hyperlinked styles by default.
+    # Only used if the `serializer_class` attribute is not set on a view.
+    'DEFAULT_MODEL_SERIALIZER_CLASS': 'rest_framework.serializers.HyperlinkedModelSerializer',
+
+    # Use Django's standard `django.contrib.auth` permissions,
+    # or allow read-only access for unauthenticated users.
+    'DEFAULT_PERMISSION_CLASSES': [
+        'rest_framework.permissions.DjangoModelPermissionsOrAnonReadOnly'
+    ]
 }
 

### 2013-10-18 tomorrow_django/tomorrow_django/urls.py

@@ -4,6 +4,9 @@ from django.conf.urls import patterns, include, url
 from django.contrib import admin
 admin.autodiscover()
 
+from oil_and_gas import views as oil_and_gas_views
+from oil_and_gas.urls import oil_and_gas_urlpatterns
+
 urlpatterns = patterns('',
     # Examples:
     # url(r'^$', 'tomorrow_django.views.home', name='home'),
@@ -11,6 +14,10 @@ urlpatterns = patterns('',
 
     # Uncomment the admin/doc line below to enable admin documentation:
     url(r'^admin/doc/', include('django.contrib.admindocs.urls')),
+
+    url(r'^api/auth/', include('rest_framework.urls', namespace='rest_framework')),
+    url(r'^api/$', oil_and_gas_views.api_root),
+    url(r'^api/', include(oil_and_gas_urlpatterns)),
 
     # Uncomment the next line to enable the admin:
     url(r'^admin/', include(admin.site.urls)),

### 2013-10-18 tomorrow_django/oil_and_gas/views.py
Fixed field list view.

@@ -73,7 +73,7 @@ class FieldList(AuthenticatedView, LoggedViewMixin, ListAPIView):
     view_name = "field list"
 
     def get_queryset(self):
-        return FieldProduction.objects.all().distinct("name")
+        return FieldProduction.objects.all().values('name').distinct()
 
 
 class FieldProductionList(AuthenticatedView, LoggedViewMixin, ListAPIView):

### 2013-10-19 tomorrow_django/oil_and_gas/static/js/production_plot.js
Simple oil production plot page.

@@ -1,0 +1,90 @@
+/**
+ * Created by Lucas-Fievet on 10/19/13.
+ */
+
+$(function () {
+
+    $.getJSON("/api/fields/").done(function (data) {
+        for (var i = 0; i < data.length; i++) {
+            $("<option value='" + data[i].name + "'>" + data[i].name + "</option>").appendTo($("#fields"));
+        }
+        $("#fields").change(function () {
+            load_field_production($("#fields").val());
+        });
+    });
+    function load_field_production(field_name) {
+        $.getJSON("/api/productions/" + field_name + "/").done(function (data) {
+            var productions = [];
+            var startYear = parseInt(data[0].date.substring(0, 4));
+            var startMonth = parseInt(data[0].date.substring(5, 7));
+            var startDay = parseInt(data[0].date.substring(8, 10));
+            for (var i = 0; i < data.length; i++) {
+                productions.push(data[i].production);
+            }
+
+            $('#container').highcharts({
+                chart: {
+                    zoomType: 'x',
+                    spacingRight: 20
+                },
+                title: {
+                    text: 'Production of the field ' + field_name
+                },
+                subtitle: {
+                    text: document.ontouchstart === undefined ?
+                        'Click and drag in the plot area to zoom in' :
+                        'Pinch the chart to zoom in'
+                },
+                xAxis: {
+                    type: 'datetime',
+                    maxZoom: 14 * 24 * 3600000, // fourteen days
+                    title: {
+                        text: null
+                    }
+                },
+                yAxis: {
+                    title: {
+                        text: 'Production in BBL'
+                    }
+                },
+                tooltip: {
+                    shared: true
+                },
+                legend: {
+                    enabled: false
+                },
+                plotOptions: {
+                    area: {
+                        fillColor: {
+                            linearGradient: { x1: 0, y1: 0, x2: 0, y2: 1},
+                            stops: [
+                                [0, Highcharts.getOptions().colors[0]],
+                                [1, Highcharts.Color(Highcharts.getOptions().colors[0]).setOpacity(0).get('rgba')]
+                            ]
+                        },
+                        lineWidth: 1,
+                        marker: {
+                            enabled: false
+                        },
+                        shadow: false,
+                        states: {
+                            hover: {
+                                lineWidth: 1
+                            }
+                        },
+                        threshold: null
+                    }
+                },
+                series: [
+                    {
+                        type: 'area',
+                        name: 'Production',
[…]

### 2013-10-19 tomorrow_django/oil_and_gas/views.py

@@ -83,3 +83,10 @@ class FieldProductionList(AuthenticatedView, LoggedViewMixin, ListAPIView):
 
     def get_queryset(self):
         return FieldProduction.objects.filter(name=self.kwargs['name'])
+
+
+def oil_and_gas_base(request):
+    if settings.DEBUG:
+        return serve(request, '/views/production.html')
+    else:
+        return serve(request, '/views/production.html')

### 2013-10-19 tomorrow_django/tomorrow_django/urls.py

@@ -21,4 +21,6 @@ urlpatterns = patterns('',
 
     # Uncomment the next line to enable the admin:
     url(r'^admin/', include(admin.site.urls)),
+
+    url(r'^$', oil_and_gas_views.oil_and_gas_base, name="tomorrow-base"),
 )

### 2013-10-20 tomorrow_django/oil_and_gas/migrations/0002_auto__add_field__del_field_countryproduction_depth__add_unique_country.py
Ensure that not duplicates can be saved to database and add retry for http calls.

@@ -1,0 +1,104 @@
+# -*- coding: utf-8 -*-
+import datetime
+from south.db import db
+from south.v2 import SchemaMigration
+from django.db import models
+
+
+class Migration(SchemaMigration):
+
+    def forwards(self, orm):
+        # Adding model 'Field'
+        db.create_table(u'oil_and_gas_field', (
+            (u'id', self.gf('django.db.models.fields.AutoField')(primary_key=True)),
+            ('name', self.gf('django.db.models.fields.CharField')(default='', unique=True, max_length=50)),
+            ('country', self.gf('django.db.models.fields.CharField')(default='', unique=True, max_length=50)),
+            ('production_oil', self.gf('django.db.models.fields.TextField')(default='')),
+            ('production_gas', self.gf('django.db.models.fields.TextField')(default='')),
+            ('production', self.gf('django.db.models.fields.TextField')(default='')),
+            ('production_oil_smooth', self.gf('django.db.models.fields.TextField')(default='')),
+            ('production_gas_smooth', self.gf('django.db.models.fields.TextField')(default='')),
+            ('production_smooth', self.gf('django.db.models.fields.TextField')(default='')),
+        ))
+        db.send_create_signal(u'oil_and_gas', ['Field'])
+
+        # Deleting field 'CountryProduction.depth'
+        db.delete_column(u'oil_and_gas_countryproduction', 'depth')
+
+        # Adding unique constraint on 'CountryProduction', fields ['date', 'name']
+        db.create_unique(u'oil_and_gas_countryproduction', ['date', 'name'])
+
+        # Adding unique constraint on 'FieldProduction', fields ['date', 'name']
+        db.create_unique(u'oil_and_gas_fieldproduction', ['date', 'name'])
+
+        # Adding unique constraint on 'WellProduction', fields ['date', 'name']
+        db.create_unique(u'oil_and_gas_wellproduction', ['date', 'name'])
+
+
+    def backwards(self, orm):
+        # Removing unique constraint on 'WellProduction', fields ['date', 'name']
+        db.delete_unique(u'oil_and_gas_wellproduction', ['date', 'name'])
+
+        # Removing unique constraint on 'FieldProduction', fields ['date', 'name']
+        db.delete_unique(u'oil_and_gas_fieldproduction', ['date', 'name'])
+
+        # Removing unique constraint on 'CountryProduction', fields ['date', 'name']
+        db.delete_unique(u'oil_and_gas_countryproduction', ['date', 'name'])
+
[…]

### 2013-10-20 tomorrow_django/oil_and_gas/models.py

@@ -6,7 +6,7 @@ logger = logging.getLogger("OilAndGas")
 
 class Production(models.Model):
     logger.info("Created production.")
-    name = models.CharField(max_length=50, default="", unique_for_date="date")
+    name = models.CharField(max_length=50, default="")
     date = models.DateField()
     # In BBL (barrels)
     production_oil = models.PositiveIntegerField(default=0)
@@ -23,6 +23,7 @@ class Production(models.Model):
 
     class Meta:
         abstract = True
+        unique_together = (("name", "date"),)
 
 
 class CountryProduction(Production):
@@ -49,3 +50,14 @@ class WellProduction(Production):
 
     def __str__(self):
         return self.name + " ; " + self.country + " ; " + self.date.__str__() + " ; "
+
+
+class Field(models.Model):
+    name = models.CharField(max_length=50, default="", unique=True)
+    country = models.CharField(max_length=50, default="", unique=True)
+    production_oil = models.TextField(default="")
+    production_gas = models.TextField(default="")
+    production = models.TextField(default="")
+    production_oil_smooth = models.TextField(default="")
+    production_gas_smooth = models.TextField(default="")
+    production_smooth = models.TextField(default="")

### 2013-10-20 tomorrow_django/oil_and_gas/static/js/helpers.js

@@ -1,0 +1,13 @@
+/**
+ * Created by Lucas-Fievet on 10/19/13.
+ */
+
+(function ($) {
+
+    $.number_month = function (date_string) {
+        var startYear = parseInt(date_string.substring(0, 4), 10),
+            startMonth = parseInt(date_string.substring(5, 7), 10);
+        return startYear * 12 + startMonth;
+    };
+
+}(window.jQuery));

### 2013-10-20 tomorrow_django/oil_and_gas/static/js/production_plot.js

@@ -12,14 +12,13 @@ $(function () {
             load_field_production($("#fields").val());
         });
     });
+
     function load_field_production(field_name) {
         $.getJSON("/api/productions/" + field_name + "/").done(function (data) {
             var productions = [];
-            var startYear = parseInt(data[0].date.substring(0, 4));
-            var startMonth = parseInt(data[0].date.substring(5, 7));
-            var startDay = parseInt(data[0].date.substring(8, 10));
+            var first_month = $.number_month(data[0].date);
             for (var i = 0; i < data.length; i++) {
-                productions.push(data[i].production);
+                productions.push([$.number_month(data[i].date) - first_month, data[i].production]);
             }
 
             $('#container').highcharts({
@@ -79,8 +78,8 @@ $(function () {
                     {
                         type: 'area',
                         name: 'Production',
-                        pointInterval: 24 * 3600 * 1000 * 30,
-                        pointStart: Date.UTC(startYear, startMonth, startDay),
+                        //pointInterval: 24 * 3600 * 1000 * 30,
+                        //pointStart: Date.UTC(startYear, startMonth, startDay),
                         data: productions
                     }
                 ]

### 2013-10-20 tomorrow_django/tomorrow_django/settings.py

@@ -17,7 +17,7 @@ DATABASES = {
         'USER': 'tomorrow',
         'PASSWORD': 't4ever!',
         'HOST': '127.0.0.1',          # Empty for localhost through domain sockets or '127.0.0.1' for localhost through TCP.
-        'PORT': '3307',                   # Set to empty string for default.
+        'PORT': '3306',                   # Set to empty string for default.
     }
 }
 
@@ -152,7 +152,7 @@ LOGGING = {
     },
     'handlers': {
         'default': {
-            'level':'INFO',
+            'level':'ERROR',
             'class':'logging.handlers.RotatingFileHandler',
             'filename': 'logs/mylog.log',
             'maxBytes': 1024*1024*5, # 5 MB
@@ -165,14 +165,14 @@ LOGGING = {
             'class': 'django.utils.log.AdminEmailHandler'
         },
         'console':{
-            'level': 'DEBUG',
+            'level': 'ERROR',
             'class': 'logging.StreamHandler'
         },
     },
     'loggers': {
         'UsLoader': {
             'handlers': ['default', 'console'],
-            'level': 'INFO',
+            'level': 'ERROR',
             'propagate': True
         },
         'django.request': {

### 2013-10-20 tomorrow_django/us_loader/models.py

@@ -1,8 +1,10 @@
 from django.db.models import Max
+from django.db import IntegrityError
 
 from datetime import date
 from BeautifulSoup import BeautifulSoup
 import requests, calendar, datetime
+from requests.adapters import HTTPAdapter
 import logging
 
 from oil_and_gas.models import FieldProduction
@@ -22,7 +24,7 @@ class UsManager():
         return last_date['date__max']
 
     def update(self):
-        logger.info("US update started.")
+        logger.error("US update started.")
         number_updates = 0
         usRequest = UsRequest()
         usRequest.year_month = self.getOldestDate()
@@ -30,9 +32,13 @@ class UsManager():
             productions = usRequest.getProductions()
             number_updates = number_updates + len(productions)
             for production in productions:
-                production.save()
+                try:
+                    production.save()
+                except (IntegrityError,):
+                    logger.error("Entry " + production.__str__() + " already exists")
             usRequest.nextMonth()
-        logger.info("US update finished.")
+            logger.error("Added " + len(productions).__str__() + " new entries.")
+        logger.error("US update finished with " + number_updates.__str__() + " new entries.")
         return number_updates
 
 
@@ -64,11 +70,15 @@ class UsRequest():
 
     def getSoup(self, page):
         r = None
-        if page == 1:
-            self.session = requests.session()
-            r = self.session.post(self.url, data=self.getPostData())
-        elif page > 1:
-            r = self.session.post(self.url, data=self.getNextPagePostData(page))
+        try :
+            if page == 1:
+                self.session = requests.session()
+                self.session.mount('http://', HTTPAdapter(max_retries=5))
+                r = self.session.post(self.url, data=self.getPostData())
+            elif page > 1:
+                r = self.session.post(self.url, data=self.getNextPagePostData(page))
+        except:
+            logger.error("Page was not accessible")
         return BeautifulSoup(r.content)
 
     def getProductions(self):
@@ -88,13 +98,13 @@ class UsRequest():
         table = soup.find('table', border=5, width=600)
 
         if table is None:
-            logger.error("Table is None.", )
+            logger.warning("Table is None.", )
             return productions
 
         trs = table.findAll('tr')
 
         if trs is None:
-            logger.error("Rows are None.")
+            logger.warning("Rows are None.")
             return productions
 
         trs = trs[2:]
@@ -108,7 +118,8 @@ class UsRequest():
     def getProduction(self, row):
         tds = row.findAll('td')
[…]

### 2013-10-20 tomorrow_django/us_loader/tasks.py

@@ -11,7 +11,7 @@ logger = logging.getLogger("UsLoader")
 
 @task()
 def updateUs(date=None):
-    logger.info("Update Bsee task started.")
+    logger.info("Update Us task started.")
     usManager = UsManager()
     if date is not None:
         usManager.update_to = date

### 2013-10-20 tomorrow_django/oil_and_gas/models.py
Improved production ploting.

@@ -61,3 +61,39 @@ class Field(models.Model):
     production_oil_smooth = models.TextField(default="")
     production_gas_smooth = models.TextField(default="")
     production_smooth = models.TextField(default="")
+
+
+class CountryAggregator():
+    def getFields(self):
+        return WellProduction.objects.filter(country="UK").values("field").distinct()
+
+    def aggregateWells(self, name):
+        return WellProduction.objects.filter(field=name).values('date').annotate(
+            total_oil=Sum('production_gas'),
+            total_gas=Sum('production_oil'),
+            total_water=Sum('production_water'),
+        )
+
+    def setFieldData(self, field, agg_well):
+        field.name = agg_well['field']
+        field.country = 'UK'
+        field.date = agg_well['date']
+        field.production_oil = agg_well['total_oil']
+        field.production_gas = agg_well['total_gas']
+        field.production_water = agg_well['total_water']
+
+    def computeFields(self, fields):
+        for field in fields:
+            fieldName = field['field']
+            agg_wells = self.aggregateWells(fieldName)
+            for agg_well in agg_wells:
+                agg_well['field'] = fieldName
+                productionDate = agg_well['date']
+                fieldProduction, created = FieldProduction.objects.get_or_create(name=fieldName, date=productionDate)
+                self.setFieldData(fieldProduction, agg_well)
+                fieldProduction.save()
+        return len(fields)
+
+    def compute(self):
+        fields = self.getFields()
+        return self.computeFields(fields)

### 2013-10-20 tomorrow_django/oil_and_gas/serializers.py

@@ -14,7 +14,7 @@ class FieldProductionSerializer(serializers.ModelSerializer):
 
     class Meta:
         model = FieldProduction
-        fields = ('date', 'production')
+        fields = ('date', 'production_oil')
 
 
 class FieldSerializer(serializers.ModelSerializer):

### 2013-10-20 tomorrow_django/oil_and_gas/static/js/helpers.js

@@ -4,10 +4,11 @@
 
 (function ($) {
 
-    $.number_month = function (date_string) {
-        var startYear = parseInt(date_string.substring(0, 4), 10),
-            startMonth = parseInt(date_string.substring(5, 7), 10);
-        return startYear * 12 + startMonth;
+    $.to_date = function (date_string) {
+        var start_year = parseInt(date_string.substring(0, 4), 10),
+            start_month = parseInt(date_string.substring(5, 7), 10),
+            start_day = parseInt(date_string.substring(8, 10), 10);
+        return Date.UTC(start_year, start_month, start_day);
     };
 
 }(window.jQuery));

### 2013-10-20 tomorrow_django/oil_and_gas/static/js/production_plot.js

@@ -16,9 +16,8 @@ $(function () {
     function load_field_production(field_name) {
         $.getJSON("/api/productions/" + field_name + "/").done(function (data) {
             var productions = [];
-            var first_month = $.number_month(data[0].date);
             for (var i = 0; i < data.length; i++) {
-                productions.push([$.number_month(data[i].date) - first_month, data[i].production]);
+                productions.push([$.to_date(data[i].date), data[i].production_oil]);
             }
 
             $('#container').highcharts({
@@ -78,8 +77,6 @@ $(function () {
                     {
                         type: 'area',
                         name: 'Production',
-                        //pointInterval: 24 * 3600 * 1000 * 30,
-                        //pointStart: Date.UTC(startYear, startMonth, startDay),
                         data: productions
                     }
                 ]

### 2014-06-17 tomorrow_django/oil_and_gas/fitting.py
Merge branch 'master' of https://github.com/LucasFievet/Tomorrow
Conflicts:
	Publications/Forecasting  Future Oil Production/peak_oil.lyx~

@@ -158,7 +158,7 @@ def fit_stretched_exponential(_x, _y=None, x_min=None, x_max=None,
     x_, y_ = prepare_xy(_x, _y)
     x_min = x_min_2_x_min(x_min, x_, y_, _x, _y)
     x, y = chop_xy(x_, y_, x_min, x_max)
-    x -= x_min
+    # x -= x_min
 
     if x_min != x_min_guess or y0 is None or tau is None or beta is None:
         x_no_zeros, y_no_zeros = remove_zeros(x, y)

### 2014-06-17 tomorrow_django/oil_and_gas/static/js/production_plot.js

@@ -34,12 +34,12 @@ $(function () {
                 if (x >= data.x_min) {
                     fit.push([
                         date,
-                        fit_function(x-data.x_min) / 30
+                        fit_function(x) / 30
                     ]);
                     fit_range.push([
                         date,
-                        fit_function(x-data.x_min) * (1 - data.error_std) / 30,// * (1 - data.error_avg),
-                        fit_function(x-data.x_min) * (1 + data.error_std) / 30// * (1 - data.error_avg)
+                        fit_function(x) * (1 - data.error_std) / 30,// * (1 - data.error_avg),
+                        fit_function(x) * (1 + data.error_std) / 30// * (1 - data.error_avg)
                     ]);
                 }
             }
@@ -135,9 +135,7 @@ $(function () {
                     }
                 },
                 subtitle: {
-                    text: document.ontouchstart === undefined ?
-                        'Click and drag in the plot area to zoom in' :
-                        'Pinch the chart to zoom in'
+                    text: ''
                 },
                 xAxis: {
                     type: 'datetime',

### 2014-06-20 Publications/Forecasting  Future Oil Production/peak_oil.tex
Merge branch 'master' of https://github.com/LucasFievet/Tomorrow
Conflicts:
	Publications/Forecasting  Future Oil Production/peak_oil.lyx~

%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside]{article}
\usepackage{mathpazo}
\renewcommand{\ttdefault}{mathpazo}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{float}
\usepackage{amstext}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=false]
 {hyperref}
\usepackage{breakurl}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

\@ifundefined{date}{}{\date{}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.2 (15/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\usepackage{lipsum}% Package to generate dummy text throughout this template

% Use the Palatino font
% Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype}% Slightly tweak font spacing for aesthetics

% Document margins
\usepackage{multicol}% Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}% Custom captions under/above floats in tables or figures
\usepackage{booktabs}% Horizontal rules in tables
\usepackage{float}% Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
% For hyperlinks in the PDF

\usepackage{lettrine}% The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist}% Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract}% Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec}% Allows customization of titles
\renewcommand{\thesection}{\Roman{section}}
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr}% Headers and footers
 % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyhead[C]{Forecasting future oil production $\bullet$ July 2014} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\makeatother

\begin{document}
\thispagestyle{fancy}

\title{Forecasting future oil production}

\author{L. Fiévet, Z. Forró, M. del Degan, P. Cauwels and D. Sornette\\
D-MTEC, ETH Zürich}

\date{05/25/14}
\maketitle
\begin{abstract}
\noindent This paper presents a new Monte-Carlo methodology to forecast
the crude oil production of Norway and the U.K. based on the current/past
performances of individual oil fields. By extrapolating the future
production of these fields and the frequency of new discoveries, the
oil production of these countries can be forecasted. The results indicate
that standard methodology tends to underestimate remaining oil reserves.
A comparison of the model to this methodology is made by making predictions
with the full data and a back-test starting in 2008. It shows that
the model gives a better description of future oil production.
\end{abstract}

\section{Introduction}

\lettrine[nindent=0em,lines=3] Forecasting future oil production
has been a topic of active interest since the beginning of the past
century because of oil's central role in our economy. Its importance
ranges from energy production, through manufacturing to pharmaceuticals
industry. As petroleum is a non-renewable and finite resource, it
is primordial to be able to forecast future oil production. The fear
of a global oil peak, beyond which production will inevitably decline,
has been growing due to stagnating supplies and high oil prices since
the crisis in 2008/2009 \cite{Murray 2012}. As any industrialized
country, Europe is strongly dependent on oil supply to maintain its
economic power. In the nowadays difficult geopolitical environment,
it is important to know how much of the oil needed in Europe will
come from reliable sources. In the past, a big share has been coming
from Norway and the U.K. which have been reliable exporters. However,
the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
production has been declining rapidly as well \cite{H=0000F6=0000F6k 2008}.\\
 The methodology behind forecasting future oil production has not
evolved much since M. King Hubbert, who in 1956 famously predicted
that the U.S. oil production would peak around 1965-1970 \cite{Hubbert 1956}.
That prediction has proven itself to be correct. His main argument
was based on the finiteness of oil reserves and use of the logistic
differential equation

\begin{equation}
\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)\label{logistic}
\end{equation}

\noindent to model oil production. The logistic differential equation
is characterized by an initial exponential growth, which then decreases
to zero as the total oil extracted reaches saturation (no more oil
is to be found). The parameter $r$ is commonly referred to as the
growth rate, and $K$ as the carrying capacity (total quantity of
oil extracted). If $P(t)$ is the the amount of oil extracted up to
time $t$, then $\frac{dP}{dt}$ is the oil production rate, the quantity
that M. King Hubbert predicted with surprising accuracy to peak. From
a methodological point of view, the Hubbert model has enjoyed a longstanding
popularity in modeling future oil production given its simplicity.
Various extensions have been studied by Brandt \cite{Brandt 2006}
to account for multi-cycled or asymmetric production curves.\\
The existing forecasts of future oil production use some form of the
Hubbert model (Brecha, Laherrère or Lynch \cite{Brecha 2012,Laherr=0000E8re 2002,Lynch 2002})
or economical model applied to aggregate production (Greiner \cite{Greiner 2012}),
but none goes into the details of studying the underlying dynamics.
The main reason for the lack of details is certainly the lack of available
data. In this paper, a new methodology is introduces to forecast future
oil production. Instead of taking the aggregate oil production profile
and fitting it with the Hubbert curve or its variants (such as the
multi-cyclic Hubbert curve), the production profile of each individual
oil field is used. By extending their production into the future and
extrapolating the future rate of discovery of new fields, the future
oil production is forecasted by the means of a Monte Carlo simulation.
To demonstrate the universality of the methodology presented here,
it is applied to 2 major oil producing countries with publicly available
data: Norway\cite{Norwegian Data} and the U.K. \cite{U.K. Data}.

\section{Methodology}

The idea behind the methodology is to model the future aggregate oil
production of a country by studying the production dynamics of its
individual constituents, the oil fields. The main benefit of this
approach, compared to working directly with aggregate production data,
is the possibility to forecast non-trivial oil production profiles
arising from the combination of all the individual field\textquoteright{}s
dynamics. In order to achieve that, one must be able 1) to \textbf{extend
the oil production of each individual field} into the future and 2)
to \textbf{extrapolate the rate of discoveries} of new oil fields.

\subsection{Extending the oil production of individual fields}

The first step to predicting the future oil production of a country
is to extrapolate the future production of existing fields and to
estimate the error on this extrapolation. The data of the fields developed
in the past shows a repeating asymmetric pattern. A good example is
the Oseberg field shown in figure \ref{regular}, with a quick ramp
up once the field is being developed, and then a peak or plateau before
the field starts decaying. The decay can take many different shapes
and is governed by a variety of geological and economical factors
which will be captured implicitly by the fitting procedure.

\subsubsection{Regular, irregular and new fields\label{sub:Regular,-irregular-and}}

To be able to forecast the oil production of each individual field,
regularity had to be found in the production's dynamics. Modeling
the whole production profile from the beginning of extraction seems
elusive due to the variety of the forms it can take. Fortunately,
modeling the decay process is sufficient in order to extrapolate future
oil production. A preliminary classification is necessary to achieve
that goal. Figures \ref{regular}, \ref{irregular} and \ref{new}
show that independently of the country, oil fields can be classified
into 3 main categories:
\begin{itemize}
\item \textbf{Regular fields} - Their decay show some regularity. 
\item \textbf{Irregular fields} - The ones that don't decay in a regular
fashion.
\item \textbf{New fields} - The ones that don't decay yet. As such, there
is no easy way to forecast their future oil production based on past
data.
\end{itemize}
\noindent All the fields have been fitted using an automated algorithm,
but the results have been subsequently checked visually to sort out
the irregular fields which could not be fitted. As of January 2014,
regular fields make up 85\% and 87\% of the number of fields and 94\%
and 71\% of the total produced oil volume in Norway and the U.K. respectively.
As such, being able to model them is crucial. To capture as many different
decay dynamics as possible, the stretched exponential

\begin{equation}
P(t)=P_{0}e^{sign(\tau)\left(\frac{t}{\left|\tau\right|}\right)^{\beta}}\label{stretched_exponential}
\end{equation}
has been fitted to the decay of regular fields. The stretched exponential
has many advantages as it generalized the power law and can therefore
capture a broad variety of distributions as shown by Laherrère and
Sornette \cite{Laherr=0000E8re and Sornette 1998}. As can be seen
on figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
is a good functional form to fit the decay process of regular fields.

For the minority of irregular fields, their decay has been modeled
as follows: $\tau$ has been picked to be the average $\tau$ over
the regular fields. Then $\beta$ has been fixed so that the sum of
the field\textquoteright{}s production over its lifetime be equal
to the official ultimate recovery estimates, when such an estimate
is available.

The minority of new fields, which did not yet enter their decay phase,
cannot be extrapolated and will therefore be treated as a new discovery.
The technical details of how to treat them as a new discovery are
discussed in section \ref{sub:Future-production-from}.

\subsubsection{Back-testing \& Error}

To determine how well the extrapolation based on the stretched exponential
predicts the future production, a complete back-testing has been performed
on each field. A single back-test is made as follows:
\begin{itemize}
\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field is
truncated at a certain date in the past $\tau\in\{0,\,\ldots,\, N\}$.
Where $\tau$ is in months since production start of the field.
\item The extrapolation$f(t)$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{\tau}\}$.
\item The future production predicted by the extrapolation $f(t)$ can be
compared to the actual production from the date $\tau$ in the past
up to now $\tau_{f}=N$. The extrapolated total production can be
computed as 
\begin{equation}
P_{e}(\tau)=\int_{\tau}^{\tau_{f}}f(t)dt
\end{equation}
and the relative error is given by
\begin{equation}
e(\tau)=\frac{P_{e}(\tau)-\sum_{i=\tau}^{\tau_{f}}p_{i}}{P_{e}(\tau)},
\end{equation}
where both are a function of the truncation time $\tau$.
\end{itemize}
Computing this back test for every month in the past since the field
starting decaying, yields a plot showing the evolution of the relative
error over time \ref{oesberg-error}. By construction, the relative
error will tend to zero as the truncation time $\tau$ goes to the
present. Nonetheless, it is a uself indicator for the stability of
the extrapolation. As can been seen for the Oseberg field, the relative
error on future production remained fairly stable during the past
decade.

From the complete back-test, on can compute the average relative error
\begin{equation}
\bar{e}=\frac{1}{N}\sum_{i=0}^{N}e(i)
\end{equation}
 the extrapolation made on the future production. Assuming that the
relative errors are normally distributed around the average relative
error, the standard deviation on the average relative error is given
by 
\begin{equation}
\sigma_{e}=\sqrt{\frac{1}{N}\sum_{i=0}^{N}\left(e(i)-\bar{e}\right)^{2}}.\label{eq:field-risk}
\end{equation}
As the average relative error is often fairly constant, the extrapolation
was corrected by the average relative error. Meaning that if the extrapolation
consistently over-estimated the production by 10\% during the back-test,
the extrapolation was reduced by 10\%. This results in an extrapolated
production $p(t)$, including a 1$\sigma$ confidence interval, as
\begin{equation}
p(t)=(1-\bar{e})\, f(t)\pm\sigma_{e}.\label{eq:extrapolation}
\end{equation}

An example of such an extrapolation including a one standard deviation
range is shown on figure \ref{irregular} for the Oesberg field.

\subsubsection{Aggregate error}

Once the individual fields have been extrapolated in the sense of
formula \ref{eq:extrapolation}, the aim is to compute the extrapolation
for the country. While it is straightforward to sum the extrapolations
of the individual fields to obtain the expected production, some care
has to be taken with respect to the confidence interval on the production
at the country level.

As shown later in section \ref{sec:Results}, the same extrapolation
including a complete monthly back-test of total future production
has been performed at the country level and the resulting relative
error is much smaller than the average error observed on the individual
fields. To account for this observation, the assumption made is that
the relative error between individual fields is uncorrelated. Therefore,
the fields can be considered as a portfolio of assets with a return
given by their extrapolation $p(t)$ (eq. \ref{eq:extrapolation})
and a risk given by $\sigma_{field}$ (eq. \ref{eq:field-risk}).
This means that the risk at the country level from the extrapolated
production is to be computed as 
\begin{equation}
\sigma_{country}^{2}=\frac{1}{\#fields}\sum_{field\in fields}\sigma_{field}^{2}.
\end{equation}

Intuitively, this models well the fact that the uncorrelated errors
among fields will mostly cancel out.

\subsection{Discovery rate of new fields\label{sub:Discovery-rate-of}}

Knowing the future production rate of existing fields is not enough
as new fields will be discovered in the future. The model describing
the discovery rate of new fields should satisfy two fundamental observations.
\begin{enumerate}
\item The rate of new discoveries should tend to zero as time goes to infinity.
This is a consequence of the finiteness of the number of oil fields. 
\item The rate of new discoveries should depend on the size of the oil fields.
As of today, giant oil fields are discovered much less frequently
than dwarfs.
\end{enumerate}

\subsubsection{Discoveries modeled as logistic growth\label{sub:Discoveries-modeled-as}}

A natural choice for such a model is a non-homogenous Poisson process.
The Poisson process is a process that generates independent events
at a rate $\lambda$. It is inhomogeneous if the rate is time-dependent,
$\lambda\rightarrow\lambda(t)$. The standard way to measure $\lambda(t)$
is to find a functional form for $N(t)$, the statistical average
of the cumulative number of events (discoveries) up to time $t$.
Then, $\lambda(t)$ is simply $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
shows $N(t)$ for Norwegian fields classified according to their size.
The logistic curve is a good fit to the data (integral form of equation
\ref{logistic}). This implies that after an initial increase, the
rate of new discoveries reaches a peak followed by a decrease until
no more oil fields are to be found, consistent with our fundamental
observations. This same approach has already been successfully applied
by Forro \cite{Forro 2012} to estimate the number of daily active
users on Zynga.

As the discovery and production dynamics are not independent of the
field size, the fields have been split into two groups: dwarfs and
giants. Unfortunately, the two logistic curves thus obtained are highly
sensitive to the splitting size. This results from the major issue,
when fitting a logistic curve to data, that the carrying capacity
can not be determined if the data does not already exhibit the slow
down in growth from reaching the carrying capacity. However, it is
mentioned in the literature that often dwarf fields have already been
discovered a long time ago, but their production has been postponed
for economical reasons \cite[p. 378]{Lynch 2002}. Therefore, it is
expected that the large oil fields have mostly been found and produced,
and that future discoveries will mostly be made up of dwarf fields.
Consequently, the splitting size has been picked as small as possible
in order to maximize the number of giant fields but large enough to
avoid recent discoveries:
\begin{itemize}
\item \textbf{Dwarfs}: Fields which produced less than $50\cdot10^{6}$
barrels.
\item \textbf{Giants}: Fields which produced more than $50\cdot10^{6}$
barrels.
\end{itemize}
The resulting plot shown in figure \ref{rate_of_discovery} pictures
the dynamics: giant oil fields have mostly been found while the discovery
process for dwarf fields is still ongoing. The logistic growth curve
fit to the giant discoveries is well constrained, however the fit
for the dwarfs is poorly constrained. As can be seen on figure \ref{rate_of_discovery},
the carrying capacity $K$ of the logistic growth model is not well
constrained by the available data. A large spectrum of values for
$K$ can lead to an equally good fit of the data. Therefore, a method
described in section \ref{sub:Likelihood-function-for} has been used
to compute the probability of different carrying capacities.

\subsubsection{Likelihood function for the number of discoveries\label{sub:Likelihood-function-for}}

To overcome the logistic fitting issue for the dwarf fields, a method
already used by Smith \cite{Smith 1980} has been implemented. This
method makes the following two postulates:
\begin{enumerate}
\item ``The discovery of reservoirs in a petroleum play can be modeled
statistically as sampling without replacement from the underlying
population of reservoirs.''
\item ``The discovery of a particular reservoir from among the existing
population is random, with a probability of discovery being dependent
on reservoir size.''
\end{enumerate}
The fields are split into $J$ size bins denoted $S_{1},\,\ldots,\, S_{J}$
occurring with frequency $n_{1},\,\ldots,\, n_{J}$. Each discovery
is considered as a step $i$ at which a field of size $I(i)\in\{S_{1},\,\ldots,\, S_{J}\}$
is found and $m_{ij}$ denotes the number of fields of size $j\in\{1,\,\ldots,\, J\}$
discovered before the $i^{th}$ step. Than the probability that the
discovery at step $i^{th}$ is of size $j$ can be expressed as
\begin{equation}
P\left(I(i)=S_{j}\right)=\frac{\left(n_{j}-m_{ij}\right)\cdot S_{j}}{\sum_{k=1}^{J}\left(n_{k}-m_{ik}\right)\cdot S_{k}}.
\end{equation}
The likelihood $L$ for a complete sequence of $N$ discoveries $\{I(1),\,\ldots,\, I(N)\}$
can than be expressed as
\begin{equation}
L=\prod_{i=1}^{N}\frac{\left(n_{I(i)}-m_{iI(i)}\right)\cdot S_{I(i)}}{\sum_{j=1}^{J}\left(n_{j}-m_{ij}\right)\cdot S_{j}}.\label{eq:likelihood}
\end{equation}
The unknown parameters are the number of fields $n_{1},\,\ldots,\, n_{J}$,
whose likelihood can now be estimated based on the existing discoveries.
Using a brute force approach, the entire space of plausible values
for the variables $n_{1},\,\ldots,\, n_{J}$ has been sampled. The
values $n_{j}$ have been sampled between the number of existing fields
$m_{Nj}$ in the bin $j$ and up to a value $n_{j}^{upper}$, such
that the scenario with the largest likelihood according to equation
\ref{eq:likelihood} has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
the bin $j$. Subsequently, the likelihood of each scenario (value
of the tuple $n_{1},\,\ldots,\, n_{J}$) has been normalized such
that the total likelihood of all generated scenarios equals one.

For the analysis of the north sea oil field discoveries, the number
of size bins has been fixed to $J=2$ splitting between dwarfs (1)
and giants (2) as described in section \ref{sub:Discoveries-modeled-as}.

\begin{table}[h]
\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.}

\centering %
\begin{tabular}{lllll}
 & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{likelihood-results}
\end{table}

The results shown in table \ref{likelihood-results} are coherent
with the intuitive expectation that discovering a new giant field
is unlikely and that future discoveries wil mostly be made up of dwarf
fields. The likelihoods obtained for the carrying capacities of dwarfs
and giants have been used to constrain the logistic curve fitted to
the discoveries. Figure \ref{rate_of_discovery} shows a sample of
fitted logistic curves, each curve being weighted by the likelihood
of its carrying capacity given by equation \ref{eq:likelihood}.

\subsubsection{Future production from discoveries\label{sub:Future-production-from}}

The goal is to compute an expected oil production coming from future
discoveries, which requires to combine the steps described in sections
\ref{sub:Discoveries-modeled-as} and \ref{sub:Likelihood-function-for}.

The method described in section \ref{sub:Likelihood-function-for}
yields probabilities for the total number of fields (including the
not yet discovered fields) in each size bins (called a scenario).
However, this likelihood method does not give the time distribution
of future discoveries. The aim is to use the likelihood function to
generate scenarios with their respective occurrence probability.

For a given scenario, the carrying capacity $K$ (= total number of
fields) is given for each size class. This is useful, because it resolves
the instability in fitting the logistic curve to the number of discovered
fields. The time distribution of the discoveries is then given for
each size class by the fitted logistic curve.

The actual size of a newly discovered field is generated according
to the size distribution of the existing fields in its size class.
The probability distribution function of field sizes in a given size
bin has been fitted by a stretched exponential function.

The production curve is computed based on the average production curve
of all existing fields in the same size category. The production curves
of the existing fields have all been normalized to a total production
of one and then have been averaged. This yields the typical production
profile including a one sigma confidence interval. For a new field,
this typical production curve is than multiplied by the size of the
field.

Superposing the production curves results in the expected production
curve from future oil field for a given scenario.

As the total parameter space is too large to be sampled entirely,
a Monte Carlo technique is applied to compute the expected production
with confidence interval from future discoveries. In a nutshell, the
algorithm works as follows:
\begin{enumerate}
\item Draw a scenario based on its probability.
\item Compute the time distribution of new discoveries by fitting a logistic
curve for each size class.
\item For each discovery generate a size and the resulting production curve
based on the size distribution and production curves of existing fields.
\item Superpose all the production curves.
\item Repeat and average over all drawn scenarios.
\end{enumerate}
The result is the expected production curve of future oil field discoveries.
The distribution of generated scenarios yields the confidence interval.

In order to account for the new fields discussed in section \ref{sub:Regular,-irregular-and},
which are already discovered but did not yet enter the decay phase,
the simulated production resulting from new discoveries has been shifted
to the date in the past where it matches the current production from
new fields.

\section{Results\label{sec:Results}}

Based on the methodology described in section \ref{sub:Discovery-rate-of},
simulating future oil production was straightforward. For each country,
the existing fields were extrapolated and the future discoveries were
simulated. Figure \ref{no_uk_2013} shows the average of 1000 simulations.
For each country, the non-symmetric shape of the production's dynamics
contrary to what Hubbert would have predicted is immediately noticeable.

\begin{table}[H]
\caption{Remaining oil reserves until 2030 in barrels for the \textbf{Fit}
of the country's production, the Monte-Carlo \textbf{Model} and the
relative difference $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$
between the two.}

\centering %
\begin{tabular}{lllr}
 & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $31\%$\tabularnewline
U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $40\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

The results in table \ref{forecast-14} show a striking difference
between the fit and the Monte-Carlo model forecast. According to the
fit (extrapolation of aggregate production), Norway's future oil production
would decay much faster than in the Monte-Carlo case. The remaining
reserves estimated with the Monte-Carlo methodology are $30\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is higher
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the Monte-Carlo one is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$40\%$ compared to the latter.

Which of the two models is more thrust worthy? Clearly, the implications
in adopting one methodology over the other are significant. The only
way to answer this question is to back-test them. In other words:
\textquotedbl{}What would have each of the models predicted had they
been used in the past?\textquotedbl{}

\section{Validation}

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the outcome of those tests. Comparing the forecast of both models
with the oil production of the subsequent 6 years, shows that the
predictive power of the Monte-Carlo model is better than a simple
extrapolation of aggregate production. The following table \ref{backtest-08}
summarizes the difference between the two approaches for the back-testing
period.

\begin{table}[H]
\caption{The fit and the Monte-Carlo model are used on the data set truncated
in 2008 and their forecast for the period 2008-2014 is compared to
the actual production.}

\centering %
\begin{tabular}{llll}
 & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
\midrule Norway & $133\cdot10^{6}$  & $108\cdot10^{6}$  & $130\cdot10^{6}$ \tabularnewline
U.K. & $79\cdot10^{6}$  & $66\cdot10^{6}$  & $75\cdot10^{9}$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}

The difference is not huge for the period from 2008 to 2014, but the
Monte-Carlo model nevertheless is noticeably closer to the actual
value. However, when considering the period after 2014, the difference
is huge as shown in table \ref{forecast-08}.

\begin{table}[H]
\caption{Remaining oil reserves forecasted for the period 2014-2030 when using
the data truncated in 2008.}

\centering %
\begin{tabular}{lllc}
 & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
\midrule Norway 08 & $25\cdot10^{6}$  & $171\cdot10^{6}$  & $-85\%$\tabularnewline
U.K. 08 & $25\cdot10^{6}$  & $91\cdot10^{9}$  & $-73\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-08}
\end{table}

The simple extrapolation decays too fast and entirely misses the fat
tails in the decay process of individual fields and the new discoveries.
As well, it must be noted that the simple extrapolation changed massively
between the back-test in table \ref{forecast-08} and the current
fit in table \ref{forecast-14} (520\% for Norway and 236\% for the
U.K.). In other words, the simple extrapolation was very unstable
in its forecast, while the Monte-Carlo forecasts remained very consistent
(less than 10\% change).

As can be seen in figure \ref{model-back-testing}, the actual production
of Norway during the back-testing period remained entirely within
the quite narrow $1\sigma$ interval of the Monte-Carlo methodology,
while totally breaking out of the $1\sigma$ interval of the simple
extrapolation. For the U.K. the Monte-Carlo methodology only performs
slightly better when considering the confidence interval, and the
confidence interval is much larger due to the uncertainty on future
discoveries and their production profile.

\section{Conclusion}

This paper presented a Monte-Carlo based methodology to forecast future
oil production. By extending the oil production of current fields
into the future and modeling the discovery rate of new fields, the
future oil production of Norway and the U.K. could be forecasted.
These forecasts are significantly different from the ones obtained
with a standard extrapolation. Indeed, our model forecasts 30\% to
40\% more remaining oil reserves than the standard extrapolation.
The back-test performed on the time period between 2008 and 2014 confirmed
that the Monte-Carlo based model better captured the production dynamics.

The results are hopeful, as they make it highly likely that the decay
of Norwegian and U.K. oil production will be much slower then one
would expect from a standard extrapolation. Nonetheless, to maintain
current levels of oil consumption in the European Union, more of it
will have to be imported from outside Europe, as the imports from
Norway will vanish (currently accounting for 11\% of E.U. oil imports
\cite{European oil imports}) and the U.K. will need to import more
oil.

\begin{table}[H]
\caption{Oil import (bbl/day) at a constant consumption of 1.5M bbl/day for
the U.K. and 0.22M bbl/day for Norway. The import for the E.U. and
Norway is a lower bound based on the changes in the U.K and Norway.}

\centering %
\begin{tabular}{lllcc}
 & 2014 & 2020 & 2025 & 2030\tabularnewline
\midrule Norway & $-1.23\cdot10^{6}$  & $-0.88\cdot10^{6}$  & $-0.58\cdot10^{6}$  & $-0.43\cdot10^{6}$ \tabularnewline
U.K. & $0.7\cdot10^{6}$  & $0.9\cdot10^{6}$  & $1.0\cdot10^{6}$  & $1.1\cdot10^{6}$ \tabularnewline
E.U.+Norway (lower bound) & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{needs}
\end{table}

As shown in table \ref{needs}, at constant consumption, the Monte-Carlo
model predicts that in 2030 the E.U. with Norway will need to increase
its oil imports by $1.3$ million barrels. Imports which will most
likely have to come from outside Europe.
\begin{thebibliography}{10}
\bibitem{Brandt 2006}Brandt, Adam R. \textquotedblleft{}Testing Hubbert.\textquotedblright{}
Energy Policy 35, no. 5 (May 2007): 3074\textendash{}88. doi:10.1016/j.enpol.2006.11.004.

\bibitem{Brecha 2012}Brecha, Robert J. \textquotedblleft{}Logistic
Curves, Extraction Costs and Effective Peak Oil.\textquotedblright{}
Energy Policy 51 (December 2012): 586\textendash{}97. doi:10.1016/j.enpol.2012.09.016.

\bibitem{European oil imports}European Commission - Energy (2014,
Mai 25). EU Crude oil imports. Retrieved from \url{http://ec.europa.eu/energy/observatory/oil/import_export_en.htm}.

\bibitem{Forro 2012}Forró, Zalán, Peter Cauwels, and Didier Sornette.
\textquotedblleft{}When Games Meet Reality: Is Zynga Overvalued?\textquotedblright{}
Journal of Investment Strategies 1 (3), 119-145 (2012).

\bibitem{U.K. Data}GOV.UK (2014, Mai 25). Exploration and production
- Oil and gas: field data. Retrieved from \url{https://www.og.decc.gov.uk/pprs/full_production.htm}.

\bibitem{Greiner 2012}Greiner, Alfred, Willi Semmler, and Tobias
Mette. \textquotedblleft{}An Economic Model of Oil Exploration and
Extraction.\textquotedblright{} Computational Economics 40, no. 4
(December 2012): 387\textendash{}99. doi:10.1007/s10614-011-9272-0.

\bibitem{H=0000F6=0000F6k 2008}Höök, Mikael, and Kjell Aleklett.
\textquotedblleft{}A Decline Rate Study of Norwegian Oil Production.\textquotedblright{}
Energy Policy 36, no. 11 (November 2008): 4262\textendash{}71. doi:10.1016/j.enpol.2008.07.039.

\bibitem{Hubbert 1956}Hubbert, Marion King. Nuclear Energy and the
Fossil Fuels. Vol. 95. Shell Development Company, Exploration and
Production Research Division Houston, TX, 1956. http://energycrisis.biz/hubbert/1956/1956.pdf.

\bibitem{Laherr=0000E8re 2002}Laherrere, Jean. \textquotedblleft{}Forecasting
Future Production from Past Discovery.\textquotedblright{} International
Journal of Global Energy Issues 18, no. 2 (2002): 218\textendash{}38.

\bibitem{Laherr=0000E8re and Sornette 1998}Laherrere, Jean, and Didier
Sornette. \textquotedblleft{}Stretched Exponential Distributions in
Nature and Economy:\textquoteleft{}fat Tails\textquoteright{} with
Characteristic Scales.\textquotedblright{} The European Physical Journal
B-Condensed Matter and Complex Systems 2, no. 4 (1998): 525\textendash{}39.

\bibitem{Lutz 2012}Lutz, Christian, Ulrike Lehr, and Kirsten S. Wiebe.
\textquotedblleft{}Economic Effects of Peak Oil.\textquotedblright{}
Energy Policy 48 (September 2012): 829\textendash{}34. doi:10.1016/j.enpol.2012.05.017.

\bibitem{Lynch 2002}Lynch, Michael C. \textquotedblleft{}Forecasting
Oil Supply: Theory and Practice.\textquotedblright{} The Quarterly
Review of Economics and Finance 42, no. 2 (2002): 373\textendash{}89.

\bibitem{Murray 2012}Murray, James W., and Jim Hansen. \textquotedblleft{}Peak
Oil and Energy Independence: Myth and Reality.\textquotedblright{}
Eos, Transactions American Geophysical Union 94, no. 28 (2013): 245\textendash{}46.

\bibitem{Norwegian Data}Norwegian Petroleum Directorate (2014, Mai
25). Factpages. Retrieved from \href{http://factpages.npd.no/factpages/}{http://factpages.npd.no/factpages/}.

\bibitem{Smith 1980}Smith, J. L. ``A Probablilistic Model of Oil
Discovery.'' The Review of Economics and Statistics, Vol. 62, No.
4 (Nov. 1980): 587-94.

\end{thebibliography}
\newpage{}

\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{oseberg}

\caption{Example of a regular field with $\beta=0.66\pm0.01$ and $\tau=-55\pm1$.}

\label{regular} 
\end{figure}

\newpage{}

\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{vale}

\caption{Example of a irregular field.}

\label{irregular} 
\end{figure}

\newpage{}

\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{skarv}

\caption{Example of a new field.}

\label{new} 
\end{figure}

\newpage{}

\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{oesberg-error}

\caption{Oesberg field - Relative error on predicted total production from
time $t$ until 2014}
\label{oesberg-error}
\end{figure}
\begin{figure}[p]
\includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
\caption{Logistic fit to the number of discoveries for Norway. The discovery
rate of new oil fields is dependent on their size.}

\label{rate_of_discovery} 
\end{figure}
\begin{figure}[p]
\includegraphics[width=1\columnwidth]{NO-future}

\includegraphics[width=1\columnwidth]{UK-future} \caption{Monte-Carlo and fit forecast based on past production data for Norway
(top) and the U.K. (bottom). In both cases, the Mont-Carlo model forecasts
a significantly slower decay than the fit.}

\label{no_uk_2013} 
\end{figure}

\noindent 
\begin{figure}[p]
\includegraphics[width=1\columnwidth]{NO-back-test}

\includegraphics[width=1\columnwidth]{UK-back-test} \caption{Monte-Carlo and Hubbert forecast based on past production data up
to 2008 for Norway (top) and the U.K. (bottom). The results can be
compared with the subsequent oil production (blue area). In both cases,
the Monte-Carlo methodology is more precise.}

\label{model-back-testing} 
\end{figure}

\end{document}

### 2014-06-20 ABM/sim-base.py

@@ -1,25 +1,0 @@
-__author__ = 'lfi'
-
-
-class SimBase(object):
-    parameters = []
-    parameter_names = {}
-    variables = []
-    variable_names = {}
-
-    def __getattr__(self, name):
-        if name in self.variable_names and len(self.variables[self.variable_names[name]]) > 0:
-            return self.variables[self.variable_names[name]][-1]
-        if name in self.parameter_names:
-            return self.parameters[self.parameter_names[name]]
-        return super(SimBase, self).__getattr__(name)
-
-    def __setattr__(self, name, value):
-        if name in self.variable_names:
-            self.variables[self.variable_names[name]].append(value)
-        if name in self.parameter_names:
-            self.parameters[self.parameter_names[name]] = value
-        return super(SimBase, self).__setattr__(name, value)
-
-    def run(self, n):
-        for key in self.variable_names.keys()

### 2014-06-20 ABM/sim_abs.py

@@ -1,0 +1,60 @@
+__author__ = 'lfi'
+
+from numpy import exp
+
+from sim_base import SimBase
+
+
+class SimABS(SimBase):
+    parameters = {
+        'r': 0.05,
+        'y': 0.1,
+        'ps': 2,
+        'v': 0.1,
+        'g': 0.1,
+        'a': 1.0,
+        'o': 1.0,
+        'b': 0.5,
+        'n': 0.5,
+        'alpha': 0.5,
+    }
+    variable = {
+        'n1': [0.5, 0.5],
+        'n2': [0.5, 0.5],
+        'p': [1.0, 1.0],
+        'U1': [0.0, 0.0],
+        'U2': [0.0, 0.0],
+    }
+
+    def update_p(self):
+        return self.n1 * (self.ps + self.v * (self.p_1 - self.ps)) +\
+            self.n2 * (self.p_1 + self.g * (self.p_1 - self.p_2)) +\
+            self.y
+
+    def update_U1(self):
+        return (self.p_1 + self.y - (1 + self.r) * self.p_2) *\
+            (self.ps + self.v * (self.p_1 - self.ps) + self.y - (1 + self.r) * self.p_2) /\
+            (self.a * self.o) + self.n * self.U1_1
+
+    def update_U2(self):
+        return (self.p_1 + self.y - (1 + self.r) * self.p_2) *\
+            (self.p_1 + self.g * (self.p_1 - self.p_2) + self.y - (1 + self.r) * self.p_2) /\
+            (self.a * self.o) + self.n * self.U2_2
+
+    def update_n1(self):
+        Z = exp(self.b * self.U1_1) + exp(self.b * self.U2_1)
+        nt2 = exp(self.b * self.U2_1) / Z
+        return 1 - nt2 * exp(-(self.p_1 - self.ps) * (self.p_1 - self.ps) / self.alpha)
+
+    def update_n2(self):
+        Z = exp(self.b * self.U1_1) + exp(self.b * self.U2_1)
+        nt2 = exp(self.b * self.U2_1) / Z
+        return nt2 * exp(-(self.p_1 - self.ps) * (self.p_1 - self.ps) / self.alpha)
+
+if __name__ == '__main__':
+    print "Simulation start"
+    sim = SimABS()
+    print sim
+    for i in range(0, 50):
+        sim.run(1)
+    sim.plot("p")

### 2014-06-20 ABM/sim_base.py

@@ -1,0 +1,58 @@
+__author__ = 'lfi'
+
+import re
+import matplotlib.pyplot as plt
+
+
+class SimBase(object):
+    parameters = {}
+    variables = {}
+
+    def __getattr__(self, name, t=-1):
+        m = re.search('(\S*)_(\d*)$', name)
+        if not m is None:
+            name = m.group(1)
+            t = -int(m.group(2))
+        if name in self.variables and len(self.variables[name]) > 0:
+            return self.variables[name][t]
+        if name in self.parameters:
+            return self.parameters[name]
+        return None
+
+    def __setattr__(self, name, value):
+        if name in self.variables:
+            self.variables[name].append(value)
+        if name in self.parameters:
+            self.parameters[name] = value
+        return super(SimBase, self).__setattr__(name, value)
+
+    def run(self, n):
+        for i in range(0, n):
+            new_values = {}
+            for key in self.variables.keys():
+                new_values[key] = self.__getattribute__("update_{0}".format(key))()
+            for key in self.variables.keys():
+                self.__setattr__(key, new_values[key])
+
+    def plot(self, *args):
+        names = args[0:]
+        fig = plt.figure(figsize=(16, 10))
+        ax = fig.add_subplot(111)
+        ax.set_xlabel("Time")
+        ax.set_ylabel("Values")
+        for name in names:
+            if name in self.variables:
+                ax.plot(self.variables[name])
+        plt.show()
+
+    def __str__(self):
+        s = ""
+        s += "\n"
+        s += "----------\n"
+        s += "Simulation\n"
+        s += "----------\n"
+        for key in self.parameters.keys():
+            s += "{0}: {1}\n".format(key, self.__getattr__(key))
+        for key in self.variables.keys():
+            s += "{0}: {1}\n".format(key, self.__getattr__(key))
+        return s

### 2014-06-20 tomorrow_django/oil_and_gas/discovery.py

@@ -117,7 +117,7 @@ class DiscoveryGenerator:
 
     def init_scenarios(self):
         db_scenarios = DiscoveryScenario.objects.filter(country=self.country)
-        # db_scenarios.delete()
+        db_scenarios.delete()
 
         if db_scenarios.count() > 0:
             self.scenarios = db_scenarios.order_by('pdf').all()

### 2014-06-20 tomorrow_django/oil_and_gas/utils.py

@@ -73,15 +73,17 @@ def make_plot(x, y, x_label, y_label, x_fit=None, y_fit=None):
 
 
 def start_plot(x_label, y_label, title):
-    fig = plt.figure(figsize=(16,10))
+    fig = plt.figure(figsize=(16, 10))
     ax = fig.add_subplot(111)
     plt.title(title)
     ax.set_xlabel(x_label)
     ax.set_ylabel(y_label)
     font = {
         'family': 'Arial',
-        'weight': 'bold',
-        'size': 26
+        'weight': 'normal',
+        'size': 28,
     }
     matplotlib.rc('font', **font)
+    matplotlib.rcParams['lines.markersize'] = 20
+    matplotlib.rcParams['axes.labelsize'] = 26
     return plt

### 2014-06-20 tomorrow_django/temp.py

@@ -1,0 +1,173 @@
+__author__ = 'lfi'
+
+import matplotlib.pyplot as plt
+
+
+model = [46000000,
+         10619762.0871,
+         558634074.635,
+         619849136.207,
+         2433123.17564,
+         433929459.862,
+         4023762.6406,
+         448263303.093,
+         18000000,
+         940060927.184,
+         30000000,
+         2917719073.7,
+         1023470079.91,
+         72511110.1662,
+         2000000,
+         288370274.506,
+         35000000,
+         1217866.7138,
+         20967411.1968,
+         128903017.51,
+         56000000,
+         640568703.647,
+         2347342699.77,
+         502706496.408,
+         234111436.726,
+         988095196.413,
+         41051598.1263,
+         76455894.4087,
+         35672420.6156,
+         2000000,
+         2383442.92785,
+         149464113.477,
+         147503577.028,
+         143586568.211,
+         8000000,
+         43359227.841,
+         2000000,
+         117600168.325,
+         91933597.047,
+         189731060.19,
+         562126607.215,
+         2528569891.44,
+         145784385.91,
+         511847928.683,
+         12930044.9675,
+         4296013.31573,
+         134317558.472,
+         12404391.0794,
+         1585869202.55,
+         3610446987.03,
+         233186044.284,
+         241321458.929,
+         57991,
+         62564773.7361,
+         66313224.2848,
+         24000000,
+         176300364.539,
+         359176817.192,
+         1982867141.46,
+         9956932.54015,
+         22996726.1572,
+         143926117.41,
+         513731534.432,
+         55988336.5572,
+         10016294.9875,
+         805275319.547,
+         103271002.117,
+         364983815.778,
+         76000000,
+         606256344.014,
+         67021196.3377,
+         204923303.507,
+         64604262.6806,
+         50000000,
+         2000553.04767,
+         ]
+
+data = [46544520,
+        10692660,
+        259139760,
+        644075520,
+        1886940,
+        433367220,
+        5660820,
+        370469220,
+        18240420,
+        932148360,
+        30191040,
+        3543044340,
+        852896880,
+        75477600,
+        2515920,
+        204418500,
+        35222880,
+        1257960,
+        18869400,
+        74219640,
+        55979220,
+        828995640,
+        2326597020,
+        383677800,
+        228948720,
+        1152920340,
+        40883700,
+        64784940,
+        32077980,
[…]

### 2014-06-23 tomorrow_django/oil_and_gas/fit_logistic.py
Latest version.

@@ -1,8 +1,9 @@
 import numpy as np
 
-from mx.DateTime.DateTime import Timestamp
+# from mx.DateTime.DateTime import Timestamp
 from numpy import array, exp, arange, where, ndarray
 from datetime import datetime
+from psycopg2._psycopg import Timestamp
 from scipy.stats import t
 from scipy.optimize import fmin
 import matplotlib.pyplot as plt

### 2014-06-23 tomorrow_django/oil_and_gas/fitting.py

@@ -18,9 +18,10 @@
 #
 
 from datetime import datetime
-from mx.DateTime.DateTime import Timestamp
+# from mx.DateTime.DateTime import Timestamp
 
 from numpy import array, ndarray, exp, log, sign, mean
+from psycopg2._psycopg import Timestamp
 
 from scipy.optimize import fmin
 from scipy.stats import linregress

### 2014-06-23 tomorrow_django/temp.py

@@ -80,94 +80,366 @@ model = [46000000,
          2000553.04767,
          ]
 
-data = [46544520,
-        10692660,
-        259139760,
-        644075520,
-        1886940,
-        433367220,
-        5660820,
-        370469220,
-        18240420,
-        932148360,
-        30191040,
-        3543044340,
-        852896880,
-        75477600,
-        2515920,
-        204418500,
-        35222880,
-        1257960,
-        18869400,
-        74219640,
-        55979220,
-        828995640,
-        2326597020,
-        383677800,
-        228948720,
-        1152920340,
-        40883700,
-        64784940,
-        32077980,
-        22014300,
-        9434700,
-        148439280,
-        144036420,
-        174856440,
-        8176740,
-        37109820,
-        2515920,
-        66042900,
-        87428220,
-        172969500,
-        572371800,
-        2447361180,
-        403176180,
-        161647860,
-        16353480,
-        4402860,
-        87428220,
-        14466540,
-        1617736560,
-        3604055400,
-        249076080,
-        230206680,
-        77364540,
-        70445760,
-        59753100,
-        24530220,
-        153471120,
-        397515360,
-        1694472120,
-        10692660,
-        20756340,
-        208192380,
-        529601160,
-        43399620,
-        16353480,
-        922713660,
-        101265780,
-        342165120,
-        76735560,
-        436512120,
-        84283320,
-        219514020,
-        59753100,
-        116361300,
-        1886940,
-        ]
+no_data = [46544520,
+           10692660,
+           259139760,
+           644075520,
+           1886940,
+           433367220,
+           5660820,
+           370469220,
+           18240420,
+           932148360,
+           30191040,
+           3543044340,
+           852896880,
+           75477600,
+           2515920,
+           204418500,
+           35222880,
+           1257960,
+           18869400,
+           74219640,
+           55979220,
+           828995640,
+           2326597020,
+           383677800,
+           228948720,
+           1152920340,
+           40883700,
+           64784940,
[…]

### 2014-06-24 Publications/Forecasting  Future Oil Production/peak_oil_3.tex
Merge branch 'master' of https://github.com/LucasFievet/Tomorrow
Conflicts:
	Publications/Forecasting  Future Oil Production/peak_oil.lyx~

--- 2014-06-20 Publications/Forecasting  Future Oil Production/peak_oil.tex
@@
 
-\title{Forecasting future oil production}
+\title{Forecasting future oil production in Norway and the UK: a general improved methodology}
 
-\author{L. Fiévet, Z. Forró, M. del Degan, P. Cauwels and D. Sornette\\
-D-MTEC, ETH Zürich}
+\author{L. Fiévet, Z. Forró, %M. del Degan, 
+P. Cauwels and D. Sornette\\
+ETH Z\"urich\\
+Department of Management, Technology and Economics\\
+Chair of Entrepreneurial Risks}
 
-\date{05/25/14}
+%\date{}
 \maketitle
 \begin{abstract}
-\noindent This paper presents a new Monte-Carlo methodology to forecast
-the crude oil production of Norway and the U.K. based on the current/past
-performances of individual oil fields. By extrapolating the future
-production of these fields and the frequency of new discoveries, the
-oil production of these countries can be forecasted. The results indicate
-that standard methodology tends to underestimate remaining oil reserves.
-A comparison of the model to this methodology is made by making predictions
-with the full data and a back-test starting in 2008. It shows that
-the model gives a better description of future oil production.
+\noindent We present a new Monte-Carlo methodology to forecast
+the crude oil production of Norway and the U.K. based on a two-step process,
+(i) the nonlinear extrapolation of the current/past
+performances of individual oil fields and (ii) a stochastic model
+of the frequency of future oil field discoveries. Compared
+with the standard methodology that tends to underestimate remaining oil reserves,
+our method gives a better description of future oil production, as validated by
+our back-tests starting in 2008. Specifically, we predict remaining reserves
+extractable until 2030 to be $188 \pm 10$ million barrels for Norway and $90 \pm10$ million
+barrels for the UK, which are respectively $45\%$ and $66\%$ above the
+predictions using the standard methodology.
 \end{abstract}
@@
 has been a topic of active interest since the beginning of the past
-century because of oil's central role in our economy. Its importance
-ranges from energy production, through manufacturing to pharmaceuticals
+century because of oil central role in our economy. Its importance
+ranges from energy production, manufacturing to the pharmaceutical
 industry. As petroleum is a non-renewable and finite resource, it
@@
 come from reliable sources. In the past, a big share has been coming
-from Norway and the U.K. which have been reliable exporters. However,
+from Norway and the U.K., which have been reliable exporters. However,
 the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
@@
 That prediction has proven itself to be correct. His main argument
-was based on the finiteness of oil reserves and use of the logistic
-differential equation
-
+was based on the finiteness of oil reserves and to what amounts
+as the use of the logistic differential equation for the total
+quantity $P(t)$ of oil extracted up to time $t$
 \begin{equation}
-\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)\label{logistic}
+\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)~.
+\label{logistic}
 \end{equation}
-
-\noindent to model oil production. The logistic differential equation
+The logistic differential equation
 is characterized by an initial exponential growth, which then decreases
@@
 growth rate, and $K$ as the carrying capacity (total quantity of
-oil extracted). If $P(t)$ is the the amount of oil extracted up to
-time $t$, then $\frac{dP}{dt}$ is the oil production rate, the quantity
+oil that can be ultimately extracted). If $P(t)$ is the amount of oil extracted up to
+time $t$, then $f(t) := \frac{dP}{dt}$ is the oil production rate, the quantity
 that M. King Hubbert predicted with surprising accuracy to peak. From
@@
 Various extensions have been studied by Brandt \cite{Brandt 2006}
-to account for multi-cycled or asymmetric production curves.\\
+to account for multi-cycled or asymmetric production curves.
 The existing forecasts of future oil production use some form of the
 Hubbert model (Brecha, Laherrère or Lynch \cite{Brecha 2012,Laherr=0000E8re 2002,Lynch 2002})
-or economical model applied to aggregate production (Greiner \cite{Greiner 2012}),
+or some economical model applied to aggregate production (Greiner \cite{Greiner 2012}),
 but none goes into the details of studying the underlying dynamics.
 The main reason for the lack of details is certainly the lack of available
-data. In this paper, a new methodology is introduces to forecast future
+data.
+
+In this article, a new methodology is introduced to forecast future
 oil production. Instead of taking the aggregate oil production profile
@@
 extrapolating the future rate of discovery of new fields, the future
-oil production is forecasted by the means of a Monte Carlo simulation.
+oil production is forecasted by means of a Monte Carlo simulation.
 To demonstrate the universality of the methodology presented here,
-it is applied to 2 major oil producing countries with publicly available
+it is applied to two major oil producing countries with publicly available
 data: Norway\cite{Norwegian Data} and the U.K. \cite{U.K. Data}.
@@
 is the possibility to forecast non-trivial oil production profiles
-arising from the combination of all the individual field\textquoteright{}s
-dynamics. In order to achieve that, one must be able 1) to \textbf{extend
+arising from the combination of all the individual field
+dynamics. This ability reflects directly that the total
+production is indeed the sum of the contributions of the individual
+oil fields in production. Thus modelling at the level of each field
+reflects more closely the reality and is likely to be over-performing
+and more reliable, as we show below.
+In order to implement our approach, one must be able 1) to \textbf{extend
 the oil production of each individual field} into the future and 2)
@@
 up once the field is being developed, and then a peak or plateau before
-the field starts decaying. The decay can take many different shapes
-and is governed by a variety of geological and economical factors
-which will be captured implicitly by the fitting procedure.
+the oil field production starts decaying. The decay can take many different shapes
+and is governed by a variety of geological and economical factors.
+The goal of the fitting procedure is to capture as much as possible the impact of
+these different factors.
 
@@
 that goal. Figures \ref{regular}, \ref{irregular} and \ref{new}
-show that independently of the country, oil fields can be classified
-into 3 main categories:
+show that, independent of the country, oil fields can be classified
+into three main categories:
 \begin{itemize}
-\item \textbf{Regular fields} - Their decay show some regularity. 
-\item \textbf{Irregular fields} - The ones that don't decay in a regular
-fashion.
-\item \textbf{New fields} - The ones that don't decay yet. As such, there
+\item \textbf{Regular fields} - Their decays show some regularity (see figure \ref{regular} ); 
+\item \textbf{Irregular fields} - The ones that do not decay in a regular
+fashion (see figure \ref{irregular});
+\item \textbf{New fields} - The ones that do not decay yet. As such, there
 is no easy way to forecast their future oil production based on past
-data.
+data (see figure \ref{new}).
 \end{itemize}
@@
 As such, being able to model them is crucial. To capture as many different
-decay dynamics as possible, the stretched exponential
-
+decay dynamics as possible, the decay part of the oil production rate
+$f(t) := \frac{dP}{dt}$ has been fitted by the stretched exponential  
 \begin{equation}
-P(t)=P_{0}e^{sign(\tau)\left(\frac{t}{\left|\tau\right|}\right)^{\beta}}\label{stretched_exponential}
+f(t)=f_{0}e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
+\label{stretched_exponential}
 \end{equation}
-has been fitted to the decay of regular fields. The stretched exponential
-has many advantages as it generalized the power law and can therefore
+The stretched exponential function
+has many advantages as it generalises the power law and can therefore
 capture a broad variety of distributions as shown by Laherrère and
-Sornette \cite{Laherr=0000E8re and Sornette 1998}. As can be seen
-on figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
+Sornette \cite{Laherr=0000E8re and Sornette 1998}. Moreover,
+Malevergne et al. \cite{malpissor05} showed that the power law
+function can be obtained as an asymptotic case of the stretched
+exponential family, allowing for asymptotically nested statistical tests.  As can be seen
+in figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
 is a good functional form to fit the decay process of regular fields.
 
-For the minority of irregular fields, their decay has been modeled
-as follows: $\tau$ has been picked to be the average $\tau$ over
+For the minority of irregular fields, we assume 
+no difference in the decay of production between giants and dwarfs, which
+has been modelled as follows: the decay time scale $\tau$ has been picked to be the average $\tau$ over
 the regular fields. Then $\beta$ has been fixed so that the sum of
-the field\textquoteright{}s production over its lifetime be equal
+the field production over its lifetime be equal
 to the official ultimate recovery estimates, when such an estimate
@@
 The minority of new fields, which did not yet enter their decay phase,
-cannot be extrapolated and will therefore be treated as a new discovery.
-The technical details of how to treat them as a new discovery are
+cannot be extrapolated and will therefore be treated as new discoveries.
+The technical details of how to treat them as new discoveries are
 discussed in section \ref{sub:Future-production-from}.
@@
 \item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field is
-truncated at a certain date in the past $\tau\in\{0,\,\ldots,\, N\}$.
-Where $\tau$ is in months since production start of the field.
-\item The extrapolation$f(t)$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{\tau}\}$.
-\item The future production predicted by the extrapolation $f(t)$ can be
-compared to the actual production from the date $\tau$ in the past
-up to now $\tau_{f}=N$. The extrapolated total production can be
+truncated at a certain date in the past $T \in\{0,\,\ldots,\, N\}$, where
+$T$ is the time counted in months since the production start of the field.
+\item The extrapolation of the oil production rate
+$f(t) := \frac{dP}{dt}$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{T}\}$.
+\item The future production predicted by the extrapolation function $f(t)$ can be
+compared to the actual production from the date $T$ in the past
+up to the present $T_{f}=N$. The extrapolated total production can be
 computed as 
 \begin{equation}
-P_{e}(\tau)=\int_{\tau}^{\tau_{f}}f(t)dt
+P_{e}(T)=\int_{T}^{T_{f}}f(t)dt
 \end{equation}
@@
 \begin{equation}
-e(\tau)=\frac{P_{e}(\tau)-\sum_{i=\tau}^{\tau_{f}}p_{i}}{P_{e}(\tau)},
+e(T)=\frac{P_{e}(T)-\sum_{i=T}^{T_{f}}p_{i}}{P_{e}(T)},
+\label{ryjryukoik}
 \end{equation}
-where both are a function of the truncation time $\tau$.
+where both $P_{e}(T)$ and $e(T)$ are functions of the truncation time $T$.
 \end{itemize}
-Computing this back test for every month in the past since the field
-starting decaying, yields a plot showing the evolution of the relative
-error over time \ref{oesberg-error}. By construction, the relative
-error will tend to zero as the truncation time $\tau$ goes to the
-present. Nonetheless, it is a uself indicator for the stability of
-the extrapolation. As can been seen for the Oseberg field, the relative
+Computing this back-test, for every month in the past since the field production
+started decaying, yields a plot showing the evolution of the relative
+error over time defined by expression (\ref{oesberg-error}). By construction, the relative
+error will tend to zero as the truncation time $T$ approaches the
+present. Nonetheless, it is a useful indicator for the stability of
+the extrapolation. As can been seen for the Oseberg field in figure \ref{oesberg-error}, the relative
 error on future production remained fairly stable during the past
@@
 
-From the complete back-test, on can compute the average relative error
+From the complete back-test, we compute the average relative error
 \begin{equation}
@@
 \end{equation}
- the extrapolation made on the future production. Assuming that the
+of the extrapolation made on the future production. Assuming that the
 relative errors are normally distributed around the average relative
@@
 As the average relative error is often fairly constant, the extrapolation
-was corrected by the average relative error. Meaning that if the extrapolation
+was corrected by the average relative error, that is, if the extrapolation
 consistently over-estimated the production by 10\% during the back-test,
 the extrapolation was reduced by 10\%. This results in an extrapolated
-production $p(t)$, including a 1$\sigma$ confidence interval, as
+production $p(t)$, including a 1$\sigma$ confidence interval, given by
 \begin{equation}
-p(t)=(1-\bar{e})\, f(t)\pm\sigma_{e}.\label{eq:extrapolation}
+p(t)=(1-\bar{e})\, f(t)\pm\sigma_{e}.
+\label{eq:extrapolation}
 \end{equation}
-
 An example of such an extrapolation including a one standard deviation
-range is shown on figure \ref{irregular} for the Oesberg field.
+range is shown in figure \ref{regular} for the Oseberg field.
 
@@
 
-Once the individual fields have been extrapolated in the sense of
-formula \ref{eq:extrapolation}, the aim is to compute the extrapolation
-for the country. While it is straightforward to sum the extrapolations
+Once the individual fields have been extrapolated using
+formula \ref{eq:extrapolation}, we compute the extrapolation of the oil production
+for the whole country. While it is straightforward to sum the extrapolations
 of the individual fields to obtain the expected production, some care
-has to be taken with respect to the confidence interval on the production
+has to be taken with respect to the confidence interval of the production
 at the country level.
@@
 fields. To account for this observation, the assumption made is that
-the relative error between individual fields is uncorrelated. Therefore,
+the relative error between individual fields is uncorrelated. 
+While one could imagine that some inter-dependence could result from
+a coordinated response of supply to a sharp increase/decrease of demand,  we have not 
+observed this to be the case at a significant level. Therefore,
 the fields can be considered as a portfolio of assets with a return
@@
 and a risk given by $\sigma_{field}$ (eq. \ref{eq:field-risk}).
-This means that the risk at the country level from the extrapolated
-production is to be computed as 
+This means that the average standard deviation per field
+at the country level from the extrapolated production can be computed as 
 \begin{equation}
@@
 \end{equation}
-
 Intuitively, this models well the fact that the uncorrelated errors
@@
 As of today, giant oil fields are discovered much less frequently
-than dwarfs.
+than dwarf oil fields.
 \end{enumerate}
@@
 The Poisson process is a process that generates independent events
-at a rate $\lambda$. It is inhomogeneous if the rate is time-dependent,
+at a rate $\lambda$. It is nonhomogeneous if the rate is time-dependent,
 $\lambda\rightarrow\lambda(t)$. The standard way to measure $\lambda(t)$
@@
 of the cumulative number of events (discoveries) up to time $t$.
-Then, $\lambda(t)$ is simply $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
-shows $N(t)$ for Norwegian fields classified according to their size.
+Then, $\lambda(t)$ is simply a smoothed estimation 
+of the observed rate $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
+shows $N(t)$ for Norwegian fields classified according to their size in two classes, 
+dwarfs and giant fields. 
 The logistic curve is a good fit to the data (integral form of equation
@@
 observations. This same approach has already been successfully applied
-by Forro \cite{Forro 2012} to estimate the number of daily active
+by Forr\'o et al. \cite{Forro 2012} to estimate the number of daily active
 users on Zynga.
@@
 when fitting a logistic curve to data, that the carrying capacity
-can not be determined if the data does not already exhibit the slow
-down in growth from reaching the carrying capacity. However, it is
+can not be determined if the data does not already exhibit the slowdown 
+in growth towards the carrying capacity. However, it is
 mentioned in the literature that often dwarf fields have already been
@@
 in order to maximize the number of giant fields but large enough to
-avoid recent discoveries:
+avoid recent discoveries. Our definition is thus:
 \begin{itemize}
@@
 \end{itemize}
+We note that this definition differs by a factor 10 from the more standard one, for which
+oil fields with an ultimate recoverable resource of 0.5 billion barrels (Gb) 
+or higher are classified as giants, while oil fields with smaller URR are considered to be dwarfs
+\cite{H=0000F6=0000F6k 2008}.
+
 The resulting plot shown in figure \ref{rate_of_discovery} pictures
@@
 fit to the giant discoveries is well constrained, however the fit
-for the dwarfs is poorly constrained. As can be seen on figure \ref{rate_of_discovery},
+for the dwarfs is poorly constrained. As can be seen in figure \ref{rate_of_discovery},
 the carrying capacity $K$ of the logistic growth model is not well
 constrained by the available data. A large spectrum of values for
-$K$ can lead to an equally good fit of the data. Therefore, a method
-described in section \ref{sub:Likelihood-function-for} has been used
+$K$ can lead to an equally good fit of the data. We have not taken
+into account the possible effect that
+newly discovered dwarf fields could become smaller and smaller 
+until the new fields have a size that is so small to be economical to drill.
+This effect would cap the carrying capacity at a value smaller than 
+the one shown in figure  \ref{rate_of_discovery}. To address these issues, the method
+described in the next section \ref{sub:Likelihood-function-for} has been used
 to compute the probability of different carrying capacities.
@@
 
-To overcome the logistic fitting issue for the dwarf fields, a method
+To overcome the poor constraint on the carrying capacity
+$K$ obtained from the fitting procedure for dwarf fields, a method
 already used by Smith \cite{Smith 1980} has been implemented. This
@@
 population is random, with a probability of discovery being dependent 
-on reservoir size.''
+on (proportional to) reservoir size.''
 \end{enumerate}
@@
 is found and $m_{ij}$ denotes the number of fields of size $j\in\{1,\,\ldots,\, J\}$
-discovered before the $i^{th}$ step. Than the probability that the
-discovery at step $i^{th}$ is of size $j$ can be expressed as
+discovered before the $i^{th}$ step. Then, the probability that the
+discovery at step $i$ is of size $j$ can be expressed as
 \begin{equation}
@@
 The likelihood $L$ for a complete sequence of $N$ discoveries $\{I(1),\,\ldots,\, I(N)\}$
-can than be expressed as
+can then be expressed as
 \begin{equation}
@@
 that the scenario with the largest likelihood according to equation
-\ref{eq:likelihood} has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
+(\ref{eq:likelihood}) has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
 the bin $j$. Subsequently, the likelihood of each scenario (value
@@
 
-For the analysis of the north sea oil field discoveries, the number
+For the analysis of the discoveries of the North Sea oil fields, the number
 of size bins has been fixed to $J=2$ splitting between dwarfs (1)
@@
 \begin{table}[h]
-\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.}
-
+\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.
+For Norway, our logistic fit suggests that up to two new giant fields could be
+ultimately discovered. For the U.K., the prediction is more bleak, suggesting
+that the most probable scenario is that no other giant field will be found.}
 \centering %
@@
 with the intuitive expectation that discovering a new giant field
-is unlikely and that future discoveries wil mostly be made up of dwarf
+is unlikely and that future discoveries will mostly be made up of dwarf
 fields. The likelihoods obtained for the carrying capacities of dwarfs
@@
 fitted logistic curves, each curve being weighted by the likelihood
-of its carrying capacity given by equation \ref{eq:likelihood}.
+of its carrying capacity given by equation (\ref{eq:likelihood}).
 
@@
 
-The goal is to compute an expected oil production coming from future
+We now compute an expected oil production coming from future
 discoveries, which requires to combine the steps described in sections
@@
 However, this likelihood method does not give the time distribution
-of future discoveries. The aim is to use the likelihood function to
+of future discoveries. We propose to use the likelihood function to
 generate scenarios with their respective occurrence probability.
@@
 For a given scenario, the carrying capacity $K$ (= total number of
-fields) is given for each size class. This is useful, because it resolves
+fields) is given for each size class. This is useful to resolve
 the instability in fitting the logistic curve to the number of discovered
@@
 \begin{enumerate}
-\item Draw a scenario based on its probability.
+\item Draw a scenario based on its probability. {\bf DIDIER: Lucas, Peter expresses
+that he does not understand how this is done. Please give more detail on 
+how a scenario is chosen according to its probability and how the probability
+is determined. A quick synthesis here would be useful.}
 \item Compute the time distribution of new discoveries by fitting a logistic
 curve for each size class.
-\item For each discovery generate a size and the resulting production curve
+\item For each discovery, generate a size and the resulting production curve
 based on the size distribution and production curves of existing fields.
@@
 to the date in the past where it matches the current production from
-new fields.
+new fields. {\bf DIDIER: what is the meaning of ``shifted''? Peter does not fully
+understand what is done here, neither do I..}
 
@@
 simulating future oil production was straightforward. For each country,
-the existing fields were extrapolated and the future discoveries were
+the existing oil field productions were extrapolated and the future discoveries were
 simulated. Figure \ref{no_uk_2013} shows the average of 1000 simulations.
-For each country, the non-symmetric shape of the production's dynamics
-contrary to what Hubbert would have predicted is immediately noticeable.
+For each country, the non-symmetric shape of the production dynamics, which
+contradicts the prediction based on Hubbert's standard approach,
+is immediately noticeable.
 
 \begin{table}[H]
-\caption{Remaining oil reserves until 2030 in barrels for the \textbf{Fit}
-of the country's production, the Monte-Carlo \textbf{Model} and the
-relative difference $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$
-between the two.}
-
+\caption{Remaining oil reserves until 2030 in barrels 
+predicted by the extrapolation of the  \textbf{Fit}
+of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
+The relative difference between these two predictions 
+is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$.}
 \centering %
@@
 The results in table \ref{forecast-14} show a striking difference
-between the fit and the Monte-Carlo model forecast. According to the
+between the extrapolation of the fit and the Monte-Carlo model forecast. According to the
 fit (extrapolation of aggregate production), Norway's future oil production
@@
 \begin{itemize}
-\item The sum of the forecast of the individual existing fields is higher
+\item The sum of the forecast of the individual existing fields is larger
 than the extrapolation of the aggregate production.
@@
 \textquotedbl{}w shape\textquotedbl{} of the oil production profile.
-However, this has not been an issue to extrapolate the decay of production
+However, this has not been an issue for us to extrapolate the decay of production
 starting at the second peak. The difference between the Hubbert-based
@@
 
-Which of the two models is more thrust worthy? Clearly, the implications
+Which of the two models is more trustworthy? Clearly, the implications
 in adopting one methodology over the other are significant. The only
 way to answer this question is to back-test them. In other words:
-\textquotedbl{}What would have each of the models predicted had they
-been used in the past?\textquotedbl{}
+\textquotedbl{}What would each of the models have predicted, had they
+been used in the past?\textquotedbl{} The next section addresses
+this question and presents the validation step of our approach.
 
@@
 long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
-shows the outcome of those tests. Comparing the forecast of both models
+shows the results of these tests. Comparing the forecast of both models,
 with the oil production of the subsequent 6 years, shows that the
-predictive power of the Monte-Carlo model is better than a simple
-extrapolation of aggregate production. The following table \ref{backtest-08}
+predictive power of the Monte-Carlo model is significantly better than a simple
+extrapolation of the aggregate past production. The Monte-Carlo model 
+is found essentially right on target, while the extrapolation of past production
+(``fit'') is under-estimating the realised production by 19\% and 16\% respectively
+for Norway and the U.K. The following table \ref{backtest-08}
 summarizes the difference between the two approaches for the back-testing
@@
 \begin{table}[H]
-\caption{The fit and the Monte-Carlo model are used on the data set truncated
-in 2008 and their forecast for the period 2008-2014 is compared to
-the actual production.}
-
+\caption{Extrapolation of past oil production (``fit'') and prediction
+using the Monte-Carlo model are used on the data set truncated
+in 2008. Their forecast for the period 2008-2014 is compared to
+the actual realised production.}
 \centering %
@@
 
-The difference is not huge for the period from 2008 to 2014, but the
-Monte-Carlo model nevertheless is noticeably closer to the actual
-value. However, when considering the period after 2014, the difference
-is huge as shown in table \ref{forecast-08}.
+While the error of the extrapolation of the past oil production method (``fit'') is not dramatic
+over these six years from 2008 to 2014, the difference between it and the Monte-Carlo
+approach becomes huge from 2014 to 2030, as shown in table \ref{forecast-08}.
 
@@
 \caption{Remaining oil reserves forecasted for the period 2014-2030 when using
-the data truncated in 2008.}
-
+the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and 
+the Monte-Carlo model. The relative difference between these two predictions 
+is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$.}
 \centering %
@@
 tails in the decay process of individual fields and the new discoveries.
-As well, it must be noted that the simple extrapolation changed massively
+Moreover, it must be noted that the simple extrapolation changed massively
 between the back-test in table \ref{forecast-08} and the current
 fit in table \ref{forecast-14} (520\% for Norway and 236\% for the
-U.K.). In other words, the simple extrapolation was very unstable
-in its forecast, while the Monte-Carlo forecasts remained very consistent
+U.K.). In other words, the simple extrapolation is very unstable
+in its forecast, while the Monte-Carlo forecasts remains very consistent
 (less than 10\% change).
@@
 
-This paper presented a Monte-Carlo based methodology to forecast future
+We have presented a Monte-Carlo based methodology to forecast future
 oil production. By extending the oil production of current fields
@@
 
-The results are hopeful, as they make it highly likely that the decay
+The results suggest that it is highly likely that the decay
 of Norwegian and U.K. oil production will be much slower then one
@@
 the U.K. and 0.22M bbl/day for Norway. The import for the E.U. and
-Norway is a lower bound based on the changes in the U.K and Norway.}
-
+Norway is a lower bound based on the changes in the U.K and Norway. Negative
+numbers for Norway represent exports.}
 \centering %
@@
 model predicts that in 2030 the E.U. with Norway will need to increase
-its oil imports by $1.3$ million barrels. Imports which will most
-likely have to come from outside Europe.
+its oil imports by $1.3$ million barrels. These imports will most
+likely have to come from outside Europe, except for  
+non-standard oil sources yet to be developed.
+
+The present methodology can be applied to many other countries
+and geological areas, as well as updated at the level of the global
+oil production. Extensions to include the new wave of shale oil and
+non-standard oil can in principle be considered and constitute
+an interesting domain of application of our methodology for the future.
+
+\vskip 0.5cm
+\noindent
+{\bf Acknowledgements}: This work derives from a Master thesis at ETH Zurich authored by 
+Marc A. Del Degan (July 2012) under the supervision of the authors, 
+entitled ``Analysis of peak oil with focus on Norwegian oil production''
+(\url{http://www.er.ethz.ch/publications/MAS_Thesis_DelDegan_final_July12.pdf}), which
+itself derived from the methodology of pricing social network companies such as Zynga developed
+by three of the authors \cite{Forro 2012}.
+
 \begin{thebibliography}{10}
@@
 
-\bibitem{Forro 2012}Forró, Zalán, Peter Cauwels, and Didier Sornette.
+\bibitem{Forro 2012} Forró, Zalán, Peter Cauwels, and Didier Sornette.
 \textquotedblleft{}When Games Meet Reality: Is Zynga Overvalued?\textquotedblright{}
@@
 
-\bibitem{Greiner 2012}Greiner, Alfred, Willi Semmler, and Tobias
+\bibitem{Greiner 2012} Greiner, Alfred, Willi Semmler, and Tobias
 Mette. \textquotedblleft{}An Economic Model of Oil Exploration and
@@
 
+\bibitem{malpissor05}
+Malevergne, Yannick, Vladilen F. Pisarenko and Didier Sornette,
+Empirical Distributions of Log-Returns: between the Stretched
+Exponential and the Power Law? Quantitative Finance 5 (4), 379401 (2005)
+
 \bibitem{Murray 2012}Murray, James W., and Jim Hansen. \textquotedblleft{}Peak
@@
 \includegraphics[width=1\columnwidth]{oseberg}
-
-\caption{Example of a regular field with $\beta=0.66\pm0.01$ and $\tau=-55\pm1$.}
-
+\caption{Example of the time evolution of the oil production per day of
+a regular field, parameterised beyond the peak in the decay
+regime by the stretched exponential 
+function (\ref{stretched_exponential}) with $\beta=0.66\pm0.01$ and $\tau=-55\pm1$ months,
+shown with the black line.  The one standard deviation given by expression
+(\ref{eq:extrapolation}) is represented by the grey band.}
 \label{regular} 
@@
 \includegraphics[width=1\columnwidth]{vale}
-
 \caption{Example of a irregular field.}
-
 \label{irregular} 
@@
 \includegraphics[width=1\columnwidth]{skarv}
-
 \caption{Example of a new field.}
-
 \label{new} 
@@
 \includegraphics[width=1\columnwidth]{oesberg-error}
-
-\caption{Oesberg field - Relative error on predicted total production from
-time $t$ until 2014}
+\caption{Oesberg field - Relative error defined by expression (\ref{ryjryukoik})
+of the predicted total production from time $t$ indicated in the abscissa until 2014.
+One can observe that the predicted future total production is over-estimated by as much
+as 70\% in 1999, then under-estimated by the same amount in 2002, while the forecast
+errors remain smaller than 20\% since 2004.}
 \label{oesberg-error}
 \end{figure}
+
 \begin{figure}[p]
 \includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
-\caption{Logistic fit to the number of discoveries for Norway. The discovery
-rate of new oil fields is dependent on their size.}
-
+\caption{Logistic fit of the function solution of expression (\ref{logistic})
+to the number of discoveries for Norway. The discovery
+rate of new oil fields is dependent on their size as explained in the main text.}
 \label{rate_of_discovery} 
 \end{figure}
+
 \begin{figure}[p]
 \includegraphics[width=1\columnwidth]{NO-future}
-
-\includegraphics[width=1\columnwidth]{UK-future} \caption{Monte-Carlo and fit forecast based on past production data for Norway
-(top) and the U.K. (bottom). In both cases, the Mont-Carlo model forecasts
-a significantly slower decay than the fit.}
+\includegraphics[width=1\columnwidth]{UK-future} 
+\caption{Monte-Carlo (green upper
+continuous line with standard deviation bank starting in 2014 onward)
+and fit forecast based on past production data (lower line and grey one standard deviation band) for Norway
+(top) and the U.K. (bottom). In both cases, the Monte-Carlo model forecasts
+a significantly slower decay than the fit by taking into account that
+new fields will come in production.}
 
@@
 
-\includegraphics[width=1\columnwidth]{UK-back-test} \caption{Monte-Carlo and Hubbert forecast based on past production data up
-to 2008 for Norway (top) and the U.K. (bottom). The results can be
+\includegraphics[width=1\columnwidth]{UK-back-test} \caption{Monte-Carlo (green upper
+continuous line with standard deviation bank starting in 2014 onward) and Hubbert forecast based 
+on past production data up to 2008 (lower line and grey one standard deviation band)
+for Norway (top) and the U.K. (bottom). The results can be
 compared with the subsequent oil production (blue area). In both cases,

### 2014-06-24 tomorrow_django/temp.py
Latest version.

@@ -1,6 +1,7 @@
 __author__ = 'lfi'
 
 import matplotlib.pyplot as plt
+from pylab import *
 
 
 model = [46000000,
@@ -421,6 +422,46 @@ uk_model = [
     4007404.343,
     ]
 
+no_sizes = [
+    10619762,
+    # 558634075,
+    2433123,
+    1217867,
+    # 128903018,
+    2000000,
+    2383443,
+    # 117600168,
+    12930045,
+    4296013,
+    57991,
+    9956933,
+    # 143926117,
+    # 67021196,
+    # 64604263,
+    2000553,
+]
+
+no_returns = [
+    1.703632265,
+    # 18.49487746,
+    1.083893075,
+    0.328372173,
+    # 2.763312843,
+    0.328170123,
+    0.493425581,
+    # 9.230202839,
+    1.822981752,
+    0.614277813,
+    0.03679632,
+    2.105237766,
+    # 5.298257944,
+    # 19.10960206,
+    # 11.82362055,
+    0.862307348,
+]
+
+fit = polyfit(no_sizes, no_returns, 1)
+
 d_no_model = []
 d_data = []
 d_uk_model = []
@@ -442,4 +483,12 @@ if __name__ == '__main__':
     ax.plot(d_data)
     ax.plot(d_uk_model)
     plt.legend(("NO - Model", "NO - Data", "UK - Model"), loc='upper left')
+    plt.show()
+
+    fig = plt.figure(figsize=(16, 10))
+    ax = fig.add_subplot(111)
+    ax.set_xlabel("Size (barrels)")
+    ax.set_ylabel("Return (Revenue/Investments)")
+    plt.plot(no_sizes, no_returns, 'bo')
+    plt.plot(no_sizes, poly1d(fit)(no_sizes), '--k')
     plt.show()

### 2014-06-27 ABM/sim_abs.py
Merge branch 'master' of https://github.com/LucasFievet/Tomorrow
Conflicts:
	Publications/Forecasting  Future Oil Production/peak_oil.lyx~

@@ -7,21 +7,21 @@ from sim_base import SimBase
 
 class SimABS(SimBase):
     parameters = {
-        'r': 0.05,
-        'y': 0.1,
-        'ps': 2,
-        'v': 0.1,
-        'g': 0.1,
+        'r': 0.001,
+        'y': 1.0,
+        'ps': 2.0,
+        'v': 1.0,
+        'g': 1.9,
         'a': 1.0,
         'o': 1.0,
-        'b': 0.5,
-        'n': 0.5,
-        'alpha': 0.5,
+        'b': 2.0,
+        'n': 0.99,
+        'alpha': 1800,
     }
-    variable = {
+    variables = {
         'n1': [0.5, 0.5],
         'n2': [0.5, 0.5],
-        'p': [1.0, 1.0],
+        'p': [1.99, 2.01],
         'U1': [0.0, 0.0],
         'U2': [0.0, 0.0],
     }
@@ -51,10 +51,12 @@ class SimABS(SimBase):
         nt2 = exp(self.b * self.U2_1) / Z
         return nt2 * exp(-(self.p_1 - self.ps) * (self.p_1 - self.ps) / self.alpha)
 
+
 if __name__ == '__main__':
     print "Simulation start"
     sim = SimABS()
     print sim
-    for i in range(0, 50):
+    for i in range(0, 5):
         sim.run(1)
+        print sim
     sim.plot("p")

### 2014-06-27 ABM/sim_base.py

@@ -51,8 +51,8 @@ class SimBase(object):
         s += "----------\n"
         s += "Simulation\n"
         s += "----------\n"
-        for key in self.parameters.keys():
-            s += "{0}: {1}\n".format(key, self.__getattr__(key))
+        # for key in self.parameters.keys():
+        #     s += "{0}: {1}\n".format(key, self.__getattr__(key))
         for key in self.variables.keys():
             s += "{0}: {1}\n".format(key, self.__getattr__(key))
         return s

### 2014-07-09 ABM/sim_abs.py
Merge branch 'master' of https://github.com/LucasFievet/Tomorrow
Conflicts:
	Publications/Forecasting  Future Oil Production/peak_oil.lyx~

@@ -1,6 +1,6 @@
 __author__ = 'lfi'
 
-from numpy import exp
+from numpy import exp, random
 
 from sim_base import SimBase
 
@@ -8,9 +8,9 @@ from sim_base import SimBase
 class SimABS(SimBase):
     parameters = {
         'r': 0.001,
-        'y': 1.0,
+        'y_avg': 1.0,
         'ps': 1000,
-        'v': 1.0,
+        'v': 0.95,
         'g': 1.9,
         'a': 1.0,
         'o': 1.0,
@@ -19,35 +19,43 @@ class SimABS(SimBase):
         'alpha': 1800,
     }
     variables = {
-        'n1': [0.5, 0.5],
-        'n2': [0.5, 0.5],
-        'p': [920, 922],
-        'U1': [0.0, 0.0],
-        'U2': [0.0, 0.0],
+        'n1': [],
+        'n2': [],
+        'p': [920, 920, 920],
+        'U1': [0.0],
+        'U2': [0.0],
+        'y': [],
     }
+    updates = [
+        ["y", "n1", "n2"],
+        ["p"],
+        ["U1", "U2"],
+    ]
 
-    def update_p(self):
-        return self.n1 * (self.ps + self.v * (self.p_1 - self.ps)) +\
-            self.n2 * (self.p_1 + self.g * (self.p_1 - self.p_2)) +\
-            self.y
-
-    def update_U1(self):
-        return (self.p_1 + self.y - (1 + self.r) * self.p_2) *\
-            (self.ps + self.v * (self.p_1 - self.ps) + self.y - (1 + self.r) * self.p_2) /\
-            (self.a * self.o) + self.n * self.U1_1
-
-    def update_U2(self):
-        return (self.p_1 + self.y - (1 + self.r) * self.p_2) *\
-            (self.p_1 + self.g * (self.p_1 - self.p_2) + self.y - (1 + self.r) * self.p_2) /\
-            (self.a * self.o) + self.n * self.U2_2
+    def update_y(self):
+        return random.normal(self.y_avg, self.o)
 
     def update_n1(self):
-        nt2 = 1 / (exp(self.b * (self.U1_1 -self.U2_1)) + 1)
-        return 1 - nt2 * exp(-(self.p_1 - self.ps) * (self.p_1 - self.ps) / self.alpha)
+        return 1 - self.update_n2()
 
     def update_n2(self):
-        nt2 = 1 / (exp(self.b * (self.U1_1 -self.U2_1)) + 1)
[…]

### 2014-07-09 ABM/sim_base.py

@@ -10,7 +10,7 @@ class SimBase(object):
 
     def __getattr__(self, name, t=-1):
         m = re.search('(\S*)_(\d*)$', name)
-        if not m is None:
+        if m is not None:
             name = m.group(1)
             t = -int(m.group(2))
         if name in self.variables and len(self.variables[name]) > 0:
@@ -28,11 +28,12 @@ class SimBase(object):
 
     def run(self, n):
         for i in range(0, n):
-            new_values = {}
-            for key in self.variables.keys():
-                new_values[key] = self.__getattribute__("update_{0}".format(key))()
-            for key in self.variables.keys():
-                self.__setattr__(key, new_values[key])
+            for keys in self.updates:
+                new_values = {}
+                for key in keys:
+                    new_values[key] = self.__getattribute__("update_{0}".format(key))()
+                for key in keys:
+                    self.__setattr__(key, new_values[key])
 
     def plot(self, *args):
         names = args[0:]

### 2014-07-09 Sudoku/sudoku.py

@@ -267,10 +267,23 @@ class Sudoku(object):
             permutations += self.get_block_row_permutation_n(i, row)
         return permutations
 
+    def get_block_col_type(self, col):
+        p = self.get_block_col_permutation(col)
+        return "B" if 0 < p.count('a') < 3 else "A"
+
+    def get_block_row_type(self, row):
+        p = self.get_block_row_permutation(row)
+        return "B" if 0 < p.count('a') < 3 else "A"
+
     def swap(self, n1, n2):
         self.solution = [e if e != n1 else 0 for e in self.solution]
         self.solution = [e if e != n2 else n1 for e in self.solution]
         self.solution = [e if e != 0 else n2 for e in self.solution]
+
+    def invert(self):
+        rows = [self.get_row(i) for i in range(0, 9)]
+        for idx, row in enumerate(rows):
+            self.set_col(idx, row)
 
     def normalize_type(self):
         row_permutation = self.get_block_row_permutation_n
@@ -289,6 +302,25 @@ class Sudoku(object):
             permutations += self.get_block_col_permutation(i) + "\n"
         return permutations
 
+    def get_super_type(self):
+        permutations = ""
+        for i in range(0, 3):
+            permutations += self.get_block_row_type(i) + " - "
+            permutations += self.get_block_col_type(i) + "\n"
+        return permutations
+
+    def get_row_super_type(self):
+        permutations = ""
+        for i in range(0, 3):
+            permutations += self.get_block_row_type(i)
+        return permutations
+
+    def get_col_super_type(self):
+        permutations = ""
+        for i in range(0, 3):
+            permutations += self.get_block_col_type(i)
+        return permutations
+
     def block_col_permute(self, col1, col2):
         for i in range(0, 3):
             col11 = self.get_col(col1 * 3 + i)
@@ -298,8 +330,22 @@ class Sudoku(object):
     def block_row_permute(self, row1, row2):
         for i in range(0, 3):
             row11 = self.get_row(row1 * 3 + i)
-            self.set_row(row1 * 3 + i, self.get_col(row2 * 3 + i))
+            self.set_row(row1 * 3 + i, self.get_row(row2 * 3 + i))
             self.set_row(row2 * 3 + i, row11)
+
+    def normalize_block_rows(self):
+        row_super_type = self.get_row_super_type()
+        col_super_type = self.get_col_super_type()
+        if row_super_type.count('A') < col_super_type.count('A'):
+            self.invert()
[…]

### 2014-07-10 Publications/Forecasting  Future Oil Production/peak_oil_5.tex
Latest version.

--- 2014-06-24 Publications/Forecasting  Future Oil Production/peak_oil_3.tex
@@
 \usepackage{esint}
+\usepackage{url}
 \usepackage[unicode=true,pdfusetitle,
@@
 
-\author{L. Fiévet, Z. Forró, %M. del Degan, 
+\author{L. Fi\'evet, Z. Forr\'o, %M. del Degan, 
 P. Cauwels and D. Sornette\\
@@
 come from reliable sources. In the past, a big share has been coming
-from Norway and the U.K., which have been reliable exporters. However,
+from Norway and the U.K., two of Europe's biggest exporters. However,
 the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
@@
 The existing forecasts of future oil production use some form of the
-Hubbert model (Brecha, Laherrère or Lynch \cite{Brecha 2012,Laherr=0000E8re 2002,Lynch 2002})
+Hubbert model (Brecha, Laherr\`{e}re or Lynch \cite{Brecha 2012,Laherr=0000E8re 2002,Lynch 2002})
 or some economical model applied to aggregate production (Greiner \cite{Greiner 2012}),
@@
 oil production is forecasted by means of a Monte Carlo simulation.
-To demonstrate the universality of the methodology presented here,
+To demonstrate the generality of the methodology presented here,
 it is applied to two major oil producing countries with publicly available
@@
 arising from the combination of all the individual field
-dynamics. This ability reflects directly that the total
-production is indeed the sum of the contributions of the individual
-oil fields in production. Thus modelling at the level of each field
+dynamics. This ability reflects directly the fact that the total
+production is the sum of the contributions of each individual
+oil field in production. Thus modelling at the level of each field
 reflects more closely the reality and is likely to be over-performing
@@
 and is governed by a variety of geological and economical factors.
-The goal of the fitting procedure is to capture as much as possible the impact of
-these different factors.
+The goal of the fitting procedure is to capture as much the impact of
+these different factors as possible.
 
@@
 \begin{equation}
-f(t)=f_{0}e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
+f(t)=f_{0}~e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
 \label{stretched_exponential}
@@
 has many advantages as it generalises the power law and can therefore
-capture a broad variety of distributions as shown by Laherrère and
+capture a broad variety of distributions as shown by Laherr\`{e}re and
 Sornette \cite{Laherr=0000E8re and Sornette 1998}. Moreover,
@@
 until the new fields have a size that is so small to be economical to drill.
-This effect would cap the carrying capacity at a value smaller than 
-the one shown in figure  \ref{rate_of_discovery}. To address these issues, the method
+
+There are in fact two competing effects that are likely to compensate each other.
+On the one hand, figure \ref{ccdffieldsizes} shows the complementary 
+cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
+Two salient properties can be observed. First, the tails 
+of the distributions are well described by power laws
+\begin{equation}
+{\rm CCDF}(S) \sim {1 \over S^{1+\alpha}}~,
+\label{ccdfsizefield}
+\end{equation}
+with exponent $\alpha = 1.2 \pm 0.1$ for Norway and $\alpha = 1.4 \pm 0.1$ for the UK.
+The fact that the estimations of the exponents $\alpha$ are larger than $1$ implies
+that the cumulative oil reserves are asymptotically controlled by the largest fields, and not the small ones
+\cite{Sornettecritbook}. However, the fact that the exponents $\alpha$ are rather close
+to $1$ (which is called ``Zipf law'') would make the many small oil field contributing 
+significantly in total. This brings us to the second important feature exhibited by figure \ref{ccdffieldsizes},
+namely the roll-overs of the CCDFs for small oil fields, likely due to an under-sampling
+of the data. Indeed, as for most data sets involving broad distributions of sizes such
+as oil fields, the distributions are in general incomplete for the small events due to the non
+exhaustive sampling. This incompleteness raises the spectre that our extrapolations
+might be grossly underestimating the large potential contributions to the total reserves
+of the many small yet undiscovered small oil fields. Assuming that the power law (\ref{ccdfsizefield})
+would hold for smaller fields down to size of 1 Million barrels leads to a number of 
+such fields 10 to 100 times larger than presently known.
+
+But there is another
+key factor that needs to be considered, namely the fact that small oil fields
+are not economically viable for exploitation, which leads to an effective truncation
+in the distribution (\ref{ccdfsizefield}) relevant for the estimation of recoverable oil.
+Taking the data from the NPD\cite{Norwegian Data}
+providing the yearly investments broken by fields in Norway, let us consider the illustrative
+case of the field GAUPE. The investment spent to 
+develop its exploitation was  \$380M, while its estimated size is 
+$\simeq 1.2Mb$, which corresponds approximately to a total
+market value of $\$120M$ at \$100 per barrel. It is thus not a surprise that
+investment to exploit this field was interrupted in 2013. 
+Figure \ref{returnNorwayfield} shows the total estimated revenue divided
+by investment as a function of the estimated ultimate recovery for a number
+of small oil fields. It can be inferred that oil fields of sizes smaller than about 10 Mb
+are not economically viable as long as the market oil price does not 
+grow much higher than \$100 per barrel. This implies that, notwithstanding
+the large number of unknown small oil fields, economic considerations
+oblige us to neglect the small oil fields, therefore providing
+a justification of our procedure. In fact, economic constraints may lead to 
+cap the carrying capacity at a value smaller than 
+the one shown in figure  \ref{rate_of_discovery}.
+
+To address these issues from a more solid angle, the method
 described in the next section \ref{sub:Likelihood-function-for} has been used
@@
 ultimately discovered. For the U.K., the prediction is more bleak, suggesting
-that the most probable scenario is that no other giant field will be found.}
+that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
 \centering %
@@
 \begin{enumerate}
-\item Draw a scenario based on its probability. {\bf DIDIER: Lucas, Peter expresses
-that he does not understand how this is done. Please give more detail on 
-how a scenario is chosen according to its probability and how the probability
-is determined. A quick synthesis here would be useful.}
+\item Draw a scenario (total number of fields in each size bin) based on its probability according to the likelihood function (\ref{eq:likelihood}).
+This is done by generating a random number $r$ between 0 and 1, and computing the scenario that is mapped to $r$ by the cumulative distribution function of all scenarios.
 \item Compute the time distribution of new discoveries by fitting a logistic
@@
 
-In order to account for the new fields discussed in section \ref{sub:Regular,-irregular-and},
-which are already discovered but did not yet enter the decay phase,
-the simulated production resulting from new discoveries has been shifted
-to the date in the past where it matches the current production from
-new fields. {\bf DIDIER: what is the meaning of ``shifted''? Peter does not fully
-understand what is done here, neither do I..}
+Last but not least, it has to be defined how the expected production from these future oil field discoveries is added to the extrapolated production from existing fields.
+The start of the simulation of new discoveries does not match up with the date of the latest production data, the reason being that the new fields (defined in section \ref{sub:Regular,-irregular-and}), which are already discovered but did not yet enter the decay phase, are not taken into account in the simulation as their final size is not known.
+The only meaningful way of treating the new fields is to consider them as a discovery.
+Therefore, the starting point in time of the simulated production resulting from new discoveries has been choosen as the date in the past where it matches the current production from
+new fields.
+In order words, the extrapolated production from regular and irregular fields added to the production from simulated future discoveries (which is shifted into the past as to match the production from new fields) must be equal to the current (latest available data) total production from all fields.
 
@@
 U.K. & $0.7\cdot10^{6}$  & $0.9\cdot10^{6}$  & $1.0\cdot10^{6}$  & $1.1\cdot10^{6}$ \tabularnewline
-E.U.+Norway (lower bound) & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
+E.U.+Norway & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
 \bottomrule  &  &  &  & \tabularnewline
@@
 
-\bibitem{Forro 2012} Forró, Zalán, Peter Cauwels, and Didier Sornette.
+\bibitem{Forro 2012} Forr\'o, Zal\'an, Peter Cauwels, and Didier Sornette.
 \textquotedblleft{}When Games Meet Reality: Is Zynga Overvalued?\textquotedblright{}
@@
 
-\bibitem{H=0000F6=0000F6k 2008}Höök, Mikael, and Kjell Aleklett.
+\bibitem{H=0000F6=0000F6k 2008}H\"o\"ok, Mikael, and Kjell Aleklett.
 \textquotedblleft{}A Decline Rate Study of Norwegian Oil Production.\textquotedblright{}
@@
 
-\bibitem{Laherr=0000E8re 2002}Laherrere, Jean. \textquotedblleft{}Forecasting
+\bibitem{Laherr=0000E8re 2002}Laherr\`{e}re, Jean. \textquotedblleft{}Forecasting
 Future Production from Past Discovery.\textquotedblright{} International
@@
 
-\bibitem{Laherr=0000E8re and Sornette 1998}Laherrere, Jean, and Didier
+\bibitem{Laherr=0000E8re and Sornette 1998}Laherr\`{e}re, Jean, and Didier
 Sornette. \textquotedblleft{}Stretched Exponential Distributions in
@@
 
+\bibitem{Sornettecritbook} Sornette, D.,
+``Critical Phenomena in Natural Sciences, Chaos, Fractals, Self-organization and Disorder: Concepts and Tools'',
+2nd ed., Springer Series in Synergetics, Heidelberg, 2004.
+
 \end{thebibliography}
@@
 
+%FIGURE 1
 \noindent 
@@
 
-\newpage{}
-
+%FIGURE 2
 \noindent 
@@
 
+%FIGURE 3
 \noindent 
@@
 
+%FIGURE 4
 \noindent 
@@
 
+%FIGURE 5
 \begin{figure}[p]
@@
 
+%FIGURE 6
+\begin{figure}[p]
+\includegraphics[width=1\columnwidth]{ccdf-field-size}
+\caption{Complementary 
+cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
+The two dashes lines visualise the power law behaviour (\ref{ccdfsizefield}) of the tail
+of the distributions. }
+\label{ccdffieldsizes}
+\end{figure}
+
+%FIGURE 7
+\begin{figure}[p]
+\includegraphics[width=1\columnwidth]{no-returns-size}
+\caption{Total estimated revenue divided by investment as a function of the estimated 
+ultimate recovery for a number of small oil fields. }
+\label{returnNorwayfield}
+\end{figure}
+
+%FIGURE 8
 \begin{figure}[p]
@@
 new fields will come in production.}
-
 \label{no_uk_2013} 
@@
 
+%FIGURE 9
 \noindent 
@@
 \includegraphics[width=1\columnwidth]{NO-back-test}
-
 \includegraphics[width=1\columnwidth]{UK-back-test} \caption{Monte-Carlo (green upper
@@
 the Monte-Carlo methodology is more precise.}
-
 \label{model-back-testing} 

### 2014-07-10 Publications/Forecasting  Future Oil Production/peak_oil_6.tex
Latest version.

--- 2014-07-10 Publications/Forecasting  Future Oil Production/peak_oil_5.tex
@@
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[twoside]{article}
+\documentclass[review]{elsarticle}
 \usepackage{mathpazo}
@@
 \usepackage{url}
-\usepackage[unicode=true,pdfusetitle,
+\usepackage[unicode=true,
  bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=false]
+ breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
  {hyperref}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-%% Because html converters don't know tabularnewline
 \providecommand{\tabularnewline}{\\}
@@
 \@ifundefined{date}{}{\date{}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.2 (15/5/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
 
@@
 
-\usepackage{fancyhdr}% Headers and footers
  % All pages have headers and footers
@@
 
+\defcitealias{Commission2014}{European Commission}
+\defcitealias{Directorate2014}{NPD}
+
 \begin{document}
@@
 
-\author{L. Fi\'evet, Z. Forr\'o, %M. del Degan, 
-P. Cauwels and D. Sornette\\
-ETH Z\"urich\\
-Department of Management, Technology and Economics\\
-Chair of Entrepreneurial Risks}
+\author[eth-mtec]{L. Fi\'evet\corref{cor1}}
+\ead{lfievet@ethz.ch}
 
-%\date{}
+\author[eth-mtec]{Z. Forr\'o}
+\ead{zforro@ethz.ch}
+
+\author[eth-mtec]{P. Cauwels}
+\ead{pcauwels@ethz.ch}
+
+\author[eth-mtec]{D. Sornette}
+\ead{dsornette@ethz.ch}
+
+\cortext[cor1]{Corresponding author. Tel.: +41 44 632 83 79.}
+
+\address[eth-mtec]{Chair of Entrepreneurial Risks, ETH Z\"urich, Scheuchzerstrasse 7 (SEC F), CH-8092 Z\"urich, Switzerland}
+
 \maketitle
@@
 our back-tests starting in 2008. Specifically, we predict remaining reserves
-extractable until 2030 to be $188 \pm 10$ million barrels for Norway and $90 \pm10$ million
+extractable until 2030 to be $188 \pm 10$ million barrels for Norway and $98 \pm10$ million
 barrels for the UK, which are respectively $45\%$ and $66\%$ above the
@@
 has been growing due to stagnating supplies and high oil prices since
-the crisis in 2008/2009 \cite{Murray 2012}. As any industrialized
+the crisis in 2008/2009 \citep{Murray2013}. As any industrialized
 country, Europe is strongly dependent on oil supply to maintain its
@@
 the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
-production has been declining rapidly as well \cite{H=0000F6=0000F6k 2008}.\\
+production has been declining rapidly as well \citep{Hook2008}.\\
  The methodology behind forecasting future oil production has not
 evolved much since M. King Hubbert, who in 1956 famously predicted
-that the U.S. oil production would peak around 1965-1970 \cite{Hubbert 1956}.
+that the U.S. oil production would peak around 1965-1970 \citep{Hubbert1956}.
 That prediction has proven itself to be correct. His main argument
@@
 popularity in modeling future oil production given its simplicity.
-Various extensions have been studied by Brandt \cite{Brandt 2006}
+Various extensions have been studied by \citet{Brandt2007}
 to account for multi-cycled or asymmetric production curves.
 The existing forecasts of future oil production use some form of the
-Hubbert model (Brecha, Laherr\`{e}re or Lynch \cite{Brecha 2012,Laherr=0000E8re 2002,Lynch 2002})
-or some economical model applied to aggregate production (Greiner \cite{Greiner 2012}),
+Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
+or some economical model applied to aggregate production \citep{Greiner2011},
 but none goes into the details of studying the underlying dynamics.
@@
 it is applied to two major oil producing countries with publicly available
-data: Norway\cite{Norwegian Data} and the U.K. \cite{U.K. Data}.
+data: Norway \citepalias{Directorate2014} and the U.K. \citep{GOVUK2014}.
 
@@
 has many advantages as it generalises the power law and can therefore
-capture a broad variety of distributions as shown by Laherr\`{e}re and
-Sornette \cite{Laherr=0000E8re and Sornette 1998}. Moreover,
-Malevergne et al. \cite{malpissor05} showed that the power law
+capture a broad variety of distributions as shown by \citet{Laherrere1998}. Moreover,
+\citet{Malevergne2005} showed that the power law
 function can be obtained as an asymptotic case of the stretched
@@
 observations. This same approach has already been successfully applied
-by Forr\'o et al. \cite{Forro 2012} to estimate the number of daily active
+by \cite{Forro2012} to estimate the number of daily active
 users on Zynga.
@@
 discovered a long time ago, but their production has been postponed
-for economical reasons \cite[p. 378]{Lynch 2002}. Therefore, it is
+for economical reasons \cite[p. 378]{Lynch2002}. Therefore, it is
 expected that the large oil fields have mostly been found and produced,
@@
 or higher are classified as giants, while oil fields with smaller URR are considered to be dwarfs
-\cite{H=0000F6=0000F6k 2008}.
+\citep{Hook2008}.
 
@@
 newly discovered dwarf fields could become smaller and smaller 
-until the new fields have a size that is so small to be economical to drill.
+until the new fields have a size that is too small to be economical to drill.
 
@@
 that the cumulative oil reserves are asymptotically controlled by the largest fields, and not the small ones
-\cite{Sornettecritbook}. However, the fact that the exponents $\alpha$ are rather close
+\citep{Sornette2004}. However, the fact that the exponents $\alpha$ are rather close
 to $1$ (which is called ``Zipf law'') would make the many small oil field contributing 
@@
 in the distribution (\ref{ccdfsizefield}) relevant for the estimation of recoverable oil.
-Taking the data from the NPD\cite{Norwegian Data}
+Taking the data from the \citetalias{Directorate2014}
 providing the yearly investments broken by fields in Norway, let us consider the illustrative
@@
 $K$ obtained from the fitting procedure for dwarf fields, a method
-already used by Smith \cite{Smith 1980} has been implemented. This
+already used by \cite{Smith1980} has been implemented. This
 method makes the following two postulates:
@@
 Superposing the production curves results in the expected production
-curve from future oil field for a given scenario.
+curve from future oil fields for a given scenario.
 
@@
 The relative difference between these two predictions 
-is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$.}
+is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
 \centering %
@@
  & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
-\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $31\%$\tabularnewline
-U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $40\%$ \tabularnewline
+\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $45\%$\tabularnewline
+U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $66\%$ \tabularnewline
 \bottomrule  &  &  & \tabularnewline
@@
 would decay much faster than in the Monte-Carlo case. The remaining
-reserves estimated with the Monte-Carlo methodology are $30\%$ larger
+reserves estimated with the Monte-Carlo methodology are $45\%$ larger
 then the estimate from the fit. This difference originates from two
@@
 case. The former underestimates the remaining oil reserves by about
-$40\%$ compared to the latter.
+$66\%$ compared to the latter.
 
@@
 the Monte-Carlo model. The relative difference between these two predictions 
-is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$.}
+is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
 \centering %
@@
  & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
-\midrule Norway 08 & $25\cdot10^{6}$  & $171\cdot10^{6}$  & $-85\%$\tabularnewline
-U.K. 08 & $25\cdot10^{6}$  & $91\cdot10^{9}$  & $-73\%$ \tabularnewline
+\midrule Norway 08 & $25\cdot10^{6}$  & $171\cdot10^{6}$  & $-584\%$\tabularnewline
+U.K. 08 & $25\cdot10^{6}$  & $91\cdot10^{9}$  & $-264\%$ \tabularnewline
 \bottomrule  &  &  & \tabularnewline
@@
 These forecasts are significantly different from the ones obtained
-with a standard extrapolation. Indeed, our model forecasts 30\% to
-40\% more remaining oil reserves than the standard extrapolation.
+with a standard extrapolation. Indeed, our model forecasts 45\% to
+66\% more remaining oil reserves than the standard extrapolation.
 The back-test performed on the time period between 2008 and 2014 confirmed
@@
 Norway will vanish (currently accounting for 11\% of E.U. oil imports
-\cite{European oil imports}) and the U.K. will need to import more
+\citepalias{Commission2014}) and the U.K. will need to import more
 oil.
@@
 itself derived from the methodology of pricing social network companies such as Zynga developed
-by three of the authors \cite{Forro 2012}.
+by three of the authors \cite{Forro2012}.
 
-\begin{thebibliography}{10}
-\bibitem{Brandt 2006}Brandt, Adam R. \textquotedblleft{}Testing Hubbert.\textquotedblright{}
-Energy Policy 35, no. 5 (May 2007): 3074\textendash{}88. doi:10.1016/j.enpol.2006.11.004.
+\bibliographystyle{model5-names}
+\biboptions{authoryear}
+\bibliography{bibliography.bib}
 
-\bibitem{Brecha 2012}Brecha, Robert J. \textquotedblleft{}Logistic
-Curves, Extraction Costs and Effective Peak Oil.\textquotedblright{}
-Energy Policy 51 (December 2012): 586\textendash{}97. doi:10.1016/j.enpol.2012.09.016.
-
-\bibitem{European oil imports}European Commission - Energy (2014,
-Mai 25). EU Crude oil imports. Retrieved from \url{http://ec.europa.eu/energy/observatory/oil/import_export_en.htm}.
-
-\bibitem{Forro 2012} Forr\'o, Zal\'an, Peter Cauwels, and Didier Sornette.
-\textquotedblleft{}When Games Meet Reality: Is Zynga Overvalued?\textquotedblright{}
-Journal of Investment Strategies 1 (3), 119-145 (2012).
-
-\bibitem{U.K. Data}GOV.UK (2014, Mai 25). Exploration and production
-- Oil and gas: field data. Retrieved from \url{https://www.og.decc.gov.uk/pprs/full_production.htm}.
-
-\bibitem{Greiner 2012} Greiner, Alfred, Willi Semmler, and Tobias
-Mette. \textquotedblleft{}An Economic Model of Oil Exploration and
-Extraction.\textquotedblright{} Computational Economics 40, no. 4
-(December 2012): 387\textendash{}99. doi:10.1007/s10614-011-9272-0.
-
-\bibitem{H=0000F6=0000F6k 2008}H\"o\"ok, Mikael, and Kjell Aleklett.
-\textquotedblleft{}A Decline Rate Study of Norwegian Oil Production.\textquotedblright{}
-Energy Policy 36, no. 11 (November 2008): 4262\textendash{}71. doi:10.1016/j.enpol.2008.07.039.
-
-\bibitem{Hubbert 1956}Hubbert, Marion King. Nuclear Energy and the
-Fossil Fuels. Vol. 95. Shell Development Company, Exploration and
-Production Research Division Houston, TX, 1956. http://energycrisis.biz/hubbert/1956/1956.pdf.
-
-\bibitem{Laherr=0000E8re 2002}Laherr\`{e}re, Jean. \textquotedblleft{}Forecasting
-Future Production from Past Discovery.\textquotedblright{} International
-Journal of Global Energy Issues 18, no. 2 (2002): 218\textendash{}38.
-
-\bibitem{Laherr=0000E8re and Sornette 1998}Laherr\`{e}re, Jean, and Didier
-Sornette. \textquotedblleft{}Stretched Exponential Distributions in
-Nature and Economy:\textquoteleft{}fat Tails\textquoteright{} with
-Characteristic Scales.\textquotedblright{} The European Physical Journal
-B-Condensed Matter and Complex Systems 2, no. 4 (1998): 525\textendash{}39.
-
-\bibitem{Lutz 2012}Lutz, Christian, Ulrike Lehr, and Kirsten S. Wiebe.
-\textquotedblleft{}Economic Effects of Peak Oil.\textquotedblright{}
-Energy Policy 48 (September 2012): 829\textendash{}34. doi:10.1016/j.enpol.2012.05.017.
-
-\bibitem{Lynch 2002}Lynch, Michael C. \textquotedblleft{}Forecasting
-Oil Supply: Theory and Practice.\textquotedblright{} The Quarterly
-Review of Economics and Finance 42, no. 2 (2002): 373\textendash{}89.
-
-\bibitem{malpissor05}
-Malevergne, Yannick, Vladilen F. Pisarenko and Didier Sornette,
-Empirical Distributions of Log-Returns: between the Stretched
-Exponential and the Power Law? Quantitative Finance 5 (4), 379401 (2005)
-
-\bibitem{Murray 2012}Murray, James W., and Jim Hansen. \textquotedblleft{}Peak
-Oil and Energy Independence: Myth and Reality.\textquotedblright{}
-Eos, Transactions American Geophysical Union 94, no. 28 (2013): 245\textendash{}46.
-
-\bibitem{Norwegian Data}Norwegian Petroleum Directorate (2014, Mai
-25). Factpages. Retrieved from \href{http://factpages.npd.no/factpages/}{http://factpages.npd.no/factpages/}.
-
-\bibitem{Smith 1980}Smith, J. L. ``A Probablilistic Model of Oil
-Discovery.'' The Review of Economics and Statistics, Vol. 62, No.
-4 (Nov. 1980): 587-94.
-
-\bibitem{Sornettecritbook} Sornette, D.,
-``Critical Phenomena in Natural Sciences, Chaos, Fractals, Self-organization and Disorder: Concepts and Tools'',
-2nd ed., Springer Series in Synergetics, Heidelberg, 2004.
-
-\end{thebibliography}
 \newpage{}
@@
 
-\newpage{}
+\pagebreak
 
@@
 \end{figure}
-
-\newpage{}
 
@@
 %FIGURE 5
-\begin{figure}[p]
+\begin{figure}[H]
 \includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
@@
 
+\pagebreak
+
 %FIGURE 6
-\begin{figure}[p]
+\begin{figure}[H]
 \includegraphics[width=1\columnwidth]{ccdf-field-size}
@@
 
+\pagebreak
+
 %FIGURE 7
-\begin{figure}[p]
+\begin{figure}[H]
 \includegraphics[width=1\columnwidth]{no-returns-size}
@@
 %FIGURE 8
-\begin{figure}[p]
+\begin{figure}[H]
 \includegraphics[width=1\columnwidth]{NO-future}
@@
 \noindent 
-\begin{figure}[p]
+\begin{figure}[H]
 \includegraphics[width=1\columnwidth]{NO-back-test}

### 2014-07-10 ABM/sim_abs.py

@@ -1,6 +1,6 @@
 __author__ = 'lfi'
 
-from numpy import exp, random
+from numpy import exp, random, arange
 
 from sim_base import SimBase
 
@@ -10,13 +10,14 @@ class SimABS(SimBase):
         'r': 0.001,
         'y_avg': 1.0,
         'ps': 1000,
-        'v': 0.95,
+        'v': 0.9,
         'g': 1.9,
         'a': 1.0,
         'o': 1.0,
         'b': 2.0,
         'n': 0.99,
         'alpha': 1800,
+        'noise': 25.0,
     }
     variables = {
         'n1': [],
@@ -25,12 +26,16 @@ class SimABS(SimBase):
         'U1': [0.0],
         'U2': [0.0],
         'y': [],
+        'epsilon': [],
     }
     updates = [
-        ["y", "n1", "n2"],
+        ["epsilon", "y", "n1", "n2"],
         ["p"],
         ["U1", "U2"],
     ]
+
+    def update_epsilon(self):
+        return random.normal(0, self.noise)
 
     def update_y(self):
         return random.normal(self.y_avg, self.o)
@@ -39,13 +44,13 @@ class SimABS(SimBase):
         return 1 - self.update_n2()
 
     def update_n2(self):
-        nt2 = 1 / (exp(self.b * (self.U1 -self.U2)) + 1)
+        nt2 = 1 / (exp(self.b * (self.U1 - self.U2)) + 1)
         return nt2 * exp(-(self.p - self.ps) * (self.p - self.ps) / self.alpha)
 
     def update_p(self):
         return self.n1 * (self.ps + self.v * (self.p - self.ps)) +\
             self.n2 * (self.p + self.g * (self.p - self.p_2)) +\
-            self.y_avg
+            self.y_avg + self.epsilon
 
     def update_U1(self):
         return (self.p + self.y - (1 + self.r) * self.p_2) *\
@@ -63,7 +68,10 @@ if __name__ == '__main__':
     sim = SimABS()
     sim.print_parameters()
     print sim
-    for i in range(0, 100):
-        sim.run(1)
-        print sim
-    sim.plot("p")
+    sim.scan_parameter("v", arange(0.5, 1.01, 0.01), "std")
+    # sim.run_until("std", 1E-4)
+    # for i in range(0, 100000):
+    #     sim.run(1)
+        # print sim
+    # sim.print_diff_information("p")
+    # sim.plot("p")

### 2014-07-10 ABM/sim_base.py

@@ -1,7 +1,11 @@
 __author__ = 'lfi'
 
+import copy
 import re
 import matplotlib.pyplot as plt
+import numpy as np
+
+from scipy.stats import skew, kurtosis
 
 
 class SimBase(object):
@@ -34,6 +38,53 @@ class SimBase(object):
                     new_values[key] = self.__getattribute__("update_{0}".format(key))()
                 for key in keys:
                     self.__setattr__(key, new_values[key])
+
+    def run_until(self, name, upper=1E-3):
+        old_value, new_value = 0.0, 1.0
+        while abs(new_value - old_value) > upper:
+            self.run(1000)
+            old_value = new_value
+            new_value = self.__getattribute__("diff_{0}".format(name))("p")
+            # print "Change {0}: {1}.".format(name, new_value - old_value)
+        return new_value
+
+    def scan_parameter(self, parameter_name, values, property_name):
+        initial_variables = copy.copy(self.variables)
+        for value in values:
+            self.variables = initial_variables
+            self.__setattr__(parameter_name, value)
+            property_value = self.run_until(property_name)
+            print "{0}={1} => {2}={3}".format(parameter_name, value, property_name, property_value)
+
+    def diff_method(self, name, method):
+        array = np.array(self.variables[name])
+        array_diff = np.diff(array)
+        return method(array_diff)
+
+    def diff_mean(self, name):
+        return self.diff_method(name, np.mean)
+
+    def diff_median(self, name):
+        return self.diff_method(name, np.median)
+
+    def diff_std(self, name):
+        return self.diff_method(name, np.std)
+
+    def diff_skewness(self, name):
+        return self.diff_method(name, skew)
+
+    def diff_kurtosis(self, name):
+        return self.diff_method(name, kurtosis)
+
+    def print_diff_information(self, name):
+        print "----------"
+        print "Variable {0}".format(name)
+        print "----------"
+        print "Mean: {0}".format(self.diff_mean(name))
+        print "Median: {0}".format(self.diff_median(name))
+        print "Standard: {0}".format(self.diff_std(name))
+        print "Skewness: {0}".format(self.diff_skewness(name))
+        print "Kurtosis: {0}".format(self.diff_kurtosis(name))
 
     def plot(self, *args):
         names = args[0:]

### 2014-07-16 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

--- 2014-07-10 Publications/Forecasting  Future Oil Production/peak_oil_6.tex
@@
 our back-tests starting in 2008. Specifically, we predict remaining reserves
-extractable until 2030 to be $188 \pm 10$ million barrels for Norway and $98 \pm10$ million
+extractable until 2030 to be $5.7 \pm 0.3$ million barrels for Norway and $3.0 \pm0.3$ billion
 barrels for the UK, which are respectively $45\%$ and $66\%$ above the
@@
 
-\begin{table}[h]
-\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.
-For Norway, our logistic fit suggests that up to two new giant fields could be
-ultimately discovered. For the U.K., the prediction is more bleak, suggesting
-that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
-\centering %
-\begin{tabular}{lllll}
- & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
-\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
-U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
-\bottomrule  &  &  &  & \tabularnewline
-\end{tabular}\label{likelihood-results}
-\end{table}
-
 The results shown in table \ref{likelihood-results} are coherent
@@
 
-\begin{table}[H]
-\caption{Remaining oil reserves until 2030 in barrels 
-predicted by the extrapolation of the  \textbf{Fit}
-of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
-The relative difference between these two predictions 
-is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
-\centering %
-\begin{tabular}{lllr}
- & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
-\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $45\%$\tabularnewline
-U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $66\%$ \tabularnewline
-\bottomrule  &  &  & \tabularnewline
-\end{tabular}\label{forecast-14}
-\end{table}
-
 The results in table \ref{forecast-14} show a striking difference
@@
 
-\begin{table}[H]
-\caption{Extrapolation of past oil production (``fit'') and prediction
-using the Monte-Carlo model are used on the data set truncated
-in 2008. Their forecast for the period 2008-2014 is compared to
-the actual realised production.}
-\centering %
-\begin{tabular}{llll}
- & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
-\midrule Norway & $133\cdot10^{6}$  & $108\cdot10^{6}$  & $130\cdot10^{6}$ \tabularnewline
-U.K. & $79\cdot10^{6}$  & $66\cdot10^{6}$  & $75\cdot10^{9}$ \tabularnewline
-\bottomrule  &  &  & \tabularnewline
-\end{tabular}\label{backtest-08}
-\end{table}
-
 While the error of the extrapolation of the past oil production method (``fit'') is not dramatic
@@
 approach becomes huge from 2014 to 2030, as shown in table \ref{forecast-08}.
-
-\begin{table}[H]
-\caption{Remaining oil reserves forecasted for the period 2014-2030 when using
-the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and 
-the Monte-Carlo model. The relative difference between these two predictions 
-is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
-\centering %
-\begin{tabular}{lllc}
- & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
-\midrule Norway 08 & $25\cdot10^{6}$  & $171\cdot10^{6}$  & $-584\%$\tabularnewline
-U.K. 08 & $25\cdot10^{6}$  & $91\cdot10^{9}$  & $-264\%$ \tabularnewline
-\bottomrule  &  &  & \tabularnewline
-\end{tabular}\label{forecast-08}
-\end{table}
 
@@
 
-\begin{table}[H]
-\caption{Oil import (bbl/day) at a constant consumption of 1.5M bbl/day for
-the U.K. and 0.22M bbl/day for Norway. The import for the E.U. and
-Norway is a lower bound based on the changes in the U.K and Norway. Negative
-numbers for Norway represent exports.}
-\centering %
-\begin{tabular}{lllcc}
- & 2014 & 2020 & 2025 & 2030\tabularnewline
-\midrule Norway & $-1.23\cdot10^{6}$  & $-0.88\cdot10^{6}$  & $-0.58\cdot10^{6}$  & $-0.43\cdot10^{6}$ \tabularnewline
-U.K. & $0.7\cdot10^{6}$  & $0.9\cdot10^{6}$  & $1.0\cdot10^{6}$  & $1.1\cdot10^{6}$ \tabularnewline
-E.U.+Norway & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
-\bottomrule  &  &  &  & \tabularnewline
-\end{tabular}\label{needs}
-\end{table}
-
 As shown in table \ref{needs}, at constant consumption, the Monte-Carlo
 model predicts that in 2030 the E.U. with Norway will need to increase
-its oil imports by $1.3$ million barrels. These imports will most
+its daily oil imports by $1.3$ million barrels. These imports will most
 likely have to come from outside Europe, except for  
@@
 \caption{Example of the time evolution of the oil production per day of
-a regular field, parameterised beyond the peak in the decay
+a regular field, parametrised beyond the peak in the decay
 regime by the stretched exponential 
@@
 
+\pagebreak
+
+%TABLE 1
+
+\begin{table}[h]
+\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.
+For Norway, our logistic fit suggests that up to two new giant fields could be
+ultimately discovered. For the U.K., the prediction is more bleak, suggesting
+that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
+\centering %
+\begin{tabular}{lllll}
+ & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
+\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
+U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
+\bottomrule  &  &  &  & \tabularnewline
+\end{tabular}\label{likelihood-results}
+\end{table}
+
+\pagebreak
+
+%TABLE 2
+
+\begin{table}[H]
+\caption{Remaining oil reserves until 2030 in barrels 
+predicted by the extrapolation of the  \textbf{Fit}
+of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
+The relative difference between these two predictions 
+is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
+\centering %
+\begin{tabular}{lllr}
+ & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
+\midrule Norway & $3.95\cdot10^{9}$  & $5.72\cdot10^{9}$  & $45\%$\tabularnewline
+U.K. & $1.79\cdot10^{9}$  & $2.98\cdot10^{9}$  & $66\%$ \tabularnewline
+\bottomrule  &  &  & \tabularnewline
+\end{tabular}\label{forecast-14}
+\end{table}
+
+\pagebreak
+
+%TABLE 3
+
+\begin{table}[H]
+\caption{Extrapolation of past oil production (``fit'') and prediction
+using the Monte-Carlo model are used on the data set truncated
+in 2008. Their forecast for the period 2008-2014 is compared to
+the actual realised production.}
+\centering %
+\begin{tabular}{llll}
+ & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
+\midrule Norway & $4.05\cdot10^{9}$  & $3.28\cdot10^{9}$  & $3.95\cdot10^{9}$ \tabularnewline
+U.K. & $2.40\cdot10^{9}$  & $2.00\cdot10^{9}$  & $2.28\cdot10^{9}$ \tabularnewline
+\bottomrule  &  &  & \tabularnewline
+\end{tabular}\label{backtest-08}
+\end{table}
+
+\pagebreak
+
+%TABLE 4
+
+\begin{table}[H]
+\caption{Remaining oil reserves forecasted for the period 2014-2030 when using
+the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and 
+the Monte-Carlo model. The relative difference between these two predictions 
+is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
+\centering %
+\begin{tabular}{lllc}
+ & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
+\midrule Norway 08 & $0.76\cdot10^{9}$  & $5.20\cdot10^{9}$  & $-584\%$\tabularnewline
+U.K. 08 & $0.76\cdot10^{9}$  & $2.77\cdot10^{9}$  & $-264\%$ \tabularnewline
+\bottomrule  &  &  & \tabularnewline
+\end{tabular}\label{forecast-08}
+\end{table}
+
+\pagebreak
+
+%TABLE 5
+
+\begin{table}[H]
+\caption{Oil import (bbl/day) at a constant consumption of 1.5M bbl/day for
+the U.K. and 0.22M bbl/day for Norway. The import for the E.U. and
+Norway is a lower bound based on the changes in the U.K and Norway. Negative
+numbers for Norway represent exports.}
+\centering %
+\begin{tabular}{lllcc}
+ & 2014 & 2020 & 2025 & 2030\tabularnewline
+\midrule Norway & $-1.23\cdot10^{6}$  & $-0.88\cdot10^{6}$  & $-0.58\cdot10^{6}$  & $-0.43\cdot10^{6}$ \tabularnewline
+U.K. & $0.7\cdot10^{6}$  & $0.9\cdot10^{6}$  & $1.0\cdot10^{6}$  & $1.1\cdot10^{6}$ \tabularnewline
+E.U.+Norway & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
+\bottomrule  &  &  &  & \tabularnewline
+\end{tabular}\label{needs}
+\end{table}
+
 \end{document}

### 2014-07-16 ABM/sim_abs.py

@@ -68,7 +68,16 @@ if __name__ == '__main__':
     sim = SimABS()
     sim.print_parameters()
     print sim
-    sim.scan_parameter("v", arange(0.5, 1.01, 0.01), "std")
+    sim.scan_parameters(
+        [
+            {"name": "v", "values": list(arange(0.5, 1.01, 0.025))},
+            {"name": "g", "values": list(arange(1.5, 2.51, 0.025))},
+        ],
+        ["std", "skewness", "kurtosis"]
+    )
+    sim.plot_scan("v", "g", "std")
+    sim.plot_scan("v", "g", "skewness")
+    sim.plot_scan("v", "g", "kurtosis")
     # sim.run_until("std", 1E-4)
     # for i in range(0, 100000):
     #     sim.run(1)

### 2014-07-16 ABM/sim_base.py

@@ -1,16 +1,39 @@
 __author__ = 'lfi'
 
+import json
+import hashlib
 import copy
 import re
+import os
 import matplotlib.pyplot as plt
 import numpy as np
 
+from mpl_toolkits.mplot3d import Axes3D
 from scipy.stats import skew, kurtosis
 
 
 class SimBase(object):
+    name = "simulation"
     parameters = {}
     variables = {}
+    scans = {}
+
+    def __init__(self):
+        self.initial_variables = copy.copy(self.variables)
+
+    def reset_variables(self):
+        self.variables = copy.copy(self.initial_variables)
+
+    def compute_hash(self, parameters, initial_variables=None):
+        if initial_variables is None:
+            initial_variables = self.initial_variables
+        h = self.name + json.dumps(parameters) + json.dumps(initial_variables)
+        m = hashlib.sha224(h)
+        return m.hexdigest()
+
+    @property
+    def hash(self):
+        return self.compute_hash(self.parameters, self.initial_variables)
 
     def __getattr__(self, name, t=-1):
         m = re.search('(\S*)_(\d*)$', name)
@@ -45,16 +68,82 @@ class SimBase(object):
             self.run(1000)
             old_value = new_value
             new_value = self.__getattribute__("diff_{0}".format(name))("p")
-            # print "Change {0}: {1}.".format(name, new_value - old_value)
         return new_value
 
-    def scan_parameter(self, parameter_name, values, property_name):
-        initial_variables = copy.copy(self.variables)
+    def scan_parameters(self, parameters, properties, scan=None):
+        if not isinstance(parameters, list) or len(parameters) == 0:
+            raise "Parameters need to be a list with at least one item"
+
+        if not "name" in parameters[0] or not "values" in parameters[0]:
+            raise "Parameter needs a name and values attribute."
+
+        cache_file_hash = self.compute_hash(parameters)
+        cache_file = "{0}.txt".format(cache_file_hash)
+        if os.path.isfile(cache_file):
+            f = open(cache_file, 'r')
+            self.scans = json.loads(f.read())
+            f.close()
+            return -1
+
+        # Pop first parameter
+        first = False
+        name = parameters[0]["name"]
+        values = parameters[0]["values"]
+        parameters = parameters[1:]
+        print parameters
+
+        # If no scan, use self
+        if scan is None:
+            first = True
+            scan = self.scans
+
+        if not name in scan:
+            scan[name] = {'values': []}
+
+        property_until = properties[0]
+        count = 0
+
         for value in values:
[…]

### 2014-07-16 tomorrow_django/oil_and_gas/_constants.py

@@ -7,9 +7,186 @@ M3_TO_BARRELS = 1. / 6.2898
 START = {'NO': Timestamp('1971-06-01'), 'UK': Timestamp('1975-06-01')}
 
 BOUNDS = {'NO': {'2013-02-01': [1e6, 1e7],
-               '2003-02-01': [1e7],},
-        'UK': {'2013-02-01': [1e6, 1e7],
-               '1998-02-01': [1e6, 1e7],
-               '1993-02-01': [1e6, 1e7],},
-       }
+                 '2003-02-01': [1e7],},
+          'UK': {'2013-02-01': [1e6, 1e7],
+                 '1998-02-01': [1e6, 1e7],
+                 '1993-02-01': [1e6, 1e7],},
+          }
 MAX_DATE = Timestamp('2260-01-01')
+
+OIL_PRICE_SINCE_2000 = [
+    25.21,
+    27.15,
+    27.49,
+    23.45,
+    27.23,
+    29.62,
+    28.16,
+    29.41,
+    32.08,
+    31.4,
+    32.33,
+    25.28,
+    25.95,
+    27.24,
+    25.02,
+    25.66,
+    27.55,
+    26.97,
+    24.8,
+    25.81,
+    25.03,
+    20.73,
+    18.69,
+    18.52,
+    19.15,
+    19.98,
+    23.64,
+    25.43,
+    25.69,
+    24.49,
+    25.75,
+    26.78,
+    28.28,
+    27.53,
+    24.79,
+    27.89,
+    30.77,
+    32.88,
+    30.36,
+    25.49,
+    26.06,
+    27.91,
+    28.59,
+    29.68,
+    26.88,
+    29.01,
+    29.12,
+    29.95,
+    31.4,
+    31.32,
+    33.67,
+    33.71,
+    37.63,
+    35.54,
+    37.93,
+    42.08,
+    41.65,
+    46.87,
+    42.23,
+    39.09,
+    42.89,
+    44.56,
+    50.93,
+    50.64,
+    47.81,
+    53.89,
+    56.37,
+    61.87,
+    61.65,
+    58.19,
+    54.98,
+    56.47,
+    62.36,
+    59.71,
+    60.93,
+    68,
+    68.61,
+    68.29,
+    72.51,
+    71.81,
+    61.97,
+    57.95,
+    58.13,
+    61,
+    53.4,
+    57.58,
+    60.6,
+    65.1,
+    65.1,
+    68.19,
+    73.67,
[…]

### 2014-07-16 tomorrow_django/oil_and_gas/fit_logistic.py

@@ -196,7 +196,7 @@ def fit_double_cyclic(dates, values, init_guesses=2, **kwargs):
                             resid = cost_cyclic(res, _dates, values)
                             output.append((res, resid))
                             i += 1
-                            if i%100 == 0:
+                            if i % 100 == 0:
                                 print i
     output = array(sorted([_res for (_res,_) in output], key=lambda x: x[1]))    
     return output
@@ -208,10 +208,34 @@ def fit_logistic_r_p(k, x, y, r=2.5, p=5.5):
         (r, p),
         args=(k, x, y),
         maxiter=100000,
-        disp=0
+        disp=0,
     )
     residual = cost_function_r_p((r, p), k, x, y)
     return results[0], results[1], residual
+
+
+def fit_d_logistic_r_k_p(x, y, r=2.5, k=3.5, p=5.5):
+    results = fmin(
+        cost_function_d_logistic,
+        (r, k, p),
+        args=(x, y),
+        maxiter=100000,
+        disp=0,
+    )
+    residual = cost_function_d_logistic((r, k, p), x, y)
+    return results[0], results[1], results[2], residual
+
+
+def fit_cyclic_logistic(x, y, r1=2.5, k1=3.5, p1=5.5, t1=0.0, r2=2.5, k2=3.5, p2=5.5, t2=1.0):
+    r = fmin(
+        cost_cyclic_logistic,
+        (r1, k1, p1, t1, r2, k2, p2, t2),
+        args=(x, y),
+        maxiter=100000,
+        disp=0,
+    )
+    residual = cost_cyclic_logistic((r1, k1, p1, t1, r2, k2, p2, t2), x, y)
+    return r[0], r[1], r[2], r[3], r[4], r[5], r[6], r[7], residual
 
 
 def get_logistic(r, k, p):
@@ -231,6 +255,12 @@ def get_d_logistic(r, k, p):
     return func
 
 
+def cost_cyclic_logistic((r1, k1, p1, t1, r2, k2, p2, t2), x, y):
+    fit = get_d_logistic(r1, k1, p1)(x - t1) + get_d_logistic(r2, k2, p2)(x - t2)
+    res = (y - fit)**2 / abs(fit)
+    return sum(res)
+
+
 def cost_cyclic((qmax1, qmax2, tmax1, tmax2, a1, a2), x, y):
[…]

### 2014-07-16 tomorrow_django/oil_and_gas/hubbert.py

@@ -1,0 +1,89 @@
+import copy
+import numpy as np
+
+from datetime import date, timedelta
+
+from ._constants import OIL_PRICE_SINCE_2000
+from .utils import add_months, make_plot, diff_months_abs, start_plot
+from .models import CountryProduction
+from .fit_logistic import (fit_d_logistic_r_k_p, get_d_logistic, get_logistic,
+                           fit_cyclic_logistic)
+
+
+class HubbertBacktest():
+    def __init__(self):
+        pass
+
+    @classmethod
+    def run(cls, name):
+        ds, nks, ukfs, fs = [], [], [], []
+        d = date(year=2000, month=1, day=1)
+        d_max = date.today() - timedelta(days=120)
+        r = -1.0
+        k = 20E9
+        p = 1E8
+        r1 = 0.03
+        k1 = 10E9
+        p1 = 4E9
+        t1 = 120
+        r2 = 0.03
+        k2 = 10E9
+        p2 = 5E9
+        t2 = 300
+
+        while d < d_max:
+            # Norway
+            ncps = CountryProduction.objects.filter(
+                name="NO",
+                date__lte=d,
+            ).order_by("date")
+            cops = np.array(map(lambda x: x.production_oil, ncps))
+            months = np.array(range(0, len(cops)))
+            r, k, p, residual = fit_d_logistic_r_k_p(months, cops, r, k, p)
+            ds.append(copy.copy(d))
+            nks.append(k)
+            fs.append(k - get_logistic(r, k, p)(
+                len(cops) + diff_months_abs(d_max, d)
+            ))
+            print "{0}: {1}".format(d, k)
+
+            # UK
+            cps = CountryProduction.objects.filter(
+                name="UK",
+                date__lte=d,
+            ).order_by("date")
+            cops = np.array(map(lambda x: x.production_oil, cps))
+            months = np.array(range(0, len(cops)))
+            r1, k1, p1, t1, r2, k2, p2, t2, residual = fit_cyclic_logistic(
+                months, cops, r1, k1, p1, t1, r2, k2, p2, t2
+            )
+            ukfs.append(k1 - get_logistic(r1, k1, p1)(
+                len(cops) + diff_months_abs(d_max, d) - t1
+            ) + k2 - get_logistic(r2, k2, p2)(
+                len(cops) + diff_months_abs(d_max, d) - t2
+            ))
[…]

### 2014-07-16 tomorrow_django/oil_and_gas/management/commands/hubbert.py

@@ -1,0 +1,11 @@
+from django.core.management.base import BaseCommand
+
+from ...hubbert import HubbertBacktest
+
+
+class Command(BaseCommand):
+    def handle(self, *args, **options):
+        country = "NO"
+        if len(args) > 0 and args[0] == "UK":
+            country = "UK"
+        HubbertBacktest.run(country)

### 2014-07-16 tomorrow_django/oil_and_gas/tests/fit_logistic_tests.py

@@ -17,7 +17,8 @@
 from django.test import TestCase
 from numpy import array
 
-from ..fit_logistic import fit_logistic_r_p, get_logistic, get_d_logistic, fit_d_logistic
+from ..fit_logistic import (fit_logistic_r_p, get_logistic, get_d_logistic,
+                            fit_d_logistic_r_k_p)
 
 
 class FittingTest(TestCase):
@@ -33,7 +34,7 @@ class FittingTest(TestCase):
         func = get_d_logistic(2.0, 3.0, 5.0)
         x = array([0, 1, 2, 3])
         y = [func(0.0), func(1.0), func(2.0), func(3.0)]
-        r, k, p = fit_d_logistic(3.0, x, y)
-        self.assertAlmostEqual(r, 2.0, delta=0.01, msg="r is incorrect")
-        self.assertAlmostEqual(k, 3.0, delta=0.01, msg="p is incorrect")
-        self.assertAlmostEqual(p, 5.0, delta=0.01, msg="p is incorrect")
+        r, k, p, residual = fit_d_logistic_r_k_p(x, y)
+        self.assertAlmostEqual(r, 2.0, delta=0.01, msg="r={0} is incorrect".format(r))
+        self.assertAlmostEqual(k, 3.0, delta=0.01, msg="k={0} is incorrect".format(k))
+        self.assertAlmostEqual(p, 5.0, delta=0.01, msg="p={0} is incorrect".format(p))

### 2014-07-16 tomorrow_django/oil_and_gas/utils.py

@@ -61,7 +61,7 @@ def diff_months_abs(date1, date2):
     return abs(diff_months(date1, date2))
 
 
-def make_plot(x, y, x_label, y_label, x_fit=None, y_fit=None):
+def make_plot(x, y, x_label, y_label, x_fit=None, y_fit=None, title=""):
     fig = plt.figure(figsize=(16,10))
     ax = fig.add_subplot(111)
     ax.plot(x, y, 'ok')
@@ -69,6 +69,15 @@ def make_plot(x, y, x_label, y_label, x_fit=None, y_fit=None):
         ax.plot(x_fit, y_fit, '-k')
     ax.set_xlabel(x_label)
     ax.set_ylabel(y_label)
+    plt.title(title)
+    font = {
+        'family': 'Arial',
+        'weight': 'normal',
+        'size': 28,
+    }
+    matplotlib.rc('font', **font)
+    matplotlib.rcParams['lines.markersize'] = 20
+    matplotlib.rcParams['axes.labelsize'] = 26
     plt.show()
 
 
@@ -86,4 +95,4 @@ def start_plot(x_label, y_label, title):
     matplotlib.rc('font', **font)
     matplotlib.rcParams['lines.markersize'] = 20
     matplotlib.rcParams['axes.labelsize'] = 26
-    return plt
+    return plt, ax, ax.twinx()

### 2014-07-16 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

@@
 to account for multi-cycled or asymmetric production curves.
-The existing forecasts of future oil production use some form of the
+
+\subsection{Back-testing Hubbert's Model}
+
+Despite the popularity of Hubbert model, the literature says little about its the out-of-sample performance.
+There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time.
+As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test has to be started when the peak is already well visible in the data.
+For each time $T$, a month between the start of the back-test and the date of the latest available date, we fitted the Hubbert model to the data truncated at time $T$.
+Based on each fit, we computed the predicted remaining recoverable reserves.
+This back-test allows us to study how the forecast of Hubbert's model evolves over time.
+A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time.
+An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects.
+This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.
+For the studied cases, namely Norway and the UK, a good starting date was chosen to be the 01/2000.
+Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which requires the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks.
+The peak of each of the two logistic functions matches up with one of the two peaks in the UK production data.
+The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time.
+In 2000, for Norway, the Hubbert model would have predicted 2.08Gb remaining recoverable reserves in 2014, while up to now the predicted amount grew to 2.83Gb (a 36\% error) and the growth did not yet slow down.
+For the UK, the prediction would have been 0.21Gb but grew all the way to 1.43Gb (an error of 581\%), with a recent slow down in growth.
+These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast.
+It has to be added that the error in total production since the beginning in the year 2000 relative too now is only 17\% for Norway and 6.5\% for the UK.
+Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.
+The aim is to understand the cause why the Hubbert model underestimates the tail of the production curve and improve upon it.
+A first argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves.
+To discuss this hypothesis the figure \ref{hubbert-backtest} shows as well the oil price.
+The forecast for Norway remains fairly stable and then starts to grow with a small delay when the oil price increases.
+The growth is continuing in the past years, even as the oil price stabilized again.
+For the UK the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in the oil price.
+One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to the oil price as new capacities had first to be build.
+On the other hand, UK could have had available or upcoming spare capacities which allow to respond rapidly to the increase in the oil price.
+Any case, it seems difficult to establish a clear link between the oil price and the increase in forecasted reserves.
+
+\subsection{Beyond Hubbert's Model}
+
+The existing forecasts of future oil production all use some form of the
 Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
-or some economical model applied to aggregate production \citep{Greiner2011},
-but none goes into the details of studying the underlying dynamics.
-The main reason for the lack of details is certainly the lack of available
-data.
+or some economical model applied to aggregate production \citep{Greiner2011}.
+ However, as just shown in the back-test, the Hubbert model does not provide a
+forecast stable in time, usually underestimating the tail in the production curve.
+Considering that none of these Hubbert bases methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
+In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data.
+The Hubbert model will certainly always remain the tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.
 
-In this article, a new methodology is introduced to forecast future
+Nonetheless, in this article, a new methodology is introduced to forecast future
 oil production. Instead of taking the aggregate oil production profile
@@
 it is applied to two major oil producing countries with publicly available
-data: Norway \citepalias{Directorate2014} and the U.K. \citep{GOVUK2014}.
+data at the field level: Norway \citepalias{Directorate2014} and the U.K. \citep{GOVUK2014}.
 
@@
 
+%FIGURE 0
+\noindent 
+\begin{figure}[H]
+\includegraphics[width=1\columnwidth]{hubbert-backtest}
+\caption{A back-test of the forecast stability of the Hubbert model.}
+\label{hubbert-backtest} 
+\end{figure}
+
+\pagebreak
+
 %FIGURE 1
@@
 \end{figure}
+
+\pagebreak
 

### 2014-07-16 tomorrow_django/oil_and_gas/hubbert.py

@@ -16,7 +16,7 @@ class HubbertBacktest():
 
     @classmethod
     def run(cls, name):
-        ds, nks, ukfs, fs = [], [], [], []
+        ds, nks, ukks, ukfs, fs = [], [], [], [], []
         d = date(year=2000, month=1, day=1)
         d_max = date.today() - timedelta(days=120)
         r = -1.0
@@ -57,6 +57,7 @@ class HubbertBacktest():
             r1, k1, p1, t1, r2, k2, p2, t2, residual = fit_cyclic_logistic(
                 months, cops, r1, k1, p1, t1, r2, k2, p2, t2
             )
+            ukks.append(k1 + k2)
             ukfs.append(k1 - get_logistic(r1, k1, p1)(
                 len(cops) + diff_months_abs(d_max, d) - t1
             ) + k2 - get_logistic(r2, k2, p2)(
@@ -78,8 +79,12 @@ class HubbertBacktest():
         plt, ax1, ax2 = start_plot(
             "Back-testing date",
             "Remaining production (barrels)",
-            "Norway's remaining recoverable reserves in 2014\nA back-test using Hubbert"
+            "Remaining recoverable reserves in 2014\nA back-test using Hubbert"
         )
+        print "{0} vs {1}".format(fs[0:2], fs[-2:-1])
+        print "{0} vs {1}".format(ukfs[0:2], ukfs[-2:-1])
+        print "{0} vs {1}".format(ukks[0], ukks[-1])
+        print "{0} vs {1}".format(nks[0], nks[-1])
         l1 = ax1.plot(ds, fs, '-b', label="Forecast Norway")
         l2 = ax1.plot(ds, ukfs, '-g', label="Forecast UK")
         l3 = ax2.plot(ds, OIL_PRICE_SINCE_2000[:len(ds)], '-r', label="Oil price ($)")

### 2014-07-18 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

@@
 
-Despite the popularity of Hubbert model, the literature says little about its the out-of-sample performance.
+Despite the popularity of the Hubbert model, the literature says little about its out-of-sample performance.
 There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time.
-As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test has to be started when the peak is already well visible in the data.
+As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test had to be started when the peak is already well visible in the data.
 For each time $T$, a month between the start of the back-test and the date of the latest available date, we fitted the Hubbert model to the data truncated at time $T$.
-Based on each fit, we computed the predicted remaining recoverable reserves.
-This back-test allows us to study how the forecast of Hubbert's model evolves over time.
+Based on each fit, we computed the predicted remaining recoverable reserves in 2014.\\
+This back-test allowed us to study how the forecast of Hubbert's model evolves over time.
 A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time.
 An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects.
-This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.
+This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.\\
 For the studied cases, namely Norway and the UK, a good starting date was chosen to be the 01/2000.
-Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which requires the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks.
-The peak of each of the two logistic functions matches up with one of the two peaks in the UK production data.
+Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which required the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks.
+The peak of each of the two logistic functions is matched up with one of the two peaks in the UK production data.\\
 The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time.
@@
 These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast.
-It has to be added that the error in total production since the beginning in the year 2000 relative too now is only 17\% for Norway and 6.5\% for the UK.
-Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.
-The aim is to understand the cause why the Hubbert model underestimates the tail of the production curve and improve upon it.
-A first argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves.
+For completeness, it has to be added that the relative error between the year 2000 and now in total production since the beginning is only 17\% for Norway and 6.5\% for the UK.
+Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.\\
+The aim is to understand the causes why the Hubbert model underestimates the tail of the production curve and improve upon it.
+An argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves.
 To discuss this hypothesis the figure \ref{hubbert-backtest} shows as well the oil price.
-The forecast for Norway remains fairly stable and then starts to grow with a small delay when the oil price increases.
+The forecast for Norway remains fairly stable in the beginning and then starts to grow with a small delay when the oil price increases.
 The growth is continuing in the past years, even as the oil price stabilized again.
@@
 On the other hand, UK could have had available or upcoming spare capacities which allow to respond rapidly to the increase in the oil price.
-Any case, it seems difficult to establish a clear link between the oil price and the increase in forecasted reserves.
+While it is certain that increases in the oil price will increase the recoverable reservers, it seems difficult to establish a mathematically defined link between the oil price and the increase in forecasted reserves.
+Consequently, an extension of the Hubbert model integrating the oil price has not been further explored.
 
@@
 forecast stable in time, usually underestimating the tail in the production curve.
-Considering that none of these Hubbert bases methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
+Considering that none of these Hubbert based methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
 In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data.
@@
 up once the field is being developed, and then a peak or plateau before
-the oil field production starts decaying. The decay can take many different shapes
-and is governed by a variety of geological and economical factors.
-The goal of the fitting procedure is to capture as much the impact of
-these different factors as possible.
+the oil field production starts decaying. The decay can take many different shapes and is governed by a variety of geological and economical factors.
+The goal of the fitting procedure is to capture as much as possible the impact of
+these different factors.
 
@@
 To be able to forecast the oil production of each individual field,
-regularity had to be found in the production's dynamics. Modeling
-the whole production profile from the beginning of extraction seems
-elusive due to the variety of the forms it can take. Fortunately,
-modeling the decay process is sufficient in order to extrapolate future
-oil production. A preliminary classification is necessary to achieve
-that goal. Figures \ref{regular}, \ref{irregular} and \ref{new}
-show that, independent of the country, oil fields can be classified
-into three main categories:
+regularity had to be found in the production's dynamics.
+Unfortunately, the whole production profile can take a large variety of shapes with often very asymmetric shapes.
+Therefore, modeling the whole production profile from the beginning of extraction with the Hubbert model seems elusive.
+It would require the use of various extensions of the Hubbert model and for each field the most appropriate extension would need to be determined.
+This is impractical when dealing with hundreds of oil field.\\
+Fortunately, modeling the decay process is sufficient in order to extrapolate future oil production.
+A preliminary classification is necessary to achieve that goal.
+Figures \ref{regular}, \ref{irregular} and \ref{new} show that, independent of the country, oil fields can be classified into three main categories:
 \begin{itemize}
@@
 
-To determine how well the extrapolation based on the stretched exponential
-predicts the future production, a complete back-testing has been performed
-on each field. A single back-test is made as follows:
+For every extrapolation, be it the Hubbert model applied to aggregate production at the country level or the stretched exponential applied to a field's decay, we want to determine how good and stable the predictive power is.
+To do so, we perform a complete back-testing in every case.
+This includes a monthly back-test for the extrapolation, using a stretched-exponential, of the decay process of every oil field.\\
+A single back-test is made as follows:
 \begin{itemize}
-\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field is
-truncated at a certain date in the past $T \in\{0,\,\ldots,\, N\}$, where
+\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field or country is truncated at a certain date in the past $T \in\{0,\,\ldots,\, N\}$, where
 $T$ is the time counted in months since the production start of the field.
@@
 \end{itemize}
-Computing this back-test, for every month in the past since the field production
+Computing this back-test, for every month in the past since the production
 started decaying, yields a plot showing the evolution of the relative
@@
 simulating future oil production was straightforward. For each country,
-the existing oil field productions were extrapolated and the future discoveries were
-simulated. Figure \ref{no_uk_2013} shows the average of 1000 simulations.
+the existing oil field productions were extrapolated and the future discoveries were simulated.
+Figure \ref{no_uk_2013} shows the predicted production resulting from the extrapolation of existing fields and the average of 1000 simulations of future discoveries.
 For each country, the non-symmetric shape of the production dynamics, which
-contradicts the prediction based on Hubbert's standard approach,
-is immediately noticeable.
+contradicts the prediction based on Hubbert's standard approach, is immediately noticeable.
 
 The results in table \ref{forecast-14} show a striking difference
-between the extrapolation of the fit and the Monte-Carlo model forecast. According to the
-fit (extrapolation of aggregate production), Norway's future oil production
-would decay much faster than in the Monte-Carlo case. The remaining
-reserves estimated with the Monte-Carlo methodology are $45\%$ larger
+between the Hubbert model, the extrapolation of the fit and the Monte-Carlo model forecast.
+The preliminary study discussed in the introduction already showed that the Hubbert model forecast was not stable over time and therefore unlikely to be accurate.
+Indeed the results confirm that the Hubbert forecast is smaller then fitting the decaying process and only half of what the Monte-Carlo model predicts.
+This is not surprising, because the Hubbert curve is symmetric and therefore the fitting procedure will be strongly biased by the rapid increase in production before the peak, therefore predicting a rapid decrease.
+Only fitting the decay process using a stretched exponential resolves this problem, because the decay is entirely decoupled from the increase before the peak.
+Consequently, we expect the simple fit of the decay to perform better.\\
+According to the fit (extrapolation of aggregate production), Norway's future oil production would still decay much faster than in the Monte-Carlo case.
+The remaining reserves estimated with the Monte-Carlo methodology are $45\%$ larger
 then the estimate from the fit. This difference originates from two
@@
 starting at the second peak. The difference between the Hubbert-based
-methodology and the Monte-Carlo one is very similar to the Norwegian
+methodology and the fit extrapolation is very similar to the Norwegian
 case. The former underestimates the remaining oil reserves by about
@@
 
-Which of the two models is more trustworthy? Clearly, the implications
+Which of the models is more trustworthy? Clearly, the implications
 in adopting one methodology over the other are significant. The only
@@
 with the oil production of the subsequent 6 years, shows that the
-predictive power of the Monte-Carlo model is significantly better than a simple
-extrapolation of the aggregate past production. The Monte-Carlo model 
-is found essentially right on target, while the extrapolation of past production
-(``fit'') is under-estimating the realised production by 19\% and 16\% respectively
-for Norway and the U.K. The following table \ref{backtest-08}
-summarizes the difference between the two approaches for the back-testing
-period.
+predictive power of the Monte-Carlo model is significantly better than the Hubbert model or a simple extrapolation of the aggregate past production.
+The Monte-Carlo model is found essentially right on target, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K.
+The following table \ref{backtest-08} summarizes the difference between the two approaches for the back-testing period.
 
-While the error of the extrapolation of the past oil production method (``fit'') is not dramatic
-over these six years from 2008 to 2014, the difference between it and the Monte-Carlo
-approach becomes huge from 2014 to 2030, as shown in table \ref{forecast-08}.
+While the error of the extrapolation of the past oil production method (``fit'') is not dramatic over these six years from 2008 to 2014, the difference between it and the Monte-Carlo approach becomes huge from 2014 to 2030, as shown in table \ref{forecast-08}.
 
@@
 \begin{table}[H]
-\caption{Remaining oil reserves until 2030 in barrels 
-predicted by the extrapolation of the  \textbf{Fit}
-of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
-The relative difference between these two predictions 
-is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
+\caption{Remaining oil reserves in billion barrels (Gb) until 2030 predicted by the extrapolation of the  \textbf{Fit} of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
+The relative difference between these two predictions is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
 \centering %
-\begin{tabular}{lllr}
- & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
-\midrule Norway & $3.95\cdot10^{9}$  & $5.72\cdot10^{9}$  & $45\%$\tabularnewline
-U.K. & $1.79\cdot10^{9}$  & $2.98\cdot10^{9}$  & $66\%$ \tabularnewline
+\begin{tabular}{llllr}
+ & Hubbert (Gb) & Fit (Gb) & Model (Gb)  & $\Delta$ \tabularnewline
+\midrule Norway & $2.83$ & $3.95$ & $5.72$ & $45\%$\tabularnewline
+U.K. & $1.43$ & $1.79$  & $2.98$ & $66\%$ \tabularnewline
 \bottomrule  &  &  & \tabularnewline
@@
 \caption{Extrapolation of past oil production (``fit'') and prediction
-using the Monte-Carlo model are used on the data set truncated
+using the Monte-Carlo model are made on the data set truncated
 in 2008. Their forecast for the period 2008-2014 is compared to
-the actual realised production.}
+the actual realised production. All units in billion barrels (Gb).}
 \centering %
-\begin{tabular}{llll}
- & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
-\midrule Norway & $4.05\cdot10^{9}$  & $3.28\cdot10^{9}$  & $3.95\cdot10^{9}$ \tabularnewline
-U.K. & $2.40\cdot10^{9}$  & $2.00\cdot10^{9}$  & $2.28\cdot10^{9}$ \tabularnewline
+\begin{tabular}{lllll}
+ & Actual (Gb) & Hubbert (Gb) & Fit (Gb) & Model (Gb)\tabularnewline
+\midrule Norway & $4.05$  & $3.29$  & $3.28$  & $3.95$ \tabularnewline
+U.K. & $2.40$  & $1.63$  & $2.00$  & $2.28$ \tabularnewline
 \bottomrule  &  &  & \tabularnewline
@@
 \begin{table}[H]
-\caption{Remaining oil reserves forecasted for the period 2014-2030 when using
-the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and 
-the Monte-Carlo model. The relative difference between these two predictions 
+\caption{Remaining oil reserves in billion barrels (Gb) forecasted for the period 2014-2030 when using the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and the Monte-Carlo model.
+The relative difference between these two predictions
 is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
 \centering %
-\begin{tabular}{lllc}
- & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
-\midrule Norway 08 & $0.76\cdot10^{9}$  & $5.20\cdot10^{9}$  & $-584\%$\tabularnewline
-U.K. 08 & $0.76\cdot10^{9}$  & $2.77\cdot10^{9}$  & $-264\%$ \tabularnewline
+\begin{tabular}{llllc}
+ & Hubbert (Gb) & Fit (Gb) & Model (Gb) & $\Delta$ \tabularnewline
+\midrule Norway 08 & $2.21$ & $0.76$ & $5.20$  & $-584\%$\tabularnewline
+U.K. 08 & $0.73$ & $0.76$ & $2.77$ & $-264\%$ \tabularnewline
 \bottomrule  &  &  & \tabularnewline
@@
 \begin{table}[H]
-\caption{Oil import (bbl/day) at a constant consumption of 1.5M bbl/day for
-the U.K. and 0.22M bbl/day for Norway. The import for the E.U. and
-Norway is a lower bound based on the changes in the U.K and Norway. Negative
-numbers for Norway represent exports.}
+\caption{Oil import in million barrels per day (Mb/day) at a constant consumption of 1.5 Mb/day for the U.K.
+and 0.22 Mb/day for Norway.
+The import for the E.U.
+and Norway is a lower bound based on the changes in the U.K and Norway.
+Negative numbers for Norway represent exports.}
 \centering %
@@
  & 2014 & 2020 & 2025 & 2030\tabularnewline
-\midrule Norway & $-1.23\cdot10^{6}$  & $-0.88\cdot10^{6}$  & $-0.58\cdot10^{6}$  & $-0.43\cdot10^{6}$ \tabularnewline
-U.K. & $0.7\cdot10^{6}$  & $0.9\cdot10^{6}$  & $1.0\cdot10^{6}$  & $1.1\cdot10^{6}$ \tabularnewline
-E.U.+Norway & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
+\midrule Norway (Mb/day) & $-1.23$  & $-0.88$  & $-0.58$  & $-0.43$ \tabularnewline
+U.K. (Mb/day) & $0.7$  & $0.9$  & $1.0$  & $1.1$ \tabularnewline
+E.U. + Norway (Mb/day) & $9.8$  & $10.45$  & $10.85$  & $11.1$ \tabularnewline
 \bottomrule  &  &  &  & \tabularnewline

### 2014-07-18 tomorrow_django/oil_and_gas/hubbert.py

@@ -17,7 +17,9 @@ class HubbertBacktest():
     @classmethod
     def run(cls, name):
         ds, nks, ukks, ukfs, fs = [], [], [], [], []
+        ukfs2008, nfs2008 = [], []
         d = date(year=2000, month=1, day=1)
+        d_2008 = date(year=2008, month=1, day=1)
         d_max = date.today() - timedelta(days=120)
         r = -1.0
         k = 20E9
@@ -45,6 +47,12 @@ class HubbertBacktest():
             fs.append(k - get_logistic(r, k, p)(
                 len(cops) + diff_months_abs(d_max, d)
             ))
+            if d <= d_2008:
+                nfs2008.append(get_logistic(r, k, p)(
+                    len(cops) + diff_months_abs(d_max, d)
+                ) - get_logistic(r, k, p)(
+                    len(cops) + diff_months_abs(d_2008, d)
+                ))
             print "{0}: {1}".format(d, k)
 
             # UK
@@ -63,6 +71,16 @@ class HubbertBacktest():
             ) + k2 - get_logistic(r2, k2, p2)(
                 len(cops) + diff_months_abs(d_max, d) - t2
             ))
+            if d <= d_2008:
+                ukfs2008.append(get_logistic(r1, k1, p1)(
+                    len(cops) + diff_months_abs(d_max, d) - t1
+                ) + get_logistic(r2, k2, p2)(
+                    len(cops) + diff_months_abs(d_max, d) - t2
+                ) - get_logistic(r1, k1, p1)(
+                    len(cops) + diff_months_abs(d_2008, d) - t1
+                ) - get_logistic(r2, k2, p2)(
+                    len(cops) + diff_months_abs(d_2008, d) - t2
+                ))
             print "{0}: {1}".format(d, k1 + k2)
 
             d = add_months(d, 1)
@@ -76,19 +94,38 @@ class HubbertBacktest():
         #     get_d_logistic(r1, k1, p1)(months - t1) + get_d_logistic(r2, k2, p2)(months - t2)
         # )
         # make_plot(ds, ks, "Month", "K")
+
+
+        # plt, ax1, ax2 = start_plot(
+        #     "Back-testing date",
+        #     "Remaining production (barrels)",
[…]

### 2014-07-21 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

@@
 barrels for the UK, which are respectively $45\%$ and $66\%$ above the
-predictions using the standard methodology.
+predictions using an extrapolation of aggregate production.
 \end{abstract}
@@
 of the many small yet undiscovered small oil fields. Assuming that the power law (\ref{ccdfsizefield})
-would hold for smaller fields down to size of 1 Million barrels leads to a number of 
+would hold for smaller fields down to size of 1 million barrels leads to a number of 
 such fields 10 to 100 times larger than presently known.
 
-But there is another
-key factor that needs to be considered, namely the fact that small oil fields
-are not economically viable for exploitation, which leads to an effective truncation
-in the distribution (\ref{ccdfsizefield}) relevant for the estimation of recoverable oil.
-Taking the data from the \citetalias{Directorate2014}
-providing the yearly investments broken by fields in Norway, let us consider the illustrative
-case of the field GAUPE. The investment spent to 
-develop its exploitation was  \$380M, while its estimated size is 
-$\simeq 1.2Mb$, which corresponds approximately to a total
-market value of $\$120M$ at \$100 per barrel. It is thus not a surprise that
-investment to exploit this field was interrupted in 2013. 
+But there is another key factor that needs to be considered, namely the fact that small oil fields are not economically viable for exploitation, which leads to an effective truncation in the distribution (\ref{ccdfsizefield}) relevant for the estimation of recoverable oil.
+In 1980, the threshold for economic viability was as high as 70 million barrels \cite[p.
+591]{Smith1980}, but with today's oil price we expect a much lower threshold.
+Taking the data from the \citetalias{Directorate2014} providing the yearly investments broken by fields in Norway, let us consider the illustrative case of the field GAUPE.
+The investment spent to develop its exploitation was  \$380M, while its estimated size is $\simeq 1.2Mb$, which corresponds approximately to a total market value of $\$120M$ at \$100 per barrel.
+It is thus not a surprise that investment to exploit this field was interrupted in 2013.
 Figure \ref{returnNorwayfield} shows the total estimated revenue divided
@@
 of its carrying capacity given by equation (\ref{eq:likelihood}).
+
+A comparison with the results in table 1 of \cite{Smith1980} show that our results a coherent.
+Back in 1980, he used $J=7$ which fortunately maps easily to our splitting using $J=2$ as his first reservoir size is 0 to 50 Mb as ours, and the sum of the 6 others yields our second reservoir size 50 Mb and above.
+His estimated total reserves were 43.2 billion barrels with 90\% confidence bounds at 38.4 and 65.0, across 203 dwarfs and 117 giants in our definition.
+Currently, we know that total estimated reserves are reaching 59 Bb, which is still within the 90\% confidence interval of Smith.
+This strongly speaks in favour of his likelihood technique.
+Given that in the past 34 years there have been major technological improvements and a strong increase in the oil price, it is not surprising that the actual reserves are tending towards his upper bound.\\
+A linear scaling by 59/43.2 (actual reserves over Smith expected reserves) of the number of fields Smith predicted, yields 277 dwarfs and 160 giants.
+This is close to the $296 \pm 21$ dwarfs and $156 \pm 2$ giants our 60\% likelihood estimation found for Norway and the UK combined.
+Such a strong coherence between the two fully independent implementations using different parameters (e.g.
+$J=7$ versus $J=2$) and different data sets (1980 versus 2014) is almost surprising and we see it as a good validation of the approach.
 
@@
 \textquotedbl{}w shape\textquotedbl{} of the oil production profile.
-However, this has not been an issue for us to extrapolate the decay of production
-starting at the second peak. The difference between the Hubbert-based
-methodology and the fit extrapolation is very similar to the Norwegian
-case. The former underestimates the remaining oil reserves by about
-$66\%$ compared to the latter.
+However, this has not been an issue to fit a double logistic (Hubbert model) or extrapolate the decay of aggregate production starting at the second peak.
+The differences between the three methodologies are very similar to the Norwegian case.
+The Hubbert model again predicts the smallest remaining reserves with about half the amount of the Monte-Carlo model.
+The fit of aggregate production underestimates the remaining oil reserves by about $66\%$ compared to the Monte-Carlo model.
 
@@
 long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
-shows the results of these tests. Comparing the forecast of both models,
-with the oil production of the subsequent 6 years, shows that the
-predictive power of the Monte-Carlo model is significantly better than the Hubbert model or a simple extrapolation of the aggregate past production.
-The Monte-Carlo model is found essentially right on target, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K.
-The following table \ref{backtest-08} summarizes the difference between the two approaches for the back-testing period.
-
-While the error of the extrapolation of the past oil production method (``fit'') is not dramatic over these six years from 2008 to 2014, the difference between it and the Monte-Carlo approach becomes huge from 2014 to 2030, as shown in table \ref{forecast-08}.
-
-The simple extrapolation decays too fast and entirely misses the fat
-tails in the decay process of individual fields and the new discoveries.
-Moreover, it must be noted that the simple extrapolation changed massively
-between the back-test in table \ref{forecast-08} and the current
-fit in table \ref{forecast-14} (520\% for Norway and 236\% for the
-U.K.). In other words, the simple extrapolation is very unstable
-in its forecast, while the Monte-Carlo forecasts remains very consistent
-(less than 10\% change).
-
+shows the resulting production curves for the fit of aggregate production and the Monte-Carlo model.\\
+Comparing the forecast of the models with the oil production of the subsequent 6 years, shows that the predictive power of the Monte-Carlo model is significantly better than the Hubbert model or a simple extrapolation of the aggregate past production.
+Table \ref{backtest-08} summarizes the difference between the approaches for the back-testing period.
+The Hubbert model has the poorest performance, underestimating production by 19\% (Norway) and 47\% (UK).
+The Monte-Carlo model is found essentially on target with less then 5\% error on the actual production, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K.
+These results confirm that modelling the production dynamics of the individual fields and the discovery of new fields yields a more accurate prediction then methodologies who just use aggregate production.\\
+The error of the Hubbert model or the extrapolation of the past oil production method (``fit'') is not dramatic over the six years from 2008 to 2014, however the difference between them and the Monte-Carlo approach becomes huge for the time period from 2014 to 2030, as shown in table \ref{forecast-08}.
+The Hubbert model and the simple extrapolation decay too fast and entirely miss the fat tails in the decay process of individual fields and the new discoveries.
+It is interesting to note that the Hubbert model performs similarly then the fit of aggregate production for the UK, but performs much better for Norway.
+This is due to the fact that the Hubbert model calibrates the decay slope based on the increase before the peak, which in the case of Norway turns out to be beneficial.\\
+The simple extrapolation changed massively between the back-test in table \ref{forecast-08} and the current fit in table \ref{forecast-14} (520\% for Norway and 236\% for the U.K.).
+In other words, the simple extrapolation is very unstable in its forecast, while the Monte-Carlo forecasts remains very consistent (less than 10\% change).\\
 As can be seen in figure \ref{model-back-testing}, the actual production
@@
 
-We have presented a Monte-Carlo based methodology to forecast future
-oil production. By extending the oil production of current fields
-into the future and modeling the discovery rate of new fields, the
-future oil production of Norway and the U.K. could be forecasted.
-These forecasts are significantly different from the ones obtained
-with a standard extrapolation. Indeed, our model forecasts 45\% to
-66\% more remaining oil reserves than the standard extrapolation.
-The back-test performed on the time period between 2008 and 2014 confirmed
-that the Monte-Carlo based model better captured the production dynamics.
+The detailed discussion of a full back-test of the Hubbert model for Norway and the UK showed that its prediction were unstable in time and therefore unlikely to be reliable.
+We argue that the Hubbert model is likely to be unable to correctly capture the fat tail in the decay process as it doesn't model the fat tails in the decay of individual fields nor the discovery of new fields.
+An easily quantifiable link with the oil price, which is the driving factor of oil production, could not be found and we therefore rejected the exploration of extensions of the Hubbert model to capture properly the tail.\\
+We preferred a two step Monte-Carlo based methodology to forecast future oil production which: (i) extends the oil production of current fields into the future and (ii) models the discovery rate of new fields.
+This methodology was then applied to forecast the future oil production of Norway and the U.K.
+These forecasts are significantly different from the ones obtained with an extrapolation of aggregate production.
+Indeed, our model forecasts 45\% to 66\% more remaining oil reserves than the standard extrapolation.
+The back-test performed on the time period between 2008 and 2014 confirmed that the Monte-Carlo based model better captured the production dynamics.
 

### 2014-07-21 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

@@
 our back-tests starting in 2008. Specifically, we predict remaining reserves
-extractable until 2030 to be $5.7 \pm 0.3$ million barrels for Norway and $3.0 \pm0.3$ billion
-barrels for the UK, which are respectively $45\%$ and $66\%$ above the
-predictions using an extrapolation of aggregate production.
+extractable until 2030 to be $5.7 \pm 0.3$ billion barrels for Norway and $3.0 \pm0.3$ billion barrels for the UK, which are respectively $45\%$ and $66\%$ above the predictions using an extrapolation of aggregate production.
 \end{abstract}

### 2014-08-14 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

@@
 \usepackage{abstract}% Allows abstract customization
+\usepackage{setspace}
 \renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
@@
 \defcitealias{Directorate2014}{NPD}
+
+\doublespacing
 
@@
 \begin{abstract}
+\doublespacing
 \noindent We present a new Monte-Carlo methodology to forecast
@@
 \end{abstract}
-
-\section{Introduction}
 
@@
 the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
-production has been declining rapidly as well \citep{Hook2008}.\\
+production has been declining rapidly as well \citep{Hook2008}.
+This motivates the novel Monte-Carlo methodology to forecast
+the crude oil production of Norway and the U.K. that we present here.
+
+\section{State of the art}
+
+\subsection{Hubbert's Model for peak oil forecasting}
+
 The methodology behind forecasting future oil production has not
@@
 
-Despite the popularity of the Hubbert model, the literature says little about its out-of-sample performance.
+Despite the popularity of the Hubbert model, the literature says little about its out-of-sample performance,
+apart from the one case that made the prediction famous, namely the prediction by Hubbert of the
+ conventional oil peak that occurred in the US in 1970.
 There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time.
@@
 A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time.
-An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects.
-This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.\\
-For the studied cases, namely Norway and the UK, a good starting date was chosen to be the 01/2000.
-Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which required the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks.
+An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects (technically, it is called ``misspecified'').
+This is of course assuming that the dynamics of the system is deterministic to a higher degree than the error the model is making in its forecast.\\
+For the studied cases, namely Norway and the UK, a good starting date was chosen to be 01/2000.
+Fitting Norway's production could be made using the simple logistic equation \ref{logistic}.
+In contrast, for the UK, the production exhibits two clear peaks, which required the use of a double-logistic function that is the sum of two independent logistic functions with time shifted peaks.
 The peak of each of the two logistic functions is matched up with one of the two peaks in the UK production data.\\
 The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time.
-In 2000, for Norway, the Hubbert model would have predicted 2.08Gb remaining recoverable reserves in 2014, while up to now the predicted amount grew to 2.83Gb (a 36\% error) and the growth did not yet slow down.
+In 2000, for Norway, the Hubbert model would have predicted 2.08Gb of remaining recoverable reserves by 2014 while, up to the time of writing (July 2014), the predicted amount grew to 2.83Gb (a 36\% error) and the growth is not showing sign of slowing down, yet.
 For the UK, the prediction would have been 0.21Gb but grew all the way to 1.43Gb (an error of 581\%), with a recent slow down in growth.
-These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast.
-For completeness, it has to be added that the relative error between the year 2000 and now in total production since the beginning is only 17\% for Norway and 6.5\% for the UK.
+These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast of recoverable reserves.
+For completeness, let us stress that the relative error between the year 2000 and July 2014 in total production since the beginning is only 17\% for Norway and 6.5\% for the UK.
 Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.\\
-The aim is to understand the causes why the Hubbert model underestimates the tail of the production curve and improve upon it.
+Our aim is first to understand why the Hubbert model underestimates the tail of the production curve and then to improve upon it.
 An argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves.
-To discuss this hypothesis the figure \ref{hubbert-backtest} shows as well the oil price.
-The forecast for Norway remains fairly stable in the beginning and then starts to grow with a small delay when the oil price increases.
-The growth is continuing in the past years, even as the oil price stabilized again.
-For the UK the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in the oil price.
-One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to the oil price as new capacities had first to be build.
-On the other hand, UK could have had available or upcoming spare capacities which allow to respond rapidly to the increase in the oil price.
-While it is certain that increases in the oil price will increase the recoverable reservers, it seems difficult to establish a mathematically defined link between the oil price and the increase in forecasted reserves.
+To discuss this hypothesis, figure \ref{hubbert-backtest} shows the oil price as well.
+The forecast for Norway remains fairly stable from 2000 to 2007 during a period when the oil
+price went from slightly above \$20 per barrel to more than \$80 per barrel. The shift to a larger predicted
+remaining production does not correlate with the peak, bottom and then rebound of the oil prices
+since 2007.
+For the UK, the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in oil prices.
+One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to oil prices as new capacities had first to be build.
+On the other hand, UK could have had available or upcoming spare capacities that allowed to respond rapidly to the increase in the oil price.
+While it is certain that increases in oil prices will augment the recoverable reserves, it seems difficult to establish a well-defined mathematical link between oil price and the increase in forecasted reserves.
 Consequently, an extension of the Hubbert model integrating the oil price has not been further explored.
@@
 
-The existing forecasts of future oil production all use some form of the
+The existing forecasts of future oil production all use some version of the
 Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
@@
  However, as just shown in the back-test, the Hubbert model does not provide a
-forecast stable in time, usually underestimating the tail in the production curve.
-Considering that none of these Hubbert based methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
+forecast that is stable in time, usually underestimating the significant tail in the production curve.
+Considering that none of these Hubbert-based methodologies go into the details of studying the underlying dynamics, it is not surprising that important features of the dynamics bearing
+on the total recoverable reserves have been missed.
 In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data.
-The Hubbert model will certainly always remain the tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.
+For this reason, the Hubbert model will likely remain a tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.
 
@@
 
-A comparison with the results in table 1 of \cite{Smith1980} show that our results a coherent.
-Back in 1980, he used $J=7$ which fortunately maps easily to our splitting using $J=2$ as his first reservoir size is 0 to 50 Mb as ours, and the sum of the 6 others yields our second reservoir size 50 Mb and above.
+A comparison with the results in table 1 of \cite{Smith1980} show that our results are coherent.
+Back in 1980, \cite{Smith1980} used $J=7$, which fortunately maps easily to our splitting using $J=2$ as his first reservoir size is 0 to 50 Mb as ours, and the sum of the 6 others yields our second reservoir size 50 Mb and above.
 His estimated total reserves were 43.2 billion barrels with 90\% confidence bounds at 38.4 and 65.0, across 203 dwarfs and 117 giants in our definition.
@@
 This strongly speaks in favour of his likelihood technique.
-Given that in the past 34 years there have been major technological improvements and a strong increase in the oil price, it is not surprising that the actual reserves are tending towards his upper bound.\\
+Given that, in the past 34 years, there have been major technological improvements and a strong increase in the oil price, it is not surprising that the actual reserves are tending towards his upper bound.\\
 A linear scaling by 59/43.2 (actual reserves over Smith expected reserves) of the number of fields Smith predicted, yields 277 dwarfs and 160 giants.
-This is close to the $296 \pm 21$ dwarfs and $156 \pm 2$ giants our 60\% likelihood estimation found for Norway and the UK combined.
+This is close to the $296 \pm 21$ dwarfs and $156 \pm 2$ giants of our 60\% likelihood estimation found for Norway and the UK combined.
 Such a strong coherence between the two fully independent implementations using different parameters (e.g.
-$J=7$ versus $J=2$) and different data sets (1980 versus 2014) is almost surprising and we see it as a good validation of the approach.
+$J=7$ versus $J=2$) and different data sets (1980 versus 2014) provides a credible validation of the approach.
 
@@
 \textquotedbl{}w shape\textquotedbl{} of the oil production profile.
-However, this has not been an issue to fit a double logistic (Hubbert model) or extrapolate the decay of aggregate production starting at the second peak.
+However, this has not been an issue as this inverted \textquotedbl{}w shape\textquotedbl{}
+can be fitted well by a double logistic (Hubbert model).
+Alternatively, we can extrapolate the decay of the aggregate production starting at the second peak.
 The differences between the three methodologies are very similar to the Norwegian case.
 The Hubbert model again predicts the smallest remaining reserves with about half the amount of the Monte-Carlo model.
-The fit of aggregate production underestimates the remaining oil reserves by about $66\%$ compared to the Monte-Carlo model.
+The fit of the aggregate production underestimates the remaining oil reserves by about $66\%$ compared to the Monte-Carlo model.
 
@@
 long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
-shows the resulting production curves for the fit of aggregate production and the Monte-Carlo model.\\
-Comparing the forecast of the models with the oil production of the subsequent 6 years, shows that the predictive power of the Monte-Carlo model is significantly better than the Hubbert model or a simple extrapolation of the aggregate past production.
+shows the resulting production curves for the fit of the aggregate production and of the Monte-Carlo model.\\
+Comparing the forecast of the models with the oil production of the subsequent 6 years shows that the predictive power of the Monte-Carlo model is significantly better than the Hubbert model or than a simple extrapolation of the aggregate past production.
 Table \ref{backtest-08} summarizes the difference between the approaches for the back-testing period.
 The Hubbert model has the poorest performance, underestimating production by 19\% (Norway) and 47\% (UK).
-The Monte-Carlo model is found essentially on target with less then 5\% error on the actual production, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K.
-These results confirm that modelling the production dynamics of the individual fields and the discovery of new fields yields a more accurate prediction then methodologies who just use aggregate production.\\
-The error of the Hubbert model or the extrapolation of the past oil production method (``fit'') is not dramatic over the six years from 2008 to 2014, however the difference between them and the Monte-Carlo approach becomes huge for the time period from 2014 to 2030, as shown in table \ref{forecast-08}.
+The Monte-Carlo model is found essentially on target with less then 5\% error compared with the actual production, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K.
+These results confirm that modelling the production dynamics of the individual fields and the discovery of new fields yields a more accurate prediction then methodologies that just use aggregate production.\\
+The errors of the Hubbert model and of the extrapolation of the past oil production method (``fit'') are not dramatic over the six years from 2008 to 2014.
+However, the difference between them and the Monte-Carlo approach becomes huge for the time period from 2014 to 2030, as shown in table \ref{forecast-08}.
 The Hubbert model and the simple extrapolation decay too fast and entirely miss the fat tails in the decay process of individual fields and the new discoveries.
@@
 The detailed discussion of a full back-test of the Hubbert model for Norway and the UK showed that its prediction were unstable in time and therefore unlikely to be reliable.
-We argue that the Hubbert model is likely to be unable to correctly capture the fat tail in the decay process as it doesn't model the fat tails in the decay of individual fields nor the discovery of new fields.
-An easily quantifiable link with the oil price, which is the driving factor of oil production, could not be found and we therefore rejected the exploration of extensions of the Hubbert model to capture properly the tail.\\
+We argue that the Hubbert model is likely to be unable to correctly capture the fat tail in the decay process as it does not model the fat tails in the decay of individual fields nor the discovery of new fields.
+An easily quantifiable link with the oil price, which is the driving factor of oil production, could not be found and we therefore did not pursue the exploration of extensions of the Hubbert model to capture properly the tails of the decay production process.\\
 We preferred a two step Monte-Carlo based methodology to forecast future oil production which: (i) extends the oil production of current fields into the future and (ii) models the discovery rate of new fields.
@@
 \includegraphics[width=1\columnwidth]{oseberg}
-\caption{Example of the time evolution of the oil production per day of
+\caption{\doublespacing
+Example of the time evolution of the oil production per day of
 a regular field, parametrised beyond the peak in the decay
@@
 \includegraphics[width=1\columnwidth]{oesberg-error}
-\caption{Oesberg field - Relative error defined by expression (\ref{ryjryukoik})
+\caption{\doublespacing
+Oesberg field - Relative error defined by expression (\ref{ryjryukoik})
 of the predicted total production from time $t$ indicated in the abscissa until 2014.
@@
 \includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
-\caption{Logistic fit of the function solution of expression (\ref{logistic})
+\caption{\doublespacing
+Logistic fit of the function solution of expression (\ref{logistic})
 to the number of discoveries for Norway. The discovery
@@
 \includegraphics[width=1\columnwidth]{ccdf-field-size}
-\caption{Complementary 
-cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
+\caption{\doublespacing
+Complementary cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
 The two dashes lines visualise the power law behaviour (\ref{ccdfsizefield}) of the tail
@@
 \includegraphics[width=1\columnwidth]{no-returns-size}
-\caption{Total estimated revenue divided by investment as a function of the estimated 
+\caption{\doublespacing
+Total estimated revenue divided by investment as a function of the estimated 
 ultimate recovery for a number of small oil fields. }
@@
 \includegraphics[width=1\columnwidth]{UK-future} 
-\caption{Monte-Carlo (green upper
-continuous line with standard deviation bank starting in 2014 onward)
-and fit forecast based on past production data (lower line and grey one standard deviation band) for Norway
-(top) and the U.K. (bottom). In both cases, the Monte-Carlo model forecasts
-a significantly slower decay than the fit by taking into account that
-new fields will come in production.}
+\caption{\doublespacing
+Monte-Carlo (green upper continuous line with standard deviation bank starting in 2014 onward) and fit forecast based on past production data (lower line and grey one standard deviation band) for Norway (top) and the U.K.
+(bottom).
+In both cases, the Monte-Carlo model forecasts a significantly slower decay than the fit by taking into account that new fields will come in production.}
 \label{no_uk_2013} 
@@
 \includegraphics[width=1\columnwidth]{NO-back-test}
-\includegraphics[width=1\columnwidth]{UK-back-test} \caption{Monte-Carlo (green upper
-continuous line with standard deviation bank starting in 2014 onward) and Hubbert forecast based 
-on past production data up to 2008 (lower line and grey one standard deviation band)
-for Norway (top) and the U.K. (bottom). The results can be
-compared with the subsequent oil production (blue area). In both cases,
-the Monte-Carlo methodology is more precise.}
+\includegraphics[width=1\columnwidth]{UK-back-test}
+\caption{\doublespacing
+Monte-Carlo (green upper continuous line with standard deviation bank starting in 2014 onward) and Hubbert forecast based on past production data up to 2008 (lower line and grey one standard deviation band) for Norway (top) and the U.K.
+(bottom).
+The results can be compared with the subsequent oil production (blue area).
+In both cases, the Monte-Carlo methodology is more precise.}
 \label{model-back-testing} 
@@
 \begin{table}[h]
-\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.
-For Norway, our logistic fit suggests that up to two new giant fields could be
-ultimately discovered. For the U.K., the prediction is more bleak, suggesting
-that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
+\caption{\doublespacing
+Likelihood estimation for the number of dwarf (1) and giant (2) fields.
+For Norway, our logistic fit suggests that up to two new giant fields could be ultimately discovered.
+For the U.K., the prediction is more bleak, suggesting that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
 \centering %
@@
 \begin{table}[H]
-\caption{Remaining oil reserves in billion barrels (Gb) until 2030 predicted by the extrapolation of the  \textbf{Fit} of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
+\caption{\doublespacing
+Remaining oil reserves in billion barrels (Gb) until 2030 predicted by the extrapolation of the  \textbf{Fit} of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
 The relative difference between these two predictions is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
@@
 \begin{table}[H]
-\caption{Extrapolation of past oil production (``fit'') and prediction
+\caption{\doublespacing
+Extrapolation of past oil production (``fit'') and prediction
 using the Monte-Carlo model are made on the data set truncated
@@
 \begin{table}[H]
-\caption{Remaining oil reserves in billion barrels (Gb) forecasted for the period 2014-2030 when using the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and the Monte-Carlo model.
+\caption{\doublespacing
+Remaining oil reserves in billion barrels (Gb) forecasted for the period 2014-2030 when using the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and the Monte-Carlo model.
 The relative difference between these two predictions
@@
 \begin{table}[H]
-\caption{Oil import in million barrels per day (Mb/day) at a constant consumption of 1.5 Mb/day for the U.K.
+\caption{\doublespacing
+Oil import in million barrels per day (Mb/day) at a constant consumption of 1.5 Mb/day for the U.K.
 and 0.22 Mb/day for Norway.

### 2014-08-14 ABM/sim_abs.py

@@ -10,19 +10,19 @@ class SimABS(SimBase):
         'r': 0.001,
         'y_avg': 1.0,
         'ps': 1000,
-        'v': 0.9,
-        'g': 1.9,
+        'v': 0.995,
+        'g': 2.5,
         'a': 1.0,
         'o': 1.0,
         'b': 2.0,
         'n': 0.99,
         'alpha': 1800,
-        'noise': 25.0,
+        'noise': 10,
     }
     variables = {
         'n1': [],
         'n2': [],
-        'p': [920, 920, 920],
+        'p': [600, 600, 600],
         'U1': [0.0],
         'U2': [0.0],
         'y': [],
@@ -70,17 +70,31 @@ if __name__ == '__main__':
     print sim
     sim.scan_parameters(
         [
-            {"name": "v", "values": list(arange(0.5, 1.01, 0.025))},
-            {"name": "g", "values": list(arange(1.5, 2.51, 0.025))},
+            {"name": "v", "values": list(arange(0.99, 1.000001, 0.0025))},
+            {"name": "g", "values": list(arange(0.5, 2.51, 0.1))},
         ],
-        ["std", "skewness", "kurtosis"]
+        [
+            {"name": "diff_mean", "variables": ("p",)},
+            {"name": "diff_std", "variables": ("p",)},
+            {"name": "diff_skewness", "variables": ("p",)},
+            {"name": "diff_kurtosis", "variables": ("p",)},
+            {"name": "mean", "variables": ("n1",)},
+            {"name": "mean", "variables": ("n2",)},
+            {"name": "diff_correlate", "variables": ("p", "n1")},
+            {"name": "diff_correlate", "variables": ("p", "n2")},
+        ]
     )
-    sim.plot_scan("v", "g", "std")
-    sim.plot_scan("v", "g", "skewness")
-    sim.plot_scan("v", "g", "kurtosis")
-    # sim.run_until("std", 1E-4)
-    # for i in range(0, 100000):
+    sim.plot_scan("v", "g", "p_diff_mean")
+    sim.plot_scan("v", "g", "p_diff_std")
+    sim.plot_scan("v", "g", "p_diff_skewness")
+    sim.plot_scan("v", "g", "p_diff_kurtosis")
+    sim.plot_scan("v", "g", "n1_mean")
+    sim.plot_scan("v", "g", "n2_mean")
+    sim.plot_scan("v", "g", "p_n1_diff_correlate")
+    sim.run_until("std")
[…]

### 2014-08-14 ABM/sim_base.py

@@ -35,11 +35,18 @@ class SimBase(object):
     def hash(self):
         return self.compute_hash(self.parameters, self.initial_variables)
 
-    def __getattr__(self, name, t=-1):
-        m = re.search('(\S*)_(\d*)$', name)
+    @staticmethod
+    def underscore_split(s):
+        one, two = s, None
+        m = re.search('(\S*)_(\d*)$', s)
         if m is not None:
-            name = m.group(1)
-            t = -int(m.group(2))
+            one = m.group(1)
+            two = m.group(2)
+        return one, two
+
+    def __getattr__(self, name, t=-1):
+        name, t = self.underscore_split(name)
+        t = -1 if t is None else -int(t)
         if name in self.variables and len(self.variables[name]) > 0:
             return self.variables[name][t]
         if name in self.parameters:
@@ -62,12 +69,13 @@ class SimBase(object):
                 for key in keys:
                     self.__setattr__(key, new_values[key])
 
-    def run_until(self, name, upper=1E-3):
+    def run_until(self, property, upper=1E-3):
+        p_variable, p_name = property['variables'], property['name']
         old_value, new_value = 0.0, 1.0
         while abs(new_value - old_value) > upper:
             self.run(1000)
             old_value = new_value
-            new_value = self.__getattribute__("diff_{0}".format(name))("p")
+            new_value = self.__getattribute__(p_name)(*p_variable)
         return new_value
 
     def scan_parameters(self, parameters, properties, scan=None):
@@ -112,11 +120,13 @@ class SimBase(object):
                 count += 1
                 scan[name]["values"].append(value)
                 for p in properties:
-                    if p not in scan[name]:
-                        scan[name][p] = {'values': []}
-                    p_value = self.__getattribute__("diff_{0}".format(p))("p")
-                    print "{0}: {1}".format(p, p_value)
-                    scan[name][p]['values'].append(p_value)
+                    p_variable, p_name = p['variables'], p['name']
+                    full_name = "{0}_{1}".format("_".join(p_variable), p_name)
+                    if full_name not in scan[name]:
+                        scan[name][full_name] = {'values': []}
+                    p_value = self.__getattribute__(p_name)(*p_variable)
+                    print "{0}: {1}".format(full_name, p_value)
+                    scan[name][full_name]['values'].append(p_value)
             else:
                 last_count = self.scan_parameters(parameters, properties, scan[name])
[…]
